---
title: "Human-Model Comparison Analysis (Clean)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This script compares computational model performance against human participant judgements on 40 held-out videos, then extends predictions to unseen video sets.

The pre-trained deepfake classifier (`model_au`) requires features scaled using its training data parameters. The LOPO model that predicts *human judgements* is trained on raw (unscaled) features, so it must also predict on raw features.

# 1. Load Libraries

```{r libraries}
library(tidyverse)
library(caret)
library(patchwork)
```

# 2. Load Data

```{r load-data}
dta_cmfts_40_nmf <- read.csv("../human_data_preprocess/Updated-Output/dta_cmfts_40_videos_nmf.csv")
dta_cmfts_40_aus <- read.csv("../human_data_preprocess/Updated-Output/dta_cmfts_40_videos_AUs.csv")

dta_full_stim_human <- read.csv("../human_data_preprocess/Updated-Output/dta_full_stim_clean.csv")
dta_pld_human       <- read.csv("../human_data_preprocess/Updated-Output/dta_pld_clean.csv")

model_nmf <- readRDS("../final_analysis/models/rf_fit_model2_nmf.RDS")
model_au  <- readRDS("../final_analysis/models/rf_fit_model1_AUs.RDS")
scaling_model <- readRDS("../final_analysis/models/scaling_model_AUs.RDS")

dta_classifier_test_new  <- read.csv("../../CSV Files/dta_classifier_test_new.csv")
dta_classifier_train_new <- read.csv("../../CSV Files/dta_classifier_train_new.csv")

dta_classifier_test_new_all  <- read.csv("../../CSV Files/dta_classifier_test_new_all.csv")
dta_classifier_train_new_all <- read.csv("../../CSV Files/dta_classifier_train_new_all.csv")

dta_cmfts_40_aus$isfakeorreal <- factor(dta_cmfts_40_aus$isfakeorreal, 
                                         levels = c("fake", "real"))
dta_classifier_test_new$isfakeorreal <- factor(dta_classifier_test_new$isfakeorreal, levels = c("fake", "real"))
dta_classifier_train_new$isfakeorreal <- factor(dta_classifier_train_new$isfakeorreal, levels = c("fake", "real"))

```

# 3. Standardise Column Names

Rename slider columns once here. Every downstream reference uses `fakeness_slider_response`.

```{r standardise-columns}
dta_full_stim_human <- dta_full_stim_human %>%
  rename(fakeness_slider_response = dfs_fakeness_slider_response)

dta_pld_human <- dta_pld_human %>%
  rename(fakeness_slider_response = dplf_fakeness_slider_response)
```

# 4. Define Scaling Function for model_au

This is ONLY used when feeding data into the pre-trained deepfake classifier (`model_au`). It is NOT used for the LOPO/LOSO human-judgement models.

```{r scaling-setup}
model_au_features <- c(
  "diff1x_pacf5_AU17_chin_raiser",
  "diff1x_pacf5_AU01_inner_brow_raiser",
  "diff2_acf1_AU01_inner_brow_raiser",
  "diff2_acf10_AU12_lip_corner_puller",
  "max_kl_shift_AU12_lip_corner_puller",
  "diff1_acf1_AU12_lip_corner_puller",
  "lumpiness_AU12_lip_corner_puller",
  "diff2_acf1_AU12_lip_corner_puller"
)

fn_scale_for_model_au <- function(df) {
  scaled <- df
  scaled[, model_au_features] <- predict(scaling_model, 
                                         newdata = df[, model_au_features])
  return(scaled)
}

# Sanity check against RAW training CSV (not model_au$trainingData)
test_scaled <- fn_scale_for_model_au(dta_classifier_train_new)
cat("Means (should be ~0):\n")
print(round(colMeans(test_scaled[, model_au_features]), 4))
cat("SDs (should be ~1):\n")
print(round(apply(test_scaled[, model_au_features], 2, sd), 4))
```

# 5. Get Deepfake Classifier Predictions on 40 Videos

```{r classifier-predictions-40}
# nmf model
preds_nmf <- predict(model_nmf, newdata = dta_cmfts_40_nmf)
dta_preds_nmf <- tibble(
  video_filename   = dta_cmfts_40_nmf$video_filename,
  isfakeorreal     = dta_cmfts_40_nmf$isfakeorreal,
  model_prediction = preds_nmf
)

# AU model (scaled)
dta_cmfts_40_aus_scaled <- fn_scale_for_model_au(dta_cmfts_40_aus)
preds_au <- predict(model_au, newdata = dta_cmfts_40_aus_scaled)
dta_preds_au <- tibble(
  video_filename   = dta_cmfts_40_aus_scaled$video_filename,
  isfakeorreal     = dta_cmfts_40_aus_scaled$isfakeorreal,
  model_prediction = preds_au
)
```

------------------------------------------------------------------------

# Analysis 1: Correspondence Analysis

Do the model and humans agree on which videos are real vs fake?

```{r correspondence-function}
fn_correspondence_analysis <- function(model_preds_df, human_data, model_name, condition) {
  
  human_aggregated <- human_data %>%
    group_by(video_filename) %>%
    summarise(mean_rating = mean(fakeness_slider_response, na.rm = TRUE), .groups = "drop") %>%
    mutate(human_prediction = ifelse(mean_rating >= 50, "fake", "real"))
  
  comparison_df <- model_preds_df %>%
    left_join(human_aggregated, by = "video_filename") %>%
    mutate(
      model_prediction = factor(model_prediction, levels = c("real", "fake")),
      human_prediction = factor(human_prediction, levels = c("real", "fake"))
    )
  
  ct <- table(Model = comparison_df$model_prediction, Human = comparison_df$human_prediction)
  chi_test <- chisq.test(ct)
  phi <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
  
  cat("--- Correspondence Analysis ---\n")
  cat("Model:", model_name, "| Condition:", condition, "\n\n")
  print(ct)
  cat("\n")
  print(chi_test)
  cat("Phi coefficient:", round(phi, 4), "\n")
  cat("---------------------------------------\n\n")
}
```

```{r correspondence-run}
fn_correspondence_analysis(dta_preds_nmf, dta_full_stim_human, "NMF Pipeline", "Full Stimuli")
fn_correspondence_analysis(dta_preds_nmf, dta_pld_human, "NMF Pipeline", "PLD")
fn_correspondence_analysis(dta_preds_au, dta_full_stim_human, "AU Pipeline", "Full Stimuli")
fn_correspondence_analysis(dta_preds_au, dta_pld_human, "AU Pipeline", "PLD")
```

------------------------------------------------------------------------

# Analysis 2: Predicting Human Judgements (LOSO & LOPO)

Can the model's features predict how a human will judge a video?

```{r predict-judgements-function}
fn_predict_human_judgements <- function(features_df, human_df, model_name, condition) {
  
  analysis_df <- human_df %>%
    mutate(human_judgement = factor(ifelse(fakeness_slider_response >= 50, "fake", "real"))) %>%
    left_join(features_df, by = "video_filename") %>%
    na.omit()
  
  predictors <- names(features_df)[!names(features_df) %in% c("video_filename", "isfakeorreal")]
  model_formula <- as.formula(paste("human_judgement ~", paste(predictors, collapse = " + ")))
  
  cat("--- Predicting Human Judgements ---\n")
  cat("Model Features:", model_name, "| Human Condition:", condition, "\n\n")
  
  # LOSO
  cat("--- LOSO ---\n")
  loso_folds <- groupKFold(analysis_df$video_filename,
                           k = length(unique(analysis_df$video_filename)))
  set.seed(12345)
  loso_fit <- train(model_formula, data = analysis_df, method = "rf",
                    trControl = trainControl(method = "cv", index = loso_folds),
                    tuneLength = 1)
  print(loso_fit)
  
  # LOPO
  cat("\n--- LOPO ---\n")
  lopo_folds <- groupKFold(analysis_df$participant_id,
                           k = length(unique(analysis_df$participant_id)))
  set.seed(12345)
  lopo_fit <- train(model_formula, data = analysis_df, method = "rf",
                    trControl = trainControl(method = "cv", index = lopo_folds),
                    tuneLength = 1)
  print(lopo_fit)
  cat("-----------------------------------------\n\n")
}
```

```{r predict-judgements-run}
fn_predict_human_judgements(dta_cmfts_40_nmf, dta_full_stim_human, "NMF Pipeline", "Full Stimulus")
fn_predict_human_judgements(dta_cmfts_40_nmf, dta_pld_human, "NMF Pipeline", "PLD")
fn_predict_human_judgements(dta_cmfts_40_aus, dta_full_stim_human, "Direct AU Pipeline", "Full Stimulus")
fn_predict_human_judgements(dta_cmfts_40_aus, dta_pld_human, "Direct AU Pipeline", "PLD")
```

------------------------------------------------------------------------

# Analysis 3: Predicting Human Accuracy

Can the model's features predict when a human will be correct or incorrect?

```{r predict-accuracy-function}
fn_predict_human_accuracy <- function(features_df, human_df, model_name, condition) {
  
  analysis_df <- human_df %>%
    mutate(human_judgement = factor(ifelse(fakeness_slider_response >= 50, "fake", "real"))) %>%
    mutate(human_is_correct = factor(ifelse(human_judgement == isfakeorreal,
                                            "correct", "incorrect"))) %>%
    left_join(features_df, by = "video_filename") %>%
    na.omit()
  
  predictors <- names(features_df)[!names(features_df) %in% c("video_filename", "isfakeorreal")]
  
  cat("--- Predicting Human Accuracy ---\n")
  cat("Model Features:", model_name, "| Human Condition:", condition, "\n\n")
  
  # LOPO (trial-level)
  cat("--- LOPO ---\n")
  lopo_folds <- groupKFold(analysis_df$participant_id,
                           k = length(unique(analysis_df$participant_id)))
  set.seed(12345)
  lopo_fit <- train(
    as.formula(paste("human_is_correct ~", paste(predictors, collapse = " + "))),
    data = analysis_df, method = "rf",
    trControl = trainControl(method = "cv", index = lopo_folds),
    tuneLength = 1
  )
  print(lopo_fit)
  
  # LOSO (aggregated per video accuracy)
  cat("\n--- LOSO ---\n")
  analysis_df_loso <- human_df %>%
    mutate(human_judgement = factor(ifelse(fakeness_slider_response >= 50, "fake", "real"))) %>%
    mutate(trial_is_correct = ifelse(human_judgement == isfakeorreal, 1, 0)) %>%
    group_by(video_filename) %>%
    summarise(human_accuracy = mean(trial_is_correct, na.rm = TRUE), .groups = "drop") %>%
    left_join(features_df, by = "video_filename") %>%
    na.omit()
  
  loso_folds <- groupKFold(analysis_df_loso$video_filename,
                           k = length(unique(analysis_df_loso$video_filename)))
  set.seed(12345)
  loso_fit <- train(
    as.formula(paste("human_accuracy ~", paste(predictors, collapse = " + "))),
    data = analysis_df_loso, method = "rf",
    trControl = trainControl(method = "cv", index = loso_folds),
    tuneLength = 1
  )
  print(loso_fit)
  cat("---------------------------------\n\n")
}
```

```{r predict-accuracy-run}
fn_predict_human_accuracy(dta_cmfts_40_nmf, dta_full_stim_human, "NMF Pipeline", "Full Stimulus")
fn_predict_human_accuracy(dta_cmfts_40_nmf, dta_pld_human, "NMF Pipeline", "PLD")
fn_predict_human_accuracy(dta_cmfts_40_aus, dta_full_stim_human, "Direct AU Pipeline", "Full Stimulus")
fn_predict_human_accuracy(dta_cmfts_40_aus, dta_pld_human, "Direct AU Pipeline", "PLD")
```

------------------------------------------------------------------------

# Analysis 4: Correctness Correspondence

Do the model and humans make mistakes on the same videos?

```{r correctness-correspondence-function}
fn_correctness_correspondence <- function(model_preds_df, human_data, model_name, condition) {
  
  model_correctness <- model_preds_df %>%
    mutate(model_is_correct = ifelse(model_prediction == isfakeorreal, "correct", "incorrect"))
  
  human_correctness <- human_data %>%
    mutate(human_judgement = ifelse(fakeness_slider_response >= 50, "fake", "real")) %>%
    mutate(trial_is_correct = ifelse(human_judgement == isfakeorreal, 1, 0)) %>%
    group_by(video_filename) %>%
    summarise(human_accuracy = mean(trial_is_correct, na.rm = TRUE), .groups = "drop") %>%
    mutate(human_is_correct = ifelse(human_accuracy > 0.5, "correct", "incorrect"))
  
  comparison_df <- model_correctness %>%
    left_join(human_correctness, by = "video_filename")
  
  ct <- table(Model = comparison_df$model_is_correct, Human = comparison_df$human_is_correct)
  chi_test <- chisq.test(ct)
  phi <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
  
  cat("--- Correctness Correspondence ---\n")
  cat("Model:", model_name, "| Condition:", condition, "\n\n")
  print(ct)
  cat("\n")
  print(chi_test)
  cat("Phi coefficient:", round(phi, 4), "\n")
  cat("---------------------------------\n\n")
}
```

```{r correctness-correspondence-run}
fn_correctness_correspondence(dta_preds_nmf, dta_full_stim_human, "NMF Pipeline", "Full Stimulus")
fn_correctness_correspondence(dta_preds_nmf, dta_pld_human, "NMF Pipeline", "PLD")
fn_correctness_correspondence(dta_preds_au, dta_full_stim_human, "Direct AU Pipeline", "Full Stimulus")
fn_correctness_correspondence(dta_preds_au, dta_pld_human, "Direct AU Pipeline", "PLD")
```

------------------------------------------------------------------------

# Human Accuracy

```{r human-accuracy}
# Full Stimuli
dta_full_stim_human <- dta_full_stim_human %>%
  mutate(
    human_prediction = ifelse(fakeness_slider_response >= 50, "fake", "real"),
    is_correct = human_prediction == isfakeorreal
  )
full_stim_acc <- round(mean(dta_full_stim_human$is_correct, na.rm = TRUE) * 100, 2)
cat("Full Stimuli accuracy:", full_stim_acc, "%\n")

# PLD
dta_pld_human <- dta_pld_human %>%
  mutate(
    human_prediction = ifelse(fakeness_slider_response >= 50, "fake", "real"),
    is_correct = human_prediction == isfakeorreal
  )
pld_acc <- round(mean(dta_pld_human$is_correct, na.rm = TRUE) * 100, 2)
pld_n   <- sum(!is.na(dta_pld_human$is_correct))
cat("PLD accuracy:", pld_acc, "%\n")

# 95% CI for PLD
pld_p  <- pld_acc / 100
pld_se <- sqrt(pld_p * (1 - pld_p) / length(unique(dta_pld_human$video_filename)))
cat(sprintf("PLD 95%% CI (per-video n=40): [%.1f%%, %.1f%%]\n",
            (pld_p - 1.96 * pld_se) * 100,
            (pld_p + 1.96 * pld_se) * 100))

```

------------------------------------------------------------------------

# Correctness Correspondence

```{r}
fn_correctness_correspondence(dta_preds_nmf, dta_full_stim_human, "NMF Pipeline", "Full Stimulus")
fn_correctness_correspondence(dta_preds_nmf, dta_pld_human,       "NMF Pipeline", "PLD")
fn_correctness_correspondence(dta_preds_au,  dta_full_stim_human, "Direct AU Pipeline", "Full Stimulus")
fn_correctness_correspondence(dta_preds_au,  dta_pld_human,       "Direct AU Pipeline", "PLD")
```

```         
--- Correctness Correspondence ---
Model: Direct AU Pipeline | Condition: PLD 

           Human
Model       correct incorrect
  correct        13         8
  incorrect       7        12


    Pearson's Chi-squared test with Yates' continuity correction

data:  ct
X-squared = 1.604, df = 1, p-value = 0.2053

Phi coefficient: 0.2003 
---------------------------------
```

Emotion subset correspondence analysis:

```{r}
add_emotion_column <- function(df) {
  df %>%
    mutate(
      emotion = case_when(
        grepl("laughing|happy|angry|disgusted|serious", 
              video_filename, ignore.case = TRUE) ~ "y",
        TRUE ~ "n"
      ),
      emotion = factor(emotion, levels = c("y", "n"))
    )
}

dta_cmfts_40_nmf <- add_emotion_column(dta_cmfts_40_nmf)
dta_cmfts_40_aus <- add_emotion_column(dta_cmfts_40_aus)

# Lookup table for joining to predictions
emotion_lookup <- dta_cmfts_40_aus %>%
  select(video_filename, emotion) %>%
  distinct()

cat("Emotion split in 40-video set:\n")
print(table(emotion_lookup$emotion))
```

```{r}
fn_correspondence_subset <- function(model_preds_df, human_data, emotion_val, label) {
  
  videos_subset <- emotion_lookup %>% filter(emotion == emotion_val)
  
  model_subset <- model_preds_df %>%
    filter(video_filename %in% videos_subset$video_filename) %>%
    mutate(model_prediction = factor(model_prediction, levels = c("real", "fake")))
  
  human_subset <- human_data %>%
    filter(video_filename %in% videos_subset$video_filename) %>%
    group_by(video_filename) %>%
    summarise(mean_rating = mean(fakeness_slider_response, na.rm = TRUE), .groups = "drop") %>%
    mutate(human_prediction = factor(
      ifelse(mean_rating >= 50, "fake", "real"), levels = c("real", "fake")
    ))
  
  comparison <- model_subset %>%
    left_join(human_subset, by = "video_filename") %>%
    mutate(model_prediction = factor(model_prediction, levels = c("real", "fake")),
           human_prediction  = factor(human_prediction,  levels = c("real", "fake")))
  
  ct       <- table(Model = comparison$model_prediction, Human = comparison$human_prediction)
  chi_test <- chisq.test(ct)
  phi      <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
  agreement_pct <- round(100 * sum(diag(ct)) / sum(ct), 1)
  
  cat("---", label, "(n =", nrow(comparison), ") ---\n")
  print(ct)
  cat(sprintf("\nAgreement: %.1f%%\n", agreement_pct))
  print(chi_test)
  cat(sprintf("Phi coefficient: %.3f\n\n", phi))
}

# Emotion videos (n=16)
fn_correspondence_subset(dta_preds_au, dta_pld_human, "y", "EMOTION videos")

# No Emotion videos (n=24)
fn_correspondence_subset(dta_preds_au, dta_pld_human, "n", "NO EMOTION videos")
```

```         
--- EMOTION videos (n = 16 ) ---
      Human
Model  real fake
  real    7    1
  fake    2    6

Agreement: 81.2%

    Pearson's Chi-squared test with Yates' continuity correction

data:  ct
X-squared = 4.0635, df = 1, p-value = 0.04382

Phi coefficient: 0.504

--- NO EMOTION videos (n = 24 ) ---
      Human
Model  real fake
  real   11    0
  fake   11    2

Agreement: 54.2%

    Pearson's Chi-squared test with Yates' continuity correction

data:  ct
X-squared = 0.38144, df = 1, p-value = 0.5368

Phi coefficient: 0.126
```

n=94 and n=464 agreement matrices stats:

```{r}
# ct_94 <- table(
#   Model = factor(dta_results_94$model_prediction, levels = c("real","fake")),
#   Human = factor(dta_results_94$predicted_human,  levels = c("real","fake"))
# )
# chi_94 <- chisq.test(ct_94)
# phi_94 <- as.numeric(sqrt(chi_94$statistic / sum(ct_94)))
# cat("n=94 Model vs Predicted Human:\n")
# print(ct_94)
# print(chi_94)
# cat(sprintf("Phi: %.3f\n\n", phi_94))
# 
# # n=464
# ct_464 <- table(
#   Model = factor(dta_results_464$model_prediction, levels = c("real","fake")),
#   Human = factor(dta_results_464$predicted_human,  levels = c("real","fake"))
# )
# chi_464 <- chisq.test(ct_464)
# phi_464 <- as.numeric(sqrt(chi_464$statistic / sum(ct_464)))
# cat("n=464 Model vs Predicted Human:\n")
# print(ct_464)
# print(chi_464)
# cat(sprintf("Phi: %.3f\n\n", phi_464))
```

Fix:

```{r}
# # Use identical logic to make_agreement_matrix to ensure consistency
# compute_agreement_stats <- function(pred1, pred2, label) {
#   ct <- table(
#     Model = factor(pred1, levels = c("real", "fake")),
#     Human = factor(pred2, levels = c("real", "fake"))
#   )
#   chi_test <- chisq.test(ct)
#   phi      <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
#   
#   cat("---", label, "---\n")
#   print(ct)
#   print(chi_test)
#   cat(sprintf("Phi: %.3f\n\n", phi))
# }
# 
# # n=94
# compute_agreement_stats(
#   as.character(dta_results_94$model_prediction),
#   as.character(dta_results_94$predicted_human),
#   "n=94 Model vs Predicted Human"
# )
# 
# # n=464
# compute_agreement_stats(
#   as.character(dta_results_464$model_prediction),
#   as.character(dta_results_464$predicted_human),
#   "n=464 Model vs Predicted Human"
# )
```

# Figure 4: LOPO Model & Extended Predictions

## Train LOPO model (PLD condition, AU features)

This model learns to predict human judgements from AU features. It is trained on RAW (unscaled) features, because the features were raw when joined from `dta_cmfts_40_aus`. Therefore it must predict on raw features too.

```{r train-lopo-model}
# training data: human trials joined with RAW AU features
dta_lopo_training <- dta_pld_human %>%
  mutate(human_judgement = factor(
    ifelse(fakeness_slider_response >= 50, "fake", "real"),
    levels = c("real", "fake")
  )) %>%
  left_join(dta_cmfts_40_aus, by = "video_filename") %>%
  na.omit()

# Clean up duplicate isfakeorreal columns from join
dta_lopo_training <- dta_lopo_training %>%
  rename(isfakeorreal = isfakeorreal.x) %>%
  select(-isfakeorreal.y)

cat("LOPO training data: ", nrow(dta_lopo_training), " trials,",
    length(unique(dta_lopo_training$participant_id)), "participants,",
    length(unique(dta_lopo_training$video_filename)), "videos\n")

# Formula using the 8 important features
lopo_formula <- as.formula(paste(
  "human_judgement ~",
  paste(model_au_features, collapse = " + ")
))

# LOPO cross-validation
lopo_folds <- groupKFold(
  dta_lopo_training$participant_id,
  k = length(unique(dta_lopo_training$participant_id))
)

lopo_control <- trainControl(
  method = "cv",
  index = lopo_folds,
  savePredictions = "final",
  classProbs = TRUE
)

# UNCOMMENT IF WANT TO TRAIN FROM SCRATCH, ELSE LOAD PRE-TRAIED
# set.seed(12345)
# lopo_fit_au_pld <- train(
#   lopo_formula,
#   data = dta_lopo_training,
#   method = "rf",
#   trControl = lopo_control,
#   tuneLength = 1
# )

lopo_fit_au_pld <- readRDS("lopo_fit_au_pld.RDS")

# Save model
# Training this will result in slightly different results each time, despite set.seed
# This is because caret is a bit weird when parallel processing is used
saveRDS(lopo_fit_au_pld, "lopo_fit_au_pld.RDS")

cat("\n=== LOPO Model Results ===\n")
print(lopo_fit_au_pld)

cat("\n=== Confusion Matrix ===\n")
conf_matrix <- confusionMatrix(
  lopo_fit_au_pld$pred$pred,
  lopo_fit_au_pld$pred$obs,
  positive = "fake"
)
print(conf_matrix)
```

```         
Accuracy   Kappa    
  0.6041667  0.1750828
```

```         
          Reference
Prediction real fake
      real 1518 1025
      fake  392  585
                                         
               Accuracy : 0.5974         
                 95% CI : (0.581, 0.6137)
    No Information Rate : 0.5426         
    P-Value [Acc > NIR] : 3.087e-11      
                                         
                  Kappa : 0.1632         
                                         
 Mcnemar's Test P-Value : < 2.2e-16      
                                         
            Sensitivity : 0.3634         
            Specificity : 0.7948         
         Pos Pred Value : 0.5988         
         Neg Pred Value : 0.5969         
             Prevalence : 0.4574         
         Detection Rate : 0.1662         
   Detection Prevalence : 0.2776         
      Balanced Accuracy : 0.5791         
                                         
       'Positive' Class : fake 
```

The LOPO model was trained on RAW features, so we predict on RAW features. `model_au` was trained on SCALED features, so we scale for that.

```{r apply-lopo-to-test-sets}
# LOPO predictions (RAW features)
predicted_human_94 <- predict(lopo_fit_au_pld, newdata = dta_classifier_test_new)

# model_au predictions (SCALED features)
dta_test_94_scaled <- fn_scale_for_model_au(dta_classifier_test_new)
model_preds_94     <- predict(model_au, newdata = dta_test_94_scaled)

# Results dataframe
dta_results_94 <- tibble(
  video_filename          = dta_classifier_test_new$video_filename,
  actual_authenticity     = factor(dta_classifier_test_new$isfakeorreal, levels = c("real", "fake")),
  predicted_human         = predicted_human_94,
  model_prediction        = model_preds_94
) %>%
  mutate(
    predicted_human_correct = predicted_human == actual_authenticity,
    model_correct           = model_prediction == actual_authenticity
  )

# Accuracy
pred_human_acc_94 <- mean(dta_results_94$predicted_human_correct) * 100
n_94 <- nrow(dta_results_94)
se_94 <- sqrt((pred_human_acc_94/100 * (1 - pred_human_acc_94/100)) / n_94)
ci_low_94  <- pred_human_acc_94 - 1.96 * se_94 * 100
ci_high_94 <- pred_human_acc_94 + 1.96 * se_94 * 100

model_acc_94 <- mean(dta_results_94$model_correct) * 100
se_model_94  <- sqrt((model_acc_94/100 * (1 - model_acc_94/100)) / n_94)
ci_low_model_94  <- model_acc_94 - 1.96 * se_model_94 * 100
ci_high_model_94 <- model_acc_94 + 1.96 * se_model_94 * 100

cat("=== n=94 Test Set ===\n")
cat(sprintf("Predicted human accuracy: %.2f%% [%.2f%%, %.2f%%]\n",
            pred_human_acc_94, ci_low_94, ci_high_94))
cat(sprintf("Model accuracy: %.2f%% [%.2f%%, %.2f%%]\n",
            model_acc_94, ci_low_model_94, ci_high_model_94))

# ============================================================
# n=464 combined set
# ============================================================

dta_combined_464 <- bind_rows(dta_classifier_train_new, dta_classifier_test_new)

# LOPO predictions (RAW features)
predicted_human_464 <- predict(lopo_fit_au_pld, newdata = dta_combined_464)

# model_au predictions (SCALED features)
dta_combined_464_scaled <- fn_scale_for_model_au(dta_combined_464)
model_preds_464         <- predict(model_au, newdata = dta_combined_464_scaled)

# Results dataframe
dta_results_464 <- tibble(
  video_filename          = dta_combined_464$video_filename,
  actual_authenticity     = factor(dta_combined_464$isfakeorreal, levels = c("real", "fake")),
  predicted_human         = predicted_human_464,
  model_prediction        = model_preds_464
) %>%
  mutate(
    predicted_human_correct = predicted_human == actual_authenticity,
    model_correct           = model_prediction == actual_authenticity
  )

# Accuracy
pred_human_acc_464 <- mean(dta_results_464$predicted_human_correct) * 100
n_464 <- nrow(dta_results_464)
se_464 <- sqrt((pred_human_acc_464/100 * (1 - pred_human_acc_464/100)) / n_464)
ci_low_464  <- pred_human_acc_464 - 1.96 * se_464 * 100
ci_high_464 <- pred_human_acc_464 + 1.96 * se_464 * 100

model_acc_464 <- mean(dta_results_464$model_correct) * 100
se_model_464  <- sqrt((model_acc_464/100 * (1 - model_acc_464/100)) / n_464)
ci_low_model_464  <- model_acc_464 - 1.96 * se_model_464 * 100
ci_high_model_464 <- model_acc_464 + 1.96 * se_model_464 * 100

cat("\n=== n=464 Combined Set ===\n")
cat(sprintf("Predicted human accuracy: %.2f%% [%.2f%%, %.2f%%]\n",
            pred_human_acc_464, ci_low_464, ci_high_464))
cat(sprintf("Model accuracy: %.2f%% [%.2f%%, %.2f%%]\n",
            model_acc_464, ci_low_model_464, ci_high_model_464))
```

------------------------------------------------------------------------

```         
=== n=94 Test Set ===
Predicted human accuracy: 49.46% [39.30%, 59.62%]
Model accuracy: 63.44% [53.65%, 73.23%]

=== n=464 Combined Set ===
Predicted human accuracy: 53.78% [49.24%, 58.32%]
Model accuracy: 92.66% [90.28%, 95.03%]
```

The model accuracy is high over 90% because the model is seeing its own training data for this combined set. This one is just for visualisaiton purposes only, so the model bar can be replaced or rmoved.

```{r}
ct_94 <- table(
  Model = factor(dta_results_94$model_prediction, levels = c("real","fake")),
  Human = factor(dta_results_94$predicted_human,  levels = c("real","fake"))
)
chi_94 <- chisq.test(ct_94)
phi_94 <- as.numeric(sqrt(chi_94$statistic / sum(ct_94)))
cat("n=94 Model vs Predicted Human:\n")
print(ct_94)
print(chi_94)
cat(sprintf("Phi: %.3f\n\n", phi_94))

# n=464
ct_464 <- table(
  Model = factor(dta_results_464$model_prediction, levels = c("real","fake")),
  Human = factor(dta_results_464$predicted_human,  levels = c("real","fake"))
)
chi_464 <- chisq.test(ct_464)
phi_464 <- as.numeric(sqrt(chi_464$statistic / sum(ct_464)))
cat("n=464 Model vs Predicted Human:\n")
print(ct_464)
print(chi_464)
cat(sprintf("Phi: %.3f\n\n", phi_464))
```

Fix:

```{r}
# Use identical logic to make_agreement_matrix to ensure consistency
compute_agreement_stats <- function(pred1, pred2, label) {
  ct <- table(
    Model = factor(pred1, levels = c("real", "fake")),
    Human = factor(pred2, levels = c("real", "fake"))
  )
  chi_test <- chisq.test(ct)
  phi      <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
  
  cat("---", label, "---\n")
  print(ct)
  print(chi_test)
  cat(sprintf("Phi: %.3f\n\n", phi))
}

# n=94
compute_agreement_stats(
  as.character(dta_results_94$model_prediction),
  as.character(dta_results_94$predicted_human),
  "n=94 Model vs Predicted Human"
)

# n=464
compute_agreement_stats(
  as.character(dta_results_464$model_prediction),
  as.character(dta_results_464$predicted_human),
  "n=464 Model vs Predicted Human"
)
```

```         
--- n=94 Model vs Predicted Human ---
      Human
Model  real fake
  real   29    6
  fake   49    9

    Pearson's Chi-squared test with Yates' continuity correction

data:  ct
X-squared = 3.0369e-30, df = 1, p-value = 1

Phi: 0.000

--- n=464 Model vs Predicted Human ---
      Human
Model  real fake
  real  178   42
  fake  180   63

    Pearson's Chi-squared test with Yates' continuity correction

data:  ct
X-squared = 2.6988, df = 1, p-value = 0.1004

Phi: 0.076
```

# Figure 4 Panels

## Prepare plotting dataframes

```{r prepare-plot-data}
# row 1
dta_plot_row1 <- dta_pld_human %>%
  group_by(video_filename, isfakeorreal) %>%
  summarise(
    mean_slider = mean(fakeness_slider_response, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    human_prediction = ifelse(mean_slider >= 50, "fake", "real"),
    human_is_correct = ifelse(human_prediction == isfakeorreal, "Correct", "Wrong")
  ) %>%
  left_join(dta_cmfts_40_aus, by = "video_filename") %>%
  mutate(
    true_class = isfakeorreal.x,
    judgement_type = case_when(
      human_prediction == "fake" & true_class == "fake" ~ "Fake Correct (TP)",
      human_prediction == "real" & true_class == "real" ~ "Real Correct (TN)",
      human_prediction == "real" & true_class == "fake" ~ "Fake Wrong (FN)",
      human_prediction == "fake" & true_class == "real" ~ "Real Wrong (FP)"
    )
  ) %>%
  na.omit()

# row 2

dta_plot_row2 <- dta_results_94 %>%
  left_join(dta_classifier_test_new, by = "video_filename") %>%
  mutate(
    true_class = as.character(actual_authenticity),
    judgement_type = case_when(
      predicted_human == "fake" & true_class == "fake" ~ "Fake Correct (TP)",
      predicted_human == "real" & true_class == "real" ~ "Real Correct (TN)",
      predicted_human == "real" & true_class == "fake" ~ "Fake Wrong (FN)",
      predicted_human == "fake" & true_class == "real" ~ "Real Wrong (FP)"
    )
  ) %>%
  na.omit()

# row 3

dta_plot_row3 <- dta_results_464 %>%
  left_join(dta_combined_464, by = "video_filename") %>%
  mutate(
    true_class = as.character(actual_authenticity),
    judgement_type = case_when(
      predicted_human == "fake" & true_class == "fake" ~ "Fake Correct (TP)",
      predicted_human == "real" & true_class == "real" ~ "Real Correct (TN)",
      predicted_human == "real" & true_class == "fake" ~ "Fake Wrong (FN)",
      predicted_human == "fake" & true_class == "real" ~ "Real Wrong (FP)"
    )
  ) %>%
  na.omit()
```

## Consistent y-axis limits

```{r y-axis-limits}
au17_range <- range(c(
  dta_plot_row1$diff1x_pacf5_AU17_chin_raiser,
  dta_plot_row2$diff1x_pacf5_AU17_chin_raiser,
  dta_plot_row3$diff1x_pacf5_AU17_chin_raiser
), na.rm = TRUE)
au17_limits <- au17_range + c(-1, 1) * diff(au17_range) * 0.05

au01_range <- range(c(
  dta_plot_row1$diff1x_pacf5_AU01_inner_brow_raiser,
  dta_plot_row2$diff1x_pacf5_AU01_inner_brow_raiser,
  dta_plot_row3$diff1x_pacf5_AU01_inner_brow_raiser
), na.rm = TRUE)
au01_limits <- au01_range + c(-1, 1) * diff(au01_range) * 0.05
```

## Common theme

```{r common-theme}
fig4_theme <- theme_classic(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = margin(5, 10, 5, 10),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

## Panel A/E/I: Accuracy bar charts

```{r fig4-accuracy-bars, fig.width=4, fig.height=4}
make_accuracy_bar <- function(human_acc, human_ci_low, human_ci_high,
                              model_acc, model_ci_low, model_ci_high,
                              human_label, title_text) {
  
  plot_data <- data.frame(
    Method   = factor(c(human_label, "Random Forest"),
                      levels = c(human_label, "Random Forest")),
    Accuracy = c(human_acc, model_acc),
    CI_low   = c(human_ci_low, model_ci_low),
    CI_high  = c(human_ci_high, model_ci_high)
  )
  
  ggplot(plot_data, aes(x = Method, y = Accuracy, fill = Method)) +
    geom_col(width = 0.6, color = "black") +
    geom_errorbar(aes(ymin = CI_low, ymax = CI_high),
                  width = 0.2, linewidth = 0.8, color = "black") +
    geom_hline(yintercept = 50, linetype = "dashed", linewidth = 0.6) +
    scale_y_continuous(limits = c(0, 100), expand = expansion(mult = c(0, 0.08))) +
    scale_fill_manual(values = setNames(c("#009E73", "#E69F00"),
                                        c(human_label, "Random Forest"))) +
    labs(title = title_text, x = NULL, y = "Accuracy (%)", fill = NULL) +
    theme_classic(base_size = 14) +
    theme(
      legend.position = "none",
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text = element_text(size = 12, color = "black"),
      plot.title = element_text(face = "bold", hjust = 0.5)
    )
}

# Panel A: Actual human vs model (n=40)
# PLD accuracy from the human accuracy section above
pld_accuracy <- round(mean(dta_pld_human$is_correct, na.rm = TRUE) * 100, 2)
pld_n <- sum(!is.na(dta_pld_human$is_correct))
# For per video accuracy use the aggregated measure
human_video_acc <- dta_plot_row1 %>%
  summarise(acc = mean(human_prediction == isfakeorreal.x) * 100) %>%
  pull(acc)
# Model accuracy on same 40 videos
model_video_acc_40 <- mean(dta_preds_au$model_prediction == dta_preds_au$isfakeorreal) * 100
n_40 <- nrow(dta_preds_au)
se_human_40 <- sqrt((human_video_acc/100 * (1 - human_video_acc/100)) / n_40)
se_model_40 <- sqrt((model_video_acc_40/100 * (1 - model_video_acc_40/100)) / n_40)

p_A <- make_accuracy_bar(
  human_video_acc, human_video_acc - 1.96*se_human_40*100, human_video_acc + 1.96*se_human_40*100,
  model_video_acc_40, model_video_acc_40 - 1.96*se_model_40*100, model_video_acc_40 + 1.96*se_model_40*100,
  "Human Observers", "Actual Human vs Model (n=40)"
)

# Panel E: Predicted human vs model (n=93)
p_E <- make_accuracy_bar(
  pred_human_acc_94, ci_low_94, ci_high_94,
  model_acc_94, ci_low_model_94, ci_high_model_94,
  "Predicted Human", "Pred. Human vs Model (n=94)"
)

# Panel I: Predicted human vs model (n=464)
p_I <- make_accuracy_bar(
  pred_human_acc_464, ci_low_464, ci_high_464,
  model_acc_464, ci_low_model_464, ci_high_model_464,
  "Predicted Human", "Pred. Human vs Model (n=464)*"
)

p_A
p_E
p_I
```

## Panel B/F/J: Agreement matrices

```{r}
cat("=== Diagnosing figure vs stats discrepancy ===\n\n")

cat("nrow(dta_results_94):", nrow(dta_results_94), "\n")
cat("nrow(dta_results_464):", nrow(dta_results_464), "\n\n")

# Check direct table from dta_results_94 (what stats chunk uses)
ct_direct_94 <- table(
  Model = factor(as.character(dta_results_94$model_prediction), levels = c("real","fake")),
  Human = factor(as.character(dta_results_94$predicted_human),  levels = c("real","fake"))
)
cat("Direct from dta_results_94:\n")
print(ct_direct_94)

# Check if dta_classifier_test_new has any duplicate video_filenames
cat("\nDuplicate video_filenames in dta_classifier_test_new:",
    sum(duplicated(dta_classifier_test_new$video_filename)), "\n")

# Check what the join in dta_plot_row2 produces
dta_plot_row2_check <- dta_results_94 %>%
  left_join(dta_classifier_test_new, by = "video_filename")
cat("nrow after left_join in dta_plot_row2:", nrow(dta_plot_row2_check), "\n")
cat("nrow after na.omit:", nrow(na.omit(dta_plot_row2_check)), "\n")

# Check isfakeorreal columns after join
cat("\nColumn names after join:\n")
print(names(dta_plot_row2_check))
```

```{r fig4-agreement-matrices, fig.width=4, fig.height=4}
make_agreement_matrix <- function(pred1, pred2, label1, label2, title_text,
                                  fill_color = "#009E73") {
  
  ct <- table(Model = factor(pred1, levels = c("real", "fake")),
              Human = factor(pred2, levels = c("real", "fake")))
  chi_test <- chisq.test(ct)
  phi <- as.numeric(sqrt(chi_test$statistic / sum(ct)))
  N <- sum(ct)
  
  plot_df <- as.data.frame(ct) %>%
    mutate(Percent = Freq / N)
  
  ggplot(plot_df, aes(x = Human, y = Model, fill = Freq)) +
    geom_tile(color = "black", linewidth = 0.6) +
    geom_text(aes(label = sprintf("%d\n(%.0f%%)", Freq, 100 * Percent)),
              size = 5, lineheight = 0.95, color = "black") +
    scale_fill_gradient(low = "white", high = fill_color, limits = c(0, NA)) +
    coord_equal() +
    labs(
      title = title_text,
      subtitle = sprintf("\u03c7\u00b2(1) = %.2f, p = %.3f, \u03c6 = %.2f",
                         chi_test$statistic, chi_test$p.value, phi),
      x = label2, y = label1, fill = "Count"
    ) +
    theme_classic(base_size = 14) +
    theme(
      legend.position = "right",
      plot.title = element_text(face = "bold", hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      axis.title = element_text(face = "bold", color = "black"),
      axis.text = element_text(face = "bold", color = "black")
    )
}

# Panel B: n=40, model vs actual human majority vote
human_pld_mv <- dta_pld_human %>%
  group_by(video_filename) %>%
  summarise(mean_slider = mean(fakeness_slider_response, na.rm = TRUE), .groups = "drop") %>%
  mutate(human_prediction = ifelse(mean_slider >= 50, "fake", "real"))

agree_40 <- dta_preds_au %>%
  left_join(human_pld_mv, by = "video_filename")

p_B <- make_agreement_matrix(
  agree_40$model_prediction, agree_40$human_prediction,
  "Model Prediction", "Human Majority Vote",
  "Model vs Actual Human (n=40)"
)

# Panel F: n=94, model vs LOPO predicted human
p_F <- make_agreement_matrix(
  as.character(dta_results_94$model_prediction),
  as.character(dta_results_94$predicted_human),
  "Model Prediction", "Pred. Human Judgement",
  "Model vs Pred. Human (n=94)"
)

# Panel J: n=464, model vs LOPO predicted human
p_J <- make_agreement_matrix(
  as.character(dta_results_464$model_prediction),
  as.character(dta_results_464$predicted_human),
  "Model Prediction", "Pred. Human Judgement",
  "Model vs Pred. Human (n=464)*"
)

p_B
p_F
p_J
```

## Panels C-D, G-H, K-L: Feature distributions

```{r fig4-feature-plots, fig.width=4, fig.height=4}
make_feature_plot <- function(df, feature_col, color_col, title_text, subtitle_text,
                              y_limits, x_label = "Detection Outcome") {
  
  ggplot(df, aes(x = judgement_type, y = .data[[feature_col]],
                 color = factor(.data[[color_col]], levels = c("real", "fake")))) +
    geom_jitter(width = 0.15, size = 3, alpha = 0.7) +
    stat_summary(fun = mean, geom = "crossbar", width = 0.5,
                 color = "black", linewidth = 0.5) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2,
                 color = "black", linewidth = 0.5) +
    scale_color_manual(values = c("real" = "#009E73", "fake" = "#E69F00"),
                       name = "True Class") +
    scale_y_continuous(limits = y_limits) +
    fig4_theme +
    labs(
      title = title_text,
      subtitle = subtitle_text,
      x = x_label,
      y = "Feature Value"
    )
}

# Row 1: Actual human (n=40)
p_C <- make_feature_plot(dta_plot_row1, "diff1x_pacf5_AU17_chin_raiser", "true_class",
                         "diff1x_pacf5_AU17", "Actual human judgements (n=40)",
                         au17_limits, "Human Detection Outcome")

p_D <- make_feature_plot(dta_plot_row1, "diff1x_pacf5_AU01_inner_brow_raiser", "true_class",
                         "diff1x_pacf5_AU01", "Actual human judgements (n=40)",
                         au01_limits, "Human Detection Outcome")

# Row 2: LOPO predicted (n=94)
p_G <- make_feature_plot(dta_plot_row2, "diff1x_pacf5_AU17_chin_raiser", "true_class",
                         "diff1x_pacf5_AU17", "Predicted human judgements (n=94)",
                         au17_limits, "Predicted Human Detection Outcome")

p_H <- make_feature_plot(dta_plot_row2, "diff1x_pacf5_AU01_inner_brow_raiser", "true_class",
                         "diff1x_pacf5_AU01", "Predicted human judgements (n=94)",
                         au01_limits, "Predicted Human Detection Outcome")

# Row 3: LOPO predicted (n=464)
p_K <- make_feature_plot(dta_plot_row3, "diff1x_pacf5_AU17_chin_raiser", "true_class",
                         "diff1x_pacf5_AU17", "Predicted human judgements (n=464)",
                         au17_limits, "Predicted Human Detection Outcome")

p_L <- make_feature_plot(dta_plot_row3, "diff1x_pacf5_AU01_inner_brow_raiser", "true_class",
                         "diff1x_pacf5_AU01", "Predicted human judgements (n=464)",
                         au01_limits, "Predicted Human Detection Outcome")

p_C
p_D
p_G
p_H
p_K
p_L
```

## Save individual panels to PDF

```{r save-panels, eval=FALSE}
panel_list <- list(
  A = p_A, B = p_B, C = p_C, D = p_D,
  E = p_E, F = p_F, G = p_G, H = p_H,
  I = p_I, J = p_J, K = p_K, L = p_L
)

for (panel_name in names(panel_list)) {
  cairo_pdf(sprintf("fig4_panel_%s.pdf", panel_name), width = 4, height = 4)
  print(panel_list[[panel_name]])
  dev.off()
}

cat("All panels saved.\n")
```

------------------------------------------------------------------------

# Diagnostic Checks

```{r diagnostics}
cat("=== Sanity Checks ===\n\n")

# Check LOPO prediction distribution
cat("LOPO predicted distribution (n=94):\n")
print(table(dta_results_94$predicted_human))

cat("\nLOPO predicted distribution (n=464):\n")
print(table(dta_results_464$predicted_human))

cat("\nActual class distribution (n=94):\n")
print(table(dta_results_94$actual_authenticity))

cat("\nActual class distribution (n=464):\n")
print(table(dta_results_464$actual_authenticity))

# Check for the real-bias issue
cat("\n\nProportion predicting 'real' (n=94):",
    round(mean(dta_results_94$predicted_human == "real") * 100, 1), "%\n")
cat("Proportion predicting 'real' (n=464):",
    round(mean(dta_results_464$predicted_human == "real") * 100, 1), "%\n")

# What were the LOPO training base rates?
cat("\nTraining base rate (human judging 'fake'):",
    round(mean(dta_lopo_training$human_judgement == "fake") * 100, 1), "%\n")
cat("Training base rate (human judging 'real'):",
    round(mean(dta_lopo_training$human_judgement == "real") * 100, 1), "%\n")

# Check that model_au predictions look sensible
cat("\n\nmodel_au prediction distribution (n=94):\n")
print(table(model_preds_94))
cat("\nmodel_au prediction distribution (n=464):\n")
print(table(model_preds_464))
```

# LOSO

```{r}
library(caret)
# Build training data: same as LOPO — human trials joined with RAW AU features
# dta_loso_training is identical to dta_lopo_training in structure
dta_loso_training <- dta_pld_human %>%
  mutate(human_judgement = factor(
    ifelse(fakeness_slider_response >= 50, "fake", "real"),
    levels = c("real", "fake")
  )) %>%
  left_join(dta_cmfts_40_aus, by = "video_filename") %>%
  na.omit()

# Clean up duplicate isfakeorreal columns from join
dta_loso_training <- dta_loso_training %>%
  rename(isfakeorreal = isfakeorreal.x) %>%
  select(-isfakeorreal.y)

cat("LOSO training data:", nrow(dta_loso_training), "trials,",
    length(unique(dta_loso_training$participant_id)), "participants,",
    length(unique(dta_loso_training$video_filename)), "videos\n")

# Same formula as LOPO — 8 important features
loso_formula <- as.formula(paste(
  "human_judgement ~",
  paste(model_au_features, collapse = " + ")
))

# LOSO cross-validation — fold on video_filename
loso_folds <- groupKFold(
  dta_loso_training$video_filename,
  k = length(unique(dta_loso_training$video_filename))
)

loso_control <- trainControl(
  method = "cv",
  index = loso_folds,
  savePredictions = "final",
  classProbs = TRUE
)

# UNCOMMENT TO RETRAIN, ELSE LOAD PRE-TRAINED
# set.seed(12345)
# loso_fit_au_pld <- train(
#   loso_formula,
#   data = dta_loso_training,
#   method = "rf",
#   trControl = loso_control,
#   tuneLength = 1
# )
# saveRDS(loso_fit_au_pld, "loso_fit_au_pld.RDS")

loso_fit_au_pld <- readRDS("loso_fit_au_pld.RDS")

cat("\n=== LOSO Model Results ===\n")
print(loso_fit_au_pld)

cat("\n=== Confusion Matrix ===\n")
conf_matrix_loso <- confusionMatrix(
  loso_fit_au_pld$pred$pred,
  loso_fit_au_pld$pred$obs,
  positive = "fake"
)
print(conf_matrix_loso)
```

```         
Accuracy  Kappa     
  0.582989  0.02466064
```

```         
         Reference
Prediction real fake
      real 1605 1211
      fake  305  399
                                          
               Accuracy : 0.5693          
                 95% CI : (0.5528, 0.5858)
    No Information Rate : 0.5426          
    P-Value [Acc > NIR] : 0.0007663       
                                          
                  Kappa : 0.0922          
                                          
 Mcnemar's Test P-Value : < 2.2e-16       
                                          
            Sensitivity : 0.2478          
            Specificity : 0.8403          
         Pos Pred Value : 0.5668          
         Neg Pred Value : 0.5700          
             Prevalence : 0.4574          
         Detection Rate : 0.1134          
   Detection Prevalence : 0.2000          
      Balanced Accuracy : 0.5441          
                                          
       'Positive' Class : fake
```
