---
title: "final_analysis"
output:
  pdf_document: default
---

# Introduction

This Rmd script will perform a sensitivity analysis to check whether the NMF step is causing a loss of predictive signal. R2 value of approx. 0.4 indicates that NMF is losing over half the variance of the original AU data. This script will check this.

Changes:

1.  Instead of NMF, one representative AU from each of the three components will be chosen based on previous basis matrix loadings (highest loading will be taken).
2.  The CMFTS features will be derived from the AU time series instead.

Highest loaded AUs:

-   Component 1 = AU12

-   Component 2 = AU17

-   Component 3 = AU01

# Load Libraries & Data

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

```{r}
# Load data
# dta_nmf_input <- read.csv("dta_nmf_input.csv")

# Load data before smoothing and normalisation
dta_preprocessed <- read.csv("../../CSV files/OpenFace files/full_dataset_preprocessed.csv")
```

# Smooth Raw AU Data

Now that I am reading in the AU data before smoothing and normalisation, I will do the smoothing step here.

```{r}
dta_preprocessed_smth <- dta_preprocessed %>%
  group_by(video_filename) %>%
  mutate(across(contains("_r"), ~ zoo::rollmean(., k = 4, fill = "extend", align = "left"), .names = "smth_{col}")) %>%
  ungroup()
```

Visualise differences:

```{r, fig.width=5, fig.height=5}
# Visualise difference
dta_preprocessed_smth %>%
  filter(video_filename == "20__exit_phone_room") %>%
  select(frame, AU07_r, smth_AU07_r) %>%
  pivot_longer(cols = c(AU07_r, smth_AU07_r), names_to = "Type", values_to = "Intensity") %>%
  ggplot(aes(x = frame, y = Intensity, color = Type)) +
  geom_line() +
  labs(
    title = "AU07 - Before vs. After Smoothing",
    x = "Frame",
    y = "AU Intensity"
  ) +
  theme_minimal()



dta_preprocessed_smth %>%
  filter(video_filename == "20__exit_phone_room") %>%
  select(frame, AU07_r, smth_AU07_r) %>%
  tidyr::pivot_longer(
    cols = c(AU07_r, smth_AU07_r),
    names_to = "Type",
    values_to = "Intensity"
  ) %>%
  ggplot(aes(x = frame, y = Intensity, color = Type)) +
  
  # --- Different styles for each line ---
  geom_line(
    aes(alpha = Type, linewidth = Type)
  ) +
  
  scale_color_manual(
    values = c(
      "AU07_r"       = "#1b9e77",  # Raw
      "smth_AU07_r"  = "#d95f02"   # Smoothed
    ),
    labels = c("AU07_r" = "Raw", "smth_AU07_r" = "Smoothed")
  ) +
  
  # --- Transparency and line width ---
  scale_alpha_manual(
    values = c("AU07_r" = 0.5, "smth_AU07_r" = 1),
    guide = "none"
  ) +
  scale_linewidth_manual(
    values = c("AU07_r" = 1.1, "smth_AU07_r" = 0.8),
    guide = "none"
  ) +
  
  labs(
    title = "AU07 - Before vs. After Smoothing",
    x = "Frame",
    y = "AU Intensity",
    color = NULL
  ) +
  
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, face = "bold", color = "black"),
    axis.title.x = element_text(size = 18, color = "black"),
    axis.text.y  = element_text(size = 18, color = "black"),
    axis.text.x  = element_text(size = 18, color = "black"),
    plot.title   = element_text(size = 18, face = "bold", color = "black", hjust = 0.5),
    plot.margin  = margin(10, 10, 10, 10)
  )

library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)

dta_long <- dta_preprocessed_smth %>%
  filter(video_filename == "20__exit_phone_room") %>%
  select(frame, AU07_r, smth_AU07_r) %>%
  tidyr::pivot_longer(
    cols = c(AU07_r, smth_AU07_r),
    names_to = "Type",
    values_to = "Intensity"
  )

base_theme <- theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 18, face = "bold", color = "black"),
    axis.title.x = element_text(size = 18, color = "black"),
    axis.text.y  = element_text(size = 16, color = "black"),
    axis.text.x  = element_text(size = 16, color = "black"),
    plot.title   = element_text(size = 18, face = "bold", color = "black", hjust = 0.5),
    plot.margin  = margin(10, 10, 10, 10)
  )

# Raw
p_raw <- dta_long %>%
  filter(Type == "AU07_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#1b9e77", alpha = 0.6, linewidth = 0.8) +
  labs(
    title = "AU07 (Raw)",
    x = NULL,
    y = "AU Intensity"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  base_theme

# Smoothed
p_smth <- dta_long %>%
  filter(Type == "smth_AU07_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#d95f02", linewidth = 0.8) +
  labs(
    title = "AU07 (Smoothed)",
    x = "Frame",
    y = "AU Intensity"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  base_theme

# Patchwork for final figure
p_final <- p_raw / p_smth +
  plot_annotation(
    title = "AU07 - Before vs. After Smoothing",
    theme = theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5, color = "black")
    )
  )

# Save with high dpi
ggsave(
  filename = "AU07_smoothing_comparison.tiff",
  plot = p_final,
  width = 5,
  height = 5,
  dpi = 600,
  compression = "lzw",
  bg = "white"
)

# PDF better for PNAS
ggsave(
  filename = "AU07_smoothing_comparison.pdf",
  plot = p_final,
  width = 5,
  height = 5,
  device = cairo_pdf
)

p_final

```

```{r, fig.width=5, fig.height=2.5}
# Save raw and smooth separately
p_raw
p_smth

ggsave(
  filename = "au07_raw.pdf",
  p_raw,
  width = 5,
  height = 2.5,
  device = cairo_pdf
)

ggsave(
  filename = "au07_smth.pdf",
  p_smth,
  width = 5,
  height = 2.5,
  device = cairo_pdf
)
```

# Split Data 80/20

First need to add the emotion column, easier to do that first:

```{r}
# Function to add emotion column
add_emotion_column <- function(df) {
  df %>%
    mutate(
      emotion = case_when(
        str_detect(video_filename, "_laughing") ~ "yes",
        str_detect(video_filename, "_happy") ~ "yes",
        str_detect(video_filename, "_serious") ~ "yes",
        str_detect(video_filename, "_angry") ~ "yes",
        str_detect(video_filename, "_disgusted") ~ "yes",
        TRUE ~ "no"
      ),
      emotion = factor(emotion, levels = c("yes", "no"))
    )
}

add_emotion_type_column <- function(df) {
  df %>%
    mutate(
      emotion_type = case_when(
        str_detect(video_filename, "_laughing")  ~ "laughing",
        str_detect(video_filename, "_happy")     ~ "happy",
        str_detect(video_filename, "_serious")   ~ "serious",
        str_detect(video_filename, "_angry")     ~ "angry",
        str_detect(video_filename, "_disgusted") ~ "disgusted",
        TRUE                                     ~ "neutral" 
      ),
      
      # Valence column
      valence = case_when(
        emotion_type %in% c("laughing", "happy") ~ "positive",
        emotion_type %in% c("angry", "disgusted") ~ "negative",
        emotion_type %in% c("serious", "neutral") ~ "neutral"
      ),
      
      emotion_type = factor(emotion_type),
      valence = factor(valence, levels = c("positive", "neutral", "negative"))
    )
}


dta_nmf_input <- add_emotion_column(dta_preprocessed_smth)
dta_nmf_input <- add_emotion_type_column(dta_nmf_input)
# Reorder columns
dta_nmf_input <- dta_nmf_input %>%
  select(video_filename, corresponding_video, scene_name, isfakeorreal, emotion, emotion_type, valence, frame, everything())
# Check order is correct
colnames(dta_nmf_input)
```

Now to create the train and test splits:

```{r}
set.seed(12345)

real_videos <- dta_nmf_input %>%
  filter(isfakeorreal == "real") %>%
  distinct(video_filename)

# Sample 80% of real videos for train set
train_real_videos <- sample(real_videos$video_filename,
                            size = floor(0.8 * nrow(real_videos)))

# Get corresponding fake videos for selected real videos
train_fake_videos <- dta_nmf_input %>%
  filter(video_filename %in% train_real_videos) %>%
  pull(corresponding_video) %>%
  unique()

# Train set
dta_nmf_input_train <- dta_nmf_input %>%
  filter(video_filename %in% c(train_real_videos, train_fake_videos))

# Test set is everything else
dta_nmf_input_test <- dta_nmf_input %>%
  filter(!video_filename %in% c(train_real_videos, train_fake_videos))


# Check distribution
print(length(unique(dta_nmf_input_train$video_filename)))
print(length(unique(dta_nmf_input_test$video_filename)))

# Check pairs are preserved
train_videos <- unique(dta_nmf_input_train$video_filename)
test_videos <- unique(dta_nmf_input_test$video_filename)

train_corresponding_videos <- dta_nmf_input %>%
  filter(video_filename %in% train_videos) %>%
  pull(corresponding_video) %>%
  unique()

if (length(intersect(train_corresponding_videos, test_videos)) == 0) {
  cat("Success.\n")
} else {
  cat("Error, the following pairs were split:\n")
  print(intersect(train_corresponding_videos, test_videos))
}
```

Training set: 370 videos

Test set: 94 videos

Additional checks confirmed the pairs are preserved and not split across train and test sets.

# Balance Classes

First we need to see the distribution of classes in both train and test sets.

```{r}
# # Training set
# cat("Training set:\n")
# table(dta_nmf_input_train$isfakeorreal, dta_nmf_input_train$emotion)
# 
# # Test set
# cat("\nTest set:\n")
# table(dta_nmf_input_test$isfakeorreal, dta_nmf_input_test$emotion)


# Get actual number of unique videos in train set.
dta_nmf_input_train %>%
  distinct(video_filename, emotion) %>%
  count(emotion)

# Get actual number of unique videos in test set.
dta_nmf_input_test %>%
  distinct(video_filename, emotion) %>%
  count(emotion)

# Emotion types
dta_nmf_input_train %>%
  distinct(video_filename, emotion_type) %>%
  count(emotion_type)
dta_nmf_input_test %>%
  distinct(video_filename, emotion_type) %>%
  count(emotion_type)

# Valence
dta_nmf_input_train %>%
  distinct(video_filename, valence) %>%
  count(valence)
dta_nmf_input_test %>%
  distinct(video_filename, valence) %>%
  count(valence)
```

In the training set:

-   222 emotion

-   148 no emotion

There is a mild imbalance of emotion vs. no emotion videos (60/40). I could undersample here but it would be throwing away a lot of data. Instead I could oversample using something like SMOTE, although this would have to be done after CMFTS feature extraction as techniques like SMOTE are not really suitable for time series, high dimensional data like this.

UPDATE:

Upsampling cannot be used at this stage due to the complexity of the data, and also cannot be done at the CMFTS feature stage due to not having associated real/fake labels for the new data. Instead, downsampling will be the only available sampling option and will have to occur at this point before NMF. This will have to preserve video pairs too.

# Normalisation

This step will correctly apply normalisation to the videos, in a way that does not lose inter-video variation.

```{r}
au_cols <- names(dta_nmf_input_train)[grepl("^smth_AU", names(dta_nmf_input_train))]

# Learn scaling params from train set only
norm_params <- list()
for (col in au_cols) {
  mean_val <- mean(dta_nmf_input_train[[col]], na.rm = T)
  sd_val <- sd(dta_nmf_input_train[[col]], na.rm = T)
  
  # Handle AUs with zero SD
  if (is.na(sd_val) || sd_val == 0) {
    sd_val <- 1  # To prevent division by zero
  }
  
  norm_params[[col]] <- c(mean = mean_val, sd = sd_val)
}

# Create function to apply z-score and non-negative shift
fn_z_score_shift <- function(x, mean_val, sd_val) {
  z <- (x - mean_val) / sd_val
  # Shift for non-negativity
  z_shifted <- z - min(z, na.rm = T)
  z_shifted[is.na(z_shifted)] <- 0
  return(z_shifted)
}


# Apply params to both train and test sets
dta_nmf_input_train <- dta_nmf_input_train
dta_nmf_input_test <- dta_nmf_input_test

for (col in au_cols) {
  params <- norm_params[[col]]
  
  # Apply to train
  dta_nmf_input_train[[col]] <- fn_z_score_shift(dta_nmf_input_train[[col]],
                                                      params["mean"], params["sd"])
  dta_nmf_input_test[[col]] <- fn_z_score_shift(dta_nmf_input_test[[col]],
                                                      params["mean"], params["sd"])
}
```

Clean up columns:

```{r}
# Remove original AU columns + confidence, success, timestamp
dta_nmf_input_train <- dta_nmf_input_train[, -(9:28)]
dta_nmf_input_test <- dta_nmf_input_test[, -(9:28)]

# Rename AU cols
dta_nmf_input_train <- dta_nmf_input_train %>%
  rename(
    AU01_inner_brow_raiser = smth_AU01_r,
    AU02_outer_brow_raiser = smth_AU02_r,
    AU04_brow_lowerer = smth_AU04_r,
    AU05_upper_lid_raiser = smth_AU05_r,
    AU06_cheek_raiser = smth_AU06_r,
    AU07_lid_tightener = smth_AU07_r,
    AU09_nose_wrinkler = smth_AU09_r,
    AU10_upper_lip_raiser = smth_AU10_r,
    AU12_lip_corner_puller = smth_AU12_r,
    AU14_dimpler = smth_AU14_r,
    AU15_lip_corner_depressor = smth_AU15_r,
    AU17_chin_raiser = smth_AU17_r,
    AU20_lip_stretcher = smth_AU20_r,
    AU23_lip_tightener = smth_AU23_r,
    AU25_lips_part = smth_AU25_r,
    AU26_jaw_drop = smth_AU26_r,
    AU45_blink = smth_AU45_r
  )

dta_nmf_input_test <- dta_nmf_input_test %>%
  rename(
    AU01_inner_brow_raiser = smth_AU01_r,
    AU02_outer_brow_raiser = smth_AU02_r,
    AU04_brow_lowerer = smth_AU04_r,
    AU05_upper_lid_raiser = smth_AU05_r,
    AU06_cheek_raiser = smth_AU06_r,
    AU07_lid_tightener = smth_AU07_r,
    AU09_nose_wrinkler = smth_AU09_r,
    AU10_upper_lip_raiser = smth_AU10_r,
    AU12_lip_corner_puller = smth_AU12_r,
    AU14_dimpler = smth_AU14_r,
    AU15_lip_corner_depressor = smth_AU15_r,
    AU17_chin_raiser = smth_AU17_r,
    AU20_lip_stretcher = smth_AU20_r,
    AU23_lip_tightener = smth_AU23_r,
    AU25_lips_part = smth_AU25_r,
    AU26_jaw_drop = smth_AU26_r,
    AU45_blink = smth_AU45_r
  )
```

# Representative AUs

```{r}
# Will keep the name dta_nmf_output for simplicity

# Only select representative AUs
dta_nmf_input_train <- dta_nmf_input_train %>%
  select(
    "video_filename",
    "isfakeorreal",
    "emotion",
    "valence",
    "frame",
    "AU01_inner_brow_raiser",
    "AU12_lip_corner_puller",
    "AU17_chin_raiser"
  )

dta_nmf_input_test <- dta_nmf_input_test %>%
  select(
    "video_filename",
    "isfakeorreal",
    "emotion",
    "valence",
    "frame",
    "AU01_inner_brow_raiser",
    "AU12_lip_corner_puller",
    "AU17_chin_raiser"
  )

# The equivalent of dta_nmf_output
dta_nmf_output <- dta_nmf_input_train

```

## Visualisations

```{r}
# Function for min-max normalisation, used in visualisations
fn_maxnormalise <- function(x, ...) {
  rng <- range(x, na.rm = TRUE)
  if (anyNA(rng) || diff(rng) == 0) {
    return(rep(0, length(x)))  # safe fallback: all 0s
  } else {
    return((x - rng[1]) / diff(rng))
  }
}
```

```{r}
dta_nmf_output %>%
  pivot_longer(
    cols = starts_with("AU"),
    names_to = "AU",
    values_to = "Activation"
  ) %>%
  group_by(video_filename, AU) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Average across videos for each frame & AU
  group_by(isfakeorreal, AU, frame) %>%
  summarise(Activation = mean(Activation, na.rm=TRUE)) %>%
  # Normalise again by AU
  group_by(AU)%>%
  mutate(Activation = fn_maxnormalise(Activation))%>%
  ungroup() %>%
  # Plot heatmap
  ggplot( aes(x = frame, y = AU, fill = Activation)) +
  geom_tile() +
  facet_wrap(~ isfakeorreal, ncol = 2) +
  scale_fill_gradientn(colours = viridis::magma(100)) +
  labs(
    title = "Heatmap of Averaged Representative AU Activations",
    x = "Frame",
    y = "AU",
    fill = "Activation"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 14)) 
```

```{r}
# Real vs. fake
dta_nmf_output %>%
  pivot_longer(
    cols = starts_with("AU"),
    names_to = "AU",
    values_to = "Activation"
  ) %>%
  # Normalise within each video and AU
  group_by(video_filename, AU) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Average across videos (by AU and frame)
  group_by(isfakeorreal, AU, frame) %>%
  summarise(Activation = mean(Activation, na.rm = TRUE)) %>%
  # Normalise averaged values within each AU
  group_by(AU) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Plot
  ggplot(aes(x = frame, y = Activation, colour = AU)) +
  geom_smooth() +
  facet_wrap(~ isfakeorreal, ncol = 1) +
  labs(
    title = "Representative AU Activations Over Time (Real vs. Fake)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 14)  # Increase font size for facet labels
  )
```

# CMFTS Features

## Install CMFTS Package

```{r}
# Install dependencies
#devtools::install_github("cbergmeir/tsExpKit")

# Install CMFTS package (uncomment if first run)
#devtools::install_github("fjbaldan/CMFTS")
```

## Perform CMFTS Feature Extraction

```{r}
dta_cmfts_input <- dta_nmf_output %>%
  select(video_filename, isfakeorreal, emotion, valence, frame, AU01_inner_brow_raiser, AU12_lip_corner_puller, AU17_chin_raiser) %>%
  pivot_longer(cols = c("AU01_inner_brow_raiser", "AU12_lip_corner_puller", "AU17_chin_raiser"),
               names_to = "AU",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_cmfts_input), value = TRUE)

head(dta_cmfts_input)
```

Function for calculating CMFTS features:

```{r}
fn_extract_cmfts_features <- function(ts_values) {
  ts_matrix <- matrix(as.numeric(ts_values), nrow = 1)
  cmfts_features <- cmfts::cmfts(dataset = ts_matrix)
  return(cmfts_features)
}
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_cmfts_output <- apply(
  dta_cmfts_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_cmfts_input[, c("video_filename", "isfakeorreal", "emotion", "valence", "AU")],
    .
  )

head(dta_cmfts_output)
```

## CMFTS Pre-Processing

First step is to concatenate the data so that each video has one row. This will result in each CMFTS feature having three columns (one for each NMF component). This is to make the data a suitable format for classifiers. More on this in `full_dataset_NMF_analysis.Rmd`.

```{r}
dta_cmfts_output_concat <- dta_cmfts_output %>%
  pivot_wider(
    names_from = AU,
    values_from = -c(video_filename, AU),
    names_glue = "{.value}_{AU}"
  )

# Remove duplicate columns
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -3:-6]
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -5:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_cmfts_output_concat)[2] <- c("isfakeorreal")
colnames(dta_cmfts_output_concat)[3] <- c("emotion")
colnames(dta_cmfts_output_concat)[4] <- c("valence")

head(dta_cmfts_output_concat)
```

### Missing Values

```{r}
dta_cmfts_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

Permutation entropy has all missing values, but some other columns also have a few missing values. I will just remove permutation entropy and impute the rest in the missForest step.

```{r}
# Drop permutation entropy columns
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -14:-16]
```

### Infinite Values

```{r}
dta_cmfts_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

Only Shannon entropy columns have infinite values. This time, there are quite a lot of infinite values, some columns are majority infinite. It is unlikely that missForest imputation will be able to successfully impute based on such little data. Therefore, I will remove Shannon entropy columns and impute other columns with NA values using missForest.

The reason why there are so many inf values is likely to be that some AU values are constant (e.g., if they are not activated in a video).

```{r}
# Remove shannon entropy CS columns
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -14:-16]
```

```{r}
# Convert Infs to NA
for(col in names(dta_cmfts_output_concat)) {
  dta_cmfts_output_concat[[col]][is.infinite(dta_cmfts_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_output_imputed <- dta_cmfts_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_output_imputed[, sapply(dta_cmfts_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE: \< 0.0001

### Features with Zero SD

Features with SD = 0 contribute no meaningful information and can be removed.

```{r}
sds_train <- sapply(dta_cmfts_output_imputed, function(x) {
  if(is.numeric(x)) sd(x, na.rm=T) else NA
})

# List features that do not meet the threshold
features_to_keep <- names(sds_train)[is.na(sds_train) | sds_train > 0]

# Drop features with zero SD
dta_cmfts_output_imputed_zerosd <- dta_cmfts_output_imputed[, features_to_keep]
```

This has removed 9 features.

# Feature Selection

## Correlation

Starting with a quick correlation-based elimination is a good way to find out which columns can be removed to reduce pairwise correlations. Columns that are highly correlated with one another are likely to be redundant for the classifier.

```{r}
# Cutoff 0.9
cat("ZeroSD\n\nCutoff = 0.9\n")
caret::findCorrelation(cor(dta_cmfts_output_imputed_zerosd[, 5:112]), cutoff = 0.9)

# Cutoff 0.95
cat("\nCutoff = 0.95\n")
caret::findCorrelation(cor(dta_cmfts_output_imputed_zerosd[, 5:112]), cutoff = 0.95)
```

A cutoff of 0.9 flagged 14 features for removal due to being highly correlated with at least one other feature. A cutoff point of 0.95 flagged 1 feature for removal.

## Recursive Feature Elimination (RFE)

```{r}
x_train_zerosd <- dta_cmfts_output_imputed_zerosd[, 5:112]
y_train_zerosd <- as.factor(dta_cmfts_output_imputed_zerosd$isfakeorreal)
# y_train_zerosd <- as.factor(dta_cmfts_output_imputed_zerosd$valence)

# Set up RFE control using random forest (fitting params)
control <- caret::rfeControl(functions = caret::rfFuncs,
                      method = "cv",
                      number = 10,
                      verbose = TRUE)  # 10-fold cross validation

# Define subset sizes to evaluate
# Model will start with 5 predictors, incrementing by 5
subsetSizes <- seq(5, ncol(x_train_zerosd), by = 5)

# RFE
set.seed(12345)
rfeResults_zerosd <- caret::rfe(x = x_train_zerosd,
                     y = y_train_zerosd,
                     sizes = subsetSizes,
                     rfeControl = control)

# Results
print(rfeResults_zerosd)
plot(rfeResults_zerosd)
```

RFE identified an optimal subset of 5 features (Accuracy = 53.74, Kappa = 0.075).

Top 5:

1.  Diff1x_pacf5_AU17
2.  Diff2x_pacf5_AU17
3.  Diff2_acf10_AU12
4.  spectral_entropy_AU12
5.  Diff1x_pacf5_AU12

## Boruta

```{r, fig.height=8, fig.width=10}
set.seed(12345)

# Create a subset of data including only the outcome and predictors
boruta_train_zsd <- dta_cmfts_output_imputed_zerosd[, c("isfakeorreal",
                                                        names(dta_cmfts_output_imputed_zerosd)[5:112])]

# boruta_train_zsd <- dta_cmfts_output_imputed_zerosd[, c("valence", 
#                                                         names(dta_cmfts_output_imputed_zerosd)[6:116])]

# Make sure outcome variable is factor
boruta_train_zsd$isfakeorreal <- as.factor(boruta_train_zsd$isfakeorreal)
# boruta_train_zsd$valence <- as.factor(boruta_train_zsd$valence)

# Boruta
boruta_output_zsd <- Boruta::Boruta(isfakeorreal ~ ., data = boruta_train_zsd, doTrace = 2)
# boruta_output_zsd <- Boruta::Boruta(valence ~ ., data = boruta_train_zsd, doTrace = 2)

# Output
# print(boruta_output_zsd)
# plot(boruta_output_zsd, las = 2)

# Solve tentative features
Boruta::TentativeRoughFix(boruta_output_zsd)

# Output
print(boruta_output_zsd)
par(mar = c(12, 4, 4, 2) + 0.1)
plot(Boruta::TentativeRoughFix(boruta_output_zsd), las = 2)

# Solve tentative features
Boruta::TentativeRoughFix(boruta_output_zsd)

# get only important features
print(Boruta::getSelectedAttributes(boruta_output_zsd, withTentative = F))

# ranked
# 1. Get statistics for all attributes, including their median importance
feature_stats <- Boruta::attStats(boruta_output_zsd)

# 2. Add the feature names as a column to the data frame
feature_stats$features <- rownames(feature_stats)

# 3. Filter for only "Confirmed" features and sort by importance (descending)
ranked_features <- feature_stats %>%
  #dplyr::filter(decision == "Confirmed") %>%
  dplyr::arrange(desc(medianImp))

# Print the ranked table
print(ranked_features)

# To get just a vector of the ranked feature names:
ranked_feature_names <- ranked_features$features
print(ranked_feature_names)
```

Boruta finds 3 important features (8 after tentative rough fixing):

1.  diff1x_pacf5_AU17
2.  diff1x_pacf5_AU01
3.  diff2_acf1_AU01
4.  diff2_acf10_AU12
5.  max_kl_shift_AU12
6.  diff1_acf1_AU12
7.  lumpiness_AU12
8.  diff2_acf1_AU12

Boruta graph:

```{r, fig.height=8}
par(mar = c(12, 5, 4, 2) + 0.1)

boruta_fixed <- Boruta::TentativeRoughFix(boruta_output_zsd)
plot(boruta_fixed, 
     las = 3,  # Keep vertical labels
     main = "",
     ylab = "Z-Score Importance",
     xlab = "",
     cex.axis = 0.8,
     cex.main = 2,
     cex.lab = 1.4
)


# Save to pdf
pdf("fig2_panel_A.pdf", width = 12, height = 8)

par(mar = c(18, 5, 4, 2) + 0.1)

boruta_fixed <- Boruta::TentativeRoughFix(boruta_output_zsd)
plot(boruta_fixed, 
     las = 3,  # Keep vertical labels
     main = "",
     ylab = "Z-Score Importance",
     xlab = "",
     cex.axis = 0.8,
     cex.main = 2,
     cex.lab = 1.4
)
dev.off()
```

# Create Test Set

## Create Final Test Set

```{r}
dta_nmf_test_output <- dta_nmf_input_test
```

## Calculate CMFTS Features

The next step is to take the projected NMF values and calculate CMFTS features in the same way as before.

```{r}
dta_cmfts_test_input <- dta_nmf_test_output %>%
  select(video_filename, isfakeorreal, emotion, valence, frame, AU01_inner_brow_raiser, AU12_lip_corner_puller, AU17_chin_raiser) %>%
  pivot_longer(cols = c("AU01_inner_brow_raiser", "AU12_lip_corner_puller", "AU17_chin_raiser"),
               names_to = "AU",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_cmfts_test_input), value = TRUE)

head(dta_cmfts_test_input)
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_cmfts_test_output <- apply(
  dta_cmfts_test_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_cmfts_test_input[, c("video_filename", "isfakeorreal", "emotion", "valence", "AU")],
    .
  )

head(dta_cmfts_test_output)
```

Now time to preprocess in the same way as before.

## CMFTS Preprocessing

```{r}
dta_cmfts_test_output_concat <- dta_cmfts_test_output %>%
  pivot_wider(
    names_from = AU,
    values_from = -c(video_filename, AU),
    names_glue = "{.value}_{AU}"
  )

# Remove duplicate columns
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -3:-6]
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -5:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_cmfts_test_output_concat)[2] <- c("isfakeorreal")
colnames(dta_cmfts_test_output_concat)[3] <- c("emotion")
colnames(dta_cmfts_test_output_concat)[4] <- c("valence")

head(dta_cmfts_test_output_concat)
```

Missing values:

```{r}
dta_cmfts_test_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

```{r}
# Drop permutation entropy columns
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -14:-16]
```

Infinite values:

```{r}
dta_cmfts_test_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

Remove shannon entropy CS columns:

```{r}
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -14:-16]
```

```{r}
# Convert Infs to NA
for(col in names(dta_cmfts_test_output_concat)) {
  dta_cmfts_test_output_concat[[col]][is.infinite(dta_cmfts_test_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_test_output_imputed <- dta_cmfts_test_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_test_output_imputed[, sapply(dta_cmfts_test_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_test_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE = \< 0.00001

Features with Zero SD:

```{r}
# sds_zerosd <- sapply(dta_cmfts_test_output_imputed, function(x) {
#   if(is.numeric(x)) sd(x, na.rm=T) else NA
# })
# 
# # List features that do not meet the threshold
# names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd == 0)]

# Drop features with zero SD
dta_cmfts_test_output_imputed_zerosd <- dta_cmfts_test_output_imputed[, features_to_keep]
```

The test set is ready: `dta_cmfts_test_output_imputed_zerosd`.

# Imbalanced Emotion Class

As we saw earlier, there is an imbalance in the emotion class (222 emotion, 148 no emotion). Now that we have the CMFTS features, we can apply SMOTE upsamping. SMOTE is better suited to the CMFTS data as it is not high dimensional or a time series, unlike the AU data before NMF.

```{r}
# # Make sure emotion column is factor
# dta_cmfts_output_imputed_zerosd$emotion <- as.factor(
#   dta_cmfts_output_imputed_zerosd$emotion
# )
# 
# # Show class imbalance before
# cat("Before SMOTE:\n")
# print(table(dta_cmfts_output_imputed_zerosd$emotion))
# 
# # SMOTE recipe
# smote_recipe <- recipes::recipe(emotion ~ ., data = dta_cmfts_output_imputed_zerosd) %>%
#   recipes::update_role(video_filename, isfakeorreal, new_role = "ID") %>%
#   themis::step_smote(emotion, over_ratio = 1) %>%  # over_ratio = 1 upsampling minority
#   recipes::prep()
# 
# dta_cmfts_emotion_balanced <- recipes::bake(smote_recipe, new_data = NULL)
# 
# # Class balance after
# cat("\nAfter SMOTE:\n")
# print(table(dta_cmfts_emotion_balanced$emotion))
```

The final balanced training set is `dta_cmfts_emotion_balanced`.

UPDATE:

This cannot be used for model training as the new rows do not have associated real/fake labels. Instead, we can oversample using bootstrapping.

```{r}
set.seed(12345)

# Separate the minority and majority
majority_data <- dta_cmfts_output_imputed_zerosd %>%
  filter(emotion == "yes")

minority_data <- dta_cmfts_output_imputed_zerosd %>%
  filter(emotion == "no")

# Bootstrapping on minority class
minority_data_oversampled <- minority_data %>%
  sample_n(size = nrow(majority_data), replace = TRUE)

# Combine to create balanced dta
dta_cmfts_emotion_balanced <- bind_rows(majority_data, minority_data_oversampled)

# Verify
cat("Before:\n")
print(table(dta_cmfts_output_imputed_zerosd$emotion))

cat("\nAfter:\n")
print(table(dta_cmfts_emotion_balanced$emotion))
```

Balancing valence:

-   Balancing the emotion_type class will be too extreme since the minority class has 6 and the majority class has 148.

```{r}
set.seed(12345)

target_size <- nrow(dta_cmfts_output_imputed_zerosd %>% filter(valence == "neutral"))

# Separate data into the three classes
neutral_data <- dta_cmfts_output_imputed_zerosd %>% filter(valence == "neutral")
positive_data <- dta_cmfts_output_imputed_zerosd %>% filter(valence == "positive")
negative_data <- dta_cmfts_output_imputed_zerosd %>% filter(valence == "negative")

# Oversample minority classes to match target size
positive_data <- positive_data %>%
  sample_n(size = target_size, replace = T)
negative_data <- negative_data %>%
  sample_n(size = target_size, replace = T)

# Combine into new df
dta_cmfts_valenced_balanced <- bind_rows(neutral_data, positive_data, negative_data)

# Check
cat("Before:\n")
print(table(dta_cmfts_output_imputed_zerosd$valence))
cat("After:\n")
print(table(dta_cmfts_valenced_balanced$valence))

```

# Classifiers

I will create two models for each classifier.

-   Model 1 = Only features confirmed as important by Boruta (n=2)

-   Model 2 = Same feature types but across all components (n=6)

The reason for this stems from the idea that the three NMF components represent interrelated aspects of facial expression, so if a feature type was most discriminative for one component, its behaviour in other components might still hold relevant information for classification.

```{r}
set.seed(12345)

# Train set (UNCOMMENT FOR IMBALANCED EMOTION)
# dta_cmfts_output_imputed_zerosd$isfakeorreal <- factor(
#   dta_cmfts_output_imputed_zerosd$isfakeorreal,
#   levels = c("real", "fake")
# )

dta_cmfts_emotion_balanced$isfakeorreal <- factor(
  dta_cmfts_emotion_balanced$isfakeorreal,
  levels = c("real", "fake")
)

# dta_cmfts_valenced_balanced$valence <- factor(
#   dta_cmfts_valenced_balanced$valence,
#   levels = c("positive", "neutral", "negative")
# )



# UNCOMMENT FOR REAL/FAKE PREDICTION
boruta_selected_features <- c(
  "diff1x_pacf5_AU17_chin_raiser",
  "diff1x_pacf5_AU01_inner_brow_raiser",
  "diff2_acf1_AU01_inner_brow_raiser",
  "diff2_acf10_AU12_lip_corner_puller",
  "max_kl_shift_AU12_lip_corner_puller",
  "diff1_acf1_AU12_lip_corner_puller",
  "lumpiness_AU12_lip_corner_puller",
  "diff2_acf1_AU12_lip_corner_puller"
)

# # Valence
# boruta_selected_features <- c(
#   "shannon_entropy_CS_nmf1",
#   "shannon_entropy_CS_nmf3",
#   "shannon_entropy_SG_nmf2",
#   "shannon_entropy_SG_nmf3",
#   "linearity_nmf3"
# )

# Select these features plus the outcome variable
dta_classifier_train_new <- dta_cmfts_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

dta_classifier_train_new_balanced <- dta_cmfts_emotion_balanced %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

# UNCOMMENT FOR REAL FAKE
# dta_classifier_train_new <- dta_cmfts_emotion_balanced %>%
#   select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

# Valence
# dta_classifier_train_new <- dta_cmfts_valenced_balanced %>%
#   select(video_filename, isfakeorreal, emotion_type, valence, all_of(boruta_selected_features))



# Test set
dta_cmfts_test_output_imputed_zerosd$isfakeorreal <- factor(
  dta_cmfts_test_output_imputed_zerosd$isfakeorreal,
  levels = c("real", "fake")
)

dta_classifier_test_new <- dta_cmfts_test_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

dta_classifier_test_new$video_filename <- dta_cmfts_test_output_imputed_zerosd$video_filename

# Valence
# dta_cmfts_test_output_imputed_zerosd$valence <- factor(
#   dta_cmfts_test_output_imputed_zerosd$valence,
#   levels = c("positive", "neutral", "negative")
# )
# 
# dta_classifier_test_new <- dta_cmfts_test_output_imputed_zerosd %>%
#   select(video_filename, isfakeorreal, emotion_type, valence, all_of(boruta_selected_features))
# 
# dta_classifier_test_new_all <- dta_cmfts_test_output_imputed_zerosd %>%
#   select(video_filename, isfakeorreal, emotion_type, valence, all_of(all_component_features))
# 
# dta_classifier_test_new$video_filename <- dta_cmfts_test_output_imputed_zerosd$video_filename
```

## Normalise Features

Non-decision tree classifiers are sensitive to the scale of the features. This section will scale the CMFTS features in the same way as the AU data previously.

```{r}
# Columns to be scaled
feature_columns <- names(dta_classifier_train_new)[!names(dta_classifier_train_new) %in% c("video_filename", "emotion", "isfakeorreal")]

# Scaling params from training set
scaling_model <- caret::preProcess(dta_classifier_train_new[, feature_columns],
                                   method = c("center", "scale"))  # z-score norm

# Save scaling model for human comparison analyses
saveRDS(scaling_model, "models/scaling_model_AUs.RDS")

model_au_features <- c(
  "diff1x_pacf5_AU17_chin_raiser",
  "diff1x_pacf5_AU01_inner_brow_raiser",
  "diff2_acf1_AU01_inner_brow_raiser",
  "diff2_acf10_AU12_lip_corner_puller",
  "max_kl_shift_AU12_lip_corner_puller",
  "diff1_acf1_AU12_lip_corner_puller",
  "lumpiness_AU12_lip_corner_puller",
  "diff2_acf1_AU12_lip_corner_puller"
)

cat("Verification of regenerated scaling_model:\n")
test <- predict(scaling_model, newdata = dta_classifier_train_new[, feature_columns])
cat("Means (should be ~0):\n")
print(round(colMeans(test[, model_au_features]), 4))
cat("SDs (should be ~1):\n")
print(round(apply(test[, model_au_features], 2, sd), 4))

# Create new scaled data by applying learned params
# Train set
dta_classifier_train_new_scaled <- dta_classifier_train_new
dta_classifier_train_new_scaled[, feature_columns] <- predict(scaling_model,
                                                              newdata = dta_classifier_train_new[, feature_columns])

# For balanced train set
dta_classifier_train_new_balanced_scaled <- dta_classifier_train_new_balanced
dta_classifier_train_new_balanced_scaled[, feature_columns] <- predict(scaling_model,
                                                              newdata = dta_classifier_train_new_balanced[, feature_columns])

# Test set
dta_classifier_test_new_scaled <- dta_classifier_test_new
dta_classifier_test_new_scaled[, feature_columns] <- predict(scaling_model,
                                                             newdata = dta_classifier_test_new[, feature_columns])




dta_classifier_train_new_scaled$isfakeorreal <- factor(
  dta_classifier_train_new_scaled$isfakeorreal,
  levels = c("fake", "real")
)
dta_classifier_train_new_balanced_scaled$isfakeorreal <- factor(
  dta_classifier_train_new_balanced_scaled$isfakeorreal,
  levels = c("fake", "real")
)
dta_classifier_test_new_scaled$isfakeorreal <- factor(
  dta_classifier_test_new_scaled$isfakeorreal,
  levels = c("fake", "real")
)
```

## Random Forest

```{r}
classifier_control <- caret::trainControl(
  method = "none",
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

# Valence
# library(MLmetrics)
# # Cross validation setup
# classifier_control <- caret::trainControl(
#   method = "none",
#   classProbs = TRUE,
#   summaryFunction = caret::multiClassSummary
# )

set.seed(12345)

rf_fit_model1 <- caret::train(
  isfakeorreal ~ ., # nmf
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "rf",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(mtry = 4)
)

# Valence
# rf_fit_model1 <- caret::train(
#   valence ~ ., # nmf
#   data = dta_classifier_train_new[, -(1:3)],
#   method = "rf",
#   metric = "ROC",
#   trControl = classifier_control,
#   tuneGrid = expand.grid(mtry = 4)
# )



print(rf_fit_model1)
plot(caret::varImp(rf_fit_model1)) # To see feature importance
```

Evaluate on unseen data:

```{r}
rf_model_1_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model1, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

# Valence
# rf_model_1_performance <- caret::confusionMatrix(
#   data = predict(rf_fit_model1, newdata = dta_classifier_test_new),
#   reference = dta_classifier_test_new$valence
# )

print(rf_model_1_performance)


# SAVE RDS
saveRDS(rf_fit_model1, "models/rf_fit_model1_AUs.RDS")
```

-   Accuracy = 68.09%

-   95% CI = (0.5767, 0.7733)

-   NIR = 0.5

-   P = 0.000294 (P \< 0.001)

-   Kappa = 0.3617

-   Sensitivity = 80.85%

-   Specificity = 55.32%

-   Precision = 64.41%

-   F1 = 71.70%

```         
         Reference
Prediction real fake
      real   26    9
      fake   21   38
```

## C5.0B

```{r}
set.seed(12345)

c50_fit_model1 <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "C5.0",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(
    model = "tree",
    trials = 20,  # number of boosting iterations
    winnow = FALSE
  )
)

# Valence
# c50_fit_model1 <- caret::train(
#   valence ~ .,
#   data = dta_classifier_train_new[, -(1:3)],
#   method = "C5.0",
#   metric = "ROC",
#   trControl = classifier_control,
#   tuneGrid = expand.grid(
#     model = "tree",
#     trials = 20,  # number of boosting iterations
#     winnow = FALSE
#   )
# )

print(c50_fit_model1)
plot(caret::varImp(c50_fit_model1)) # To see feature importanc
```

Evaluate on unseen data:

```{r}
c50_model_1_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model1, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

# Valence
# c50_model_1_performance <- caret::confusionMatrix(
#   data = predict(c50_fit_model1, newdata = dta_classifier_test_new),
#   reference = dta_classifier_test_new$valence
# )

print(c50_model_1_performance)
```

```         
   Reference
Prediction real fake
      real   24   10
      fake   23   37
                                          
               Accuracy : 0.6489          
                 95% CI : (0.5436, 0.7446)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002539        
                                          
                  Kappa : 0.2979          
                                          
 Mcnemar's Test P-Value : 0.036714        
                                          
            Sensitivity : 0.7872          
            Specificity : 0.5106          
         Pos Pred Value : 0.6167          
         Neg Pred Value : 0.7059          
             Prevalence : 0.5000          
         Detection Rate : 0.3936          
   Detection Prevalence : 0.6383          
      Balanced Accuracy : 0.6489  
      F1: 69.16%
```

# Repeated Cross-Validation

The next step is to train the classifiers again but this time using repeated CV. We did not do this before due to the NMF data leakage problem. However, we will do this now to compare results.

## Random Forest

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

rf_fit_model1_cv <- caret::train(
  isfakeorreal ~ ., # nmf
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "rf",
  metric = "ROC",
  trControl = classifier_control,
  tuneLength = 5
)

print(rf_fit_model1_cv)
plot(caret::varImp(rf_fit_model1_cv)) # To see feature importance
```

Evaluate on unseen data:

```{r}
rf_model_1_cv_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

5 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   29   12
      fake   18   35
                                          
               Accuracy : 0.6809          
                 95% CI : (0.5767, 0.7733)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.000294        
                                          
                  Kappa : 0.3617          
                                          
 Mcnemar's Test P-Value : 0.361310        
                                          
            Sensitivity : 0.7447          
            Specificity : 0.6170          
         Pos Pred Value : 0.6604          
         Neg Pred Value : 0.7073          
             Prevalence : 0.5000          
         Detection Rate : 0.3723          
   Detection Prevalence : 0.5638          
      Balanced Accuracy : 0.6809 
```

10 folds, 3 repeats:

```         
         Reference
Prediction real fake
      real   26   12
      fake   21   35
                                          
               Accuracy : 0.6489          
                 95% CI : (0.5436, 0.7446)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002539        
                                          
                  Kappa : 0.2979          
                                          
 Mcnemar's Test P-Value : 0.163734        
                                          
            Sensitivity : 0.7447          
            Specificity : 0.5532          
         Pos Pred Value : 0.6250          
         Neg Pred Value : 0.6842          
             Prevalence : 0.5000          
         Detection Rate : 0.3723          
   Detection Prevalence : 0.5957          
      Balanced Accuracy : 0.6489 
```

5 folds, 3 repeats (scaled features) FINAL:

```         
          Reference
Prediction fake real
      fake   34   19
      real   13   28
                                          
               Accuracy : 0.6596          
                 95% CI : (0.5546, 0.7542)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.001294        
                                          
                  Kappa : 0.3191          
                                          
 Mcnemar's Test P-Value : 0.376759        
                                          
            Sensitivity : 0.7234          
            Specificity : 0.5957          
         Pos Pred Value : 0.6415          
         Neg Pred Value : 0.6829          
             Prevalence : 0.5000          
         Detection Rate : 0.3617          
   Detection Prevalence : 0.5638          
      Balanced Accuracy : 0.6596          
                                          
       'Positive' Class : fake   
```

## C5.0B

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

c50_fit_model1_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "C5.0",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(
    model = "tree",
    trials = c(1, 10, 20),  # number of boosting iterations
    winnow = FALSE
  )
)

print(c50_fit_model1_cv)
plot(caret::varImp(c50_fit_model1_cv)) # To see feature importanc
```

Evaluate on unseen data:

```{r}
c50_model_1_cv_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(c50_model_1_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

5 folds, 3 repeats scaled FINAL:

```         
          Reference
Prediction real fake
      real   24   10
      fake   23   37
                                          
               Accuracy : 0.6489          
                 95% CI : (0.5436, 0.7446)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002539        
                                          
                  Kappa : 0.2979          
                                          
 Mcnemar's Test P-Value : 0.036714        
                                          
            Sensitivity : 0.7872          
            Specificity : 0.5106          
         Pos Pred Value : 0.6167          
         Neg Pred Value : 0.7059          
             Prevalence : 0.5000          
         Detection Rate : 0.3936          
   Detection Prevalence : 0.6383          
      Balanced Accuracy : 0.6489   
```

10 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   24   10
      fake   23   37
                                          
               Accuracy : 0.6489          
                 95% CI : (0.5436, 0.7446)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002539        
                                          
                  Kappa : 0.2979          
                                          
 Mcnemar's Test P-Value : 0.036714        
                                          
            Sensitivity : 0.7872          
            Specificity : 0.5106          
         Pos Pred Value : 0.6167          
         Neg Pred Value : 0.7059          
             Prevalence : 0.5000          
         Detection Rate : 0.3936          
   Detection Prevalence : 0.6383          
      Balanced Accuracy : 0.6489   
```

# 40 Held-Out Videos Evaluation

I will test all models on the 40 held-out videos from the human participant experiment as another test set.

But first, I will have to derive the CMFTS features for these videos.

## Load and Preprocess AU Data

```{r}
dta_video_au_40 <- read.csv("../human_data_preprocess/video_aus.csv")

# Make sure max frames is 241
dta_video_au_40 <- dta_video_au_40 %>%
  filter(frame < 242)

# Keep relevant columns
dta_video_au_40 <- dta_video_au_40 %>%
  select(
    video_source,
    frame,
    emotion_yes_no_new,
    isfakeorreal,
    contains("_r_")
  )

# Smooth AU values
dta_video_au_40_smth <- dta_video_au_40 %>%
  group_by(video_source) %>%
  mutate(across(contains("_r_"), ~ zoo::rollmean(., k = 4, fill = "extend", align = "left"), .names = "smth_{col}")) %>%
  ungroup()

# Visualise difference
dta_video_au_40_smth %>%
  filter(video_source == "27__kitchen_pan") %>%
  select(frame, au07_r_lid_tightener, smth_au07_r_lid_tightener) %>%
  pivot_longer(cols = c(au07_r_lid_tightener, smth_au07_r_lid_tightener), names_to = "Type", values_to = "Intensity") %>%
  ggplot(aes(x = frame, y = Intensity, color = Type)) +
  geom_line() +
  labs(
    title = "AU07 - Before vs. After Smoothing",
    x = "Frame",
    y = "AU Intensity"
  ) +
  theme_minimal()
```

Clean up column names:

```{r}
dta_video_au_40_input <- dta_video_au_40_smth %>%
  select(
    video_source,
    frame,
    emotion_yes_no_new,
    isfakeorreal,
    starts_with("smth_")
  ) %>%
  rename(
    smth_AU01_r = smth_au01_r_inner_brow_raiser,
    smth_AU02_r = smth_au02_r_outer_brow_raiser,
    smth_AU04_r = smth_au04_r_brow_lowerer,
    smth_AU05_r = smth_au05_r_upper_lid_raiser,
    smth_AU06_r = smth_au06_r_cheek_raiser,
    smth_AU07_r = smth_au07_r_lid_tightener,
    smth_AU09_r = smth_au09_r_nose_wrinkler,
    smth_AU10_r = smth_au10_r_upper_lip_raiser,
    smth_AU12_r = smth_au12_r_lip_corner_puller,
    smth_AU14_r = smth_au14_r_dimpler,
    smth_AU15_r = smth_au15_r_lip_corner_depressor,
    smth_AU17_r = smth_au17_r_chin_raiser,
    smth_AU20_r = smth_au20_r_lip_stretcher,
    smth_AU23_r = smth_au23_r_lip_tightener,
    smth_AU25_r = smth_au25_r_lips_part,
    smth_AU26_r = smth_au26_r_jaw_drop,
    smth_AU45_r = smth_au45_r_blink
  )

# I have renamed them to match the same format as the original normalisation code. 
```

## Normalisation

Here I will normalise the data in the same way as before - by using the scaling parameters learnt from the training data.

```{r}
for (col in au_cols) {
  params <- norm_params[[col]]
  
  dta_video_au_40_input[[col]] <- fn_z_score_shift(
    dta_video_au_40_input[[col]],
    params["mean"],
    params["sd"]
  )
}
```

## Select Representative AUs

Now that the data has been normalised properly, I will select the representative AU columns and remove the other AU columns.

```{r}
dta_40_test_output <- dta_video_au_40_input %>%
  select(
    video_source,
    isfakeorreal,
    emotion_yes_no_new,
    frame,
    smth_AU01_r,
    smth_AU12_r,
    smth_AU17_r
  ) %>%
  rename(
    video_filename = video_source,
    emotion = emotion_yes_no_new,
    AU01_inner_brow_raiser = smth_AU01_r,
    AU12_lip_corner_puller = smth_AU12_r,
    AU17_chin_raiser = smth_AU17_r
  )

```

## Derive CMFTS Features

```{r}
dta_40_cmfts_input <- dta_40_test_output %>%
  select(video_filename, isfakeorreal, emotion, frame, AU01_inner_brow_raiser, AU12_lip_corner_puller, AU17_chin_raiser) %>%
  pivot_longer(cols = c("AU01_inner_brow_raiser", "AU12_lip_corner_puller", "AU17_chin_raiser"),
               names_to = "AU",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_40_cmfts_input), value = TRUE)

head(dta_40_cmfts_input)
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_40_cmfts_output <- apply(
  dta_40_cmfts_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_40_cmfts_input[, c("video_filename", "isfakeorreal", "emotion", "AU")],
    .
  )

head(dta_40_cmfts_output)
```

## CMFTS Preprocessing

```{r}
dta_40_cmfts_output_concat <- dta_40_cmfts_output %>%
  pivot_wider(
    names_from = AU,
    values_from = -c(video_filename, AU),
    names_glue = "{.value}_{AU}"
  )

# Remove duplicate columns
dta_40_cmfts_output_concat <- dta_40_cmfts_output_concat[, -3:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_40_cmfts_output_concat)[2] <- c("isfakeorreal")
colnames(dta_40_cmfts_output_concat)[3] <- c("emotion")

head(dta_40_cmfts_output_concat)
```

Missing values:

```{r}
dta_40_cmfts_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

```{r}
# Drop permutation entropy columns
dta_40_cmfts_output_concat <- dta_40_cmfts_output_concat[, -13:-15]
```

Infinite values:

```{r}
dta_40_cmfts_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

Drop Shannon entropy CS columns as most values are infinite.

```{r}
dta_40_cmfts_output_concat <- dta_40_cmfts_output_concat[, -13:-15]
```

```{r}
# Convert Infs to NA
for(col in names(dta_40_cmfts_output_concat)) {
  dta_40_cmfts_output_concat[[col]][is.infinite(dta_40_cmfts_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_40_test_output_imputed <- dta_40_cmfts_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_40_test_output_imputed[, sapply(dta_cmfts_40_test_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_40_test_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE = 0.011

Features with zero SD:

```{r}
sds_zerosd <- sapply(dta_cmfts_40_test_output_imputed, function(x) {
  if(is.numeric(x)) sd(x, na.rm=T) else NA
})

# List features that do not meet the threshold
names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd == 0)]

# Drop features with zero SD
dta_cmfts_40_test_output_imputed_zerosd <- dta_cmfts_40_test_output_imputed[, names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd > 0)]]
```

The final dataset for this is ready: `dta_cmfts_40_test_output_imputed_zerosd`.

## Model Predictions

First to normalise the CMFTS features so they are on the same scale as the model was trained on.

```{r}
# Rename for simplicity
dta_classifier_40_test <- dta_cmfts_40_test_output_imputed_zerosd

# Test set
dta_classifier_40_test$isfakeorreal <- factor(
  dta_classifier_40_test$isfakeorreal,
  levels = c("fake", "real")
)

# Normalise
feature_columns <- names(dta_classifier_train_new)[!names(dta_classifier_train_new) %in% c("video_filename", "emotion", "isfakeorreal")]

dta_classifier_40_test_scaled <- dta_classifier_40_test
# Apply the scaling model that was learned from the training data
dta_classifier_40_test_scaled[, feature_columns] <- predict(scaling_model,
                                                             newdata = dta_classifier_40_test[, feature_columns])


head(dta_classifier_40_test_scaled)
```

### Random Forest

```{r}
rf_model_1_40 <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled),
  reference = dta_classifier_40_test_scaled$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_40)

# ROC AUC
y_test  <- factor(dta_classifier_40_test_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

ROC-AUC = 0.5475 [0.3624-0.7326]

Accuracy = 55% [39.8-69.3]

### C5.0B

```{r}
c50_model_1_40 <- caret::confusionMatrix(
  data = predict(c50_fit_model1_cv, newdata = dta_classifier_40_test_scaled),
  reference = dta_classifier_40_test_scaled$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_40)

# ROC AUC
y_test  <- factor(dta_classifier_40_test_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(c50_fit_model1_cv, newdata = dta_classifier_40_test_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

# Emotion Vs No Emotion

## 20% Test Set:

### Random Forest

Emotion:

```{r}
dta_classifier_test_new_emotion <- dta_classifier_test_new_scaled %>%
  filter(emotion == "yes")

rf_model_1_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(rf_model_1_emotion)

# ROC AUC
y_test  <- factor(dta_classifier_test_new_emotion$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_emotion, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

Without CV:

```         
          Reference
Prediction real fake
      real   21    7
      fake    8   22
                                          
               Accuracy : 0.7414          
                 95% CI : (0.6096, 0.8474)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.0001535       
                                          
                  Kappa : 0.4828          
                                          
 Mcnemar's Test P-Value : 1.0000000       
                                          
            Sensitivity : 0.7586          
            Specificity : 0.7241          
         Pos Pred Value : 0.7333          
         Neg Pred Value : 0.7500          
             Prevalence : 0.5000          
         Detection Rate : 0.3793          
   Detection Prevalence : 0.5172          
      Balanced Accuracy : 0.7414 
```

With CV:

```         
          Reference
Prediction real fake
      real   22   10
      fake    7   19
                                          
               Accuracy : 0.7069          
                 95% CI : (0.5727, 0.8191)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.001116        
                                          
                  Kappa : 0.4138          
                                          
 Mcnemar's Test P-Value : 0.627626        
                                          
            Sensitivity : 0.6552          
            Specificity : 0.7586          
         Pos Pred Value : 0.7308          
         Neg Pred Value : 0.6875          
             Prevalence : 0.5000          
         Detection Rate : 0.3276          
   Detection Prevalence : 0.4483          
      Balanced Accuracy : 0.7069 
```

With CV FINAL:

```         
          Reference
Prediction fake real
      fake   19    6
      real   10   23
                                         
               Accuracy : 0.7241         
                 95% CI : (0.591, 0.8334)
    No Information Rate : 0.5            
    P-Value [Acc > NIR] : 0.0004309      
                                         
                  Kappa : 0.4483         
                                         
 Mcnemar's Test P-Value : 0.4532547      
                                         
            Sensitivity : 0.6552         
            Specificity : 0.7931         
         Pos Pred Value : 0.7600         
         Neg Pred Value : 0.6970         
             Prevalence : 0.5000         
         Detection Rate : 0.3276         
   Detection Prevalence : 0.4310         
      Balanced Accuracy : 0.7241         
                                         
       'Positive' Class : fake           
                                         
Setting direction: controls < cases
Area under the curve: 0.7652
95% CI: 0.6415-0.8888 (DeLong)
```

Bar chart:

```{r fig.width=6, fig.height=6}
ggplot(
  data = data.frame(
    Condition = factor(c("Emotion Videos", "No Emotion Videos"),
                       levels = c("Emotion Videos", "No Emotion Videos")),
    Accuracy  = c(72.4, 55.6),
    CI_low    = c(59.8, 39.6),
    CI_high   = c(82.2, 70.5),
    p_label   = c("p < .001", "p = .309")
  ),
  aes(x = Condition, y = Accuracy)
) +
  geom_col(width = 0.6, fill = "gray50") +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high),
                width = 0.2, size = 0.8) +
  geom_hline(yintercept = 50, linetype = "dashed",
             colour = "red", size = 0.6, alpha = 0.6) +
  annotate("text", x = 2, y = 50, label = "NIR (0.5)",
           hjust = -1.11, vjust = -0.6, size = 4, colour = "red") +
  geom_text(aes(label = p_label), vjust = -4, size = 8) +
  scale_y_continuous(limits = c(0, 100),
                     expand = expansion(mult = c(0, 0.08))) +
  labs(y = "Accuracy (%)", x = NULL) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x  = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    axis.title.y = element_text(size = 18, face = "bold"),
    plot.margin  = margin(10, 10, 10, 10)
  )

```

No emotion:

```{r}
set.seed(12345)
dta_classifier_test_new_no_emotion <- dta_classifier_test_new_scaled %>%
  filter(emotion == "no")

rf_model_1_no_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(rf_model_1_no_emotion)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_no_emotion$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_no_emotion, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

Without CV:

```         
          Reference
Prediction real fake
      real    6    2
      fake   12   16
                                          
               Accuracy : 0.6111          
                 95% CI : (0.4346, 0.7686)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.12149         
                                          
                  Kappa : 0.2222          
                                          
 Mcnemar's Test P-Value : 0.01616         
                                          
            Sensitivity : 0.8889          
            Specificity : 0.3333          
         Pos Pred Value : 0.5714          
         Neg Pred Value : 0.7500          
             Prevalence : 0.5000          
         Detection Rate : 0.4444          
   Detection Prevalence : 0.7778          
      Balanced Accuracy : 0.6111          
                                          
       'Positive' Class : fake  
```

With CV:

```         
          Reference
Prediction real fake
      real    7    2
      fake   11   16
                                          
               Accuracy : 0.6389          
                 95% CI : (0.4622, 0.7918)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.06625         
                                          
                  Kappa : 0.2778          
                                          
 Mcnemar's Test P-Value : 0.02650         
                                          
            Sensitivity : 0.8889          
            Specificity : 0.3889          
         Pos Pred Value : 0.5926          
         Neg Pred Value : 0.7778          
             Prevalence : 0.5000          
         Detection Rate : 0.4444          
   Detection Prevalence : 0.7500          
      Balanced Accuracy : 0.6389  
```

With CV FINAL:

```         
          Reference
Prediction fake real
      fake   15   13
      real    3    5
                                         
               Accuracy : 0.5556         
                 95% CI : (0.381, 0.7206)
    No Information Rate : 0.5            
    P-Value [Acc > NIR] : 0.30886        
                                         
                  Kappa : 0.1111         
                                         
 Mcnemar's Test P-Value : 0.02445        
                                         
            Sensitivity : 0.8333         
            Specificity : 0.2778         
         Pos Pred Value : 0.5357         
         Neg Pred Value : 0.6250         
             Prevalence : 0.5000         
         Detection Rate : 0.4167         
   Detection Prevalence : 0.7778         
      Balanced Accuracy : 0.5556         
                                         
       'Positive' Class : fake           
                                         
Setting direction: controls < cases
Area under the curve: 0.6049
95% CI: 0.4071-0.8028 (DeLong)
```

Is the reported accuracy for emotion vs no emotion significantly different?

```{r}
# First, test accuracy

# Two-sample test of proportions + Fisher's exact
test <- dta_classifier_test_new_scaled %>% 
  mutate(
    pred_class = predict(rf_fit_model1_cv, newdata = ., type = "raw"),
    pred_fake = predict(rf_fit_model1_cv, newdata = ., type = "prob")[,"fake"],
    correct = as.integer(pred_class == isfakeorreal)
)

emotion_test <- filter(test, emotion == "yes")
no_emotion_test <- filter(test, emotion == "no")

# Accuracy comparison
x <- c(sum(emotion_test$correct), sum(no_emotion_test$correct))  # number correct in each group
n <- c(nrow(emotion_test), nrow(no_emotion_test))  # group sizes

prop.test(x, n, correct = TRUE)  # chi-square with yates
fisher.test(matrix(c(x[1], n[1]-x[1], x[2], n[2]-x[2]), nrow = 2, byrow = T))  # exact

```

The observed gap between emotion accuracy (72.4%) and no emotion accuracy (55.6%) of 16.9 percentage points is not statistically significant at α = 0.05.

Yates p = 0.146, Fisher p = 0.119.

Bootstrap CI for accuracy difference:

```{r}
set.seed(12345)
B <- 1000
boot_diff <- replicate(B, {
  em_i <- emotion_test[sample.int(nrow(emotion_test), replace = T), ]
  no_em_i <- no_emotion_test[sample.int(nrow(no_emotion_test), replace = T), ]
  mean(em_i$correct) - mean(no_em_i$correct)
})
quantile(boot_diff, c(.025, .5, .975))
```

The CI crosses zero, so cannot rule out no true difference with this sample.

Is reported ROC-AUC significantly different for emotion vs. no emotion?

```{r}
roc_emotion <- pROC::roc(emotion_test$isfakeorreal, emotion_test$pred_fake,
                         levels = c("real", "fake"), direction = "<")
roc_no_emotion <- pROC::roc(no_emotion_test$isfakeorreal, no_emotion_test$pred_fake,
                            levels = c("real", "fake"), direction = "<")

# AUC & 95% CI for both
pROC::auc(roc_emotion); pROC::ci.auc(roc_emotion)
pROC::auc(roc_no_emotion); pROC::ci.auc(roc_no_emotion)

# Test AUCs between groups
pROC::roc.test(roc_emotion, roc_no_emotion, method = "delong", paired = F)
```

Emotion AUC = 0.765 (95% CI = 0.642-0.889)

No emotion AUC = 0.605 (95% CI = 0.407-0.803)

DeLong test for the gap: p = .183

A suggestive but non-significant improvement in Emotion videos, with wide CIs due to low sample size.

Sensitivity & specificity comparisons to see how performance differs:

```{r}
# Sensitivity (fake vids only)
emotion_fake <- filter(test, emotion == "yes", isfakeorreal == "fake")
no_emotion_fake <- filter(test, emotion == "no", isfakeorreal == "fake")
prop.test(c(sum(emotion_fake$pred_class == "fake"), sum(no_emotion_fake$pred_class == "fake")),
          c(nrow(emotion_fake), nrow(no_emotion_fake)))

# Specificity (real vids only)
emotion_real <- filter(test, emotion == "yes", isfakeorreal == "real")
no_emotion_real <- filter(test, emotion == "no", isfakeorreal == "real")
prop.test(c(sum(emotion_real$pred_class == "real"), sum(no_emotion_real$pred_class == "real")),
          c(nrow(emotion_real), nrow(no_emotion_real)))
```

Sensitivity (fake videos):

-   Emotion 0.655 vs no emotion 0.833: p = .321

-   No clear evidence of a difference.

Specificity (real videos):

-   Emotion 0.793 vs no emotion 0.278: p = .0014 (p \< .01)

-   Clear, significant drop in specificity on No Emotion. The model produces many more false positives (calls real videos fake) in no emotion videos.

-   Accuracy is lower in No Emotion mainly because specificity collapses, even though sensitivity is a bit higher.

On the held-out 20% test set, RF performance appears better on Emotion videos (Accuracy = 72.4% vs 55.6%; AUC = 0.77 vs 0.61), but neither difference is statistically significant (Accuracy Fisher p = .119; AUC DeLong p = .183).

However, specificity is significantly lower on No Emotion videos (0.278 vs 0.793; p = .001), indicating substantially more false positives on real videos without overt emotion.

Confusion Matrices:

```{r}
# Filter subsets
emotion_test <- dta_classifier_test_new_scaled %>% filter(emotion == "yes")
no_emotion_test <- dta_classifier_test_new_scaled %>% filter(emotion == "no")

# Model predictions
pred_emotion    <- predict(rf_fit_model1_cv, newdata = emotion_test)
pred_no_emotion <- predict(rf_fit_model1_cv, newdata = no_emotion_test)

# Raw confusion matrices
cm_emotion <- table(Predicted = pred_emotion, Actual = emotion_test$isfakeorreal)
cm_no_emotion <- table(Predicted = pred_no_emotion, Actual = no_emotion_test$isfakeorreal)

# Convert to tidy format
tidy_cm <- function(cm, title) {
  as.data.frame(cm) %>%
    mutate(Condition = title)
}

df_cm <- bind_rows(
  tidy_cm(cm_emotion,    "Emotion Videos"),
  tidy_cm(cm_no_emotion, "No Emotion Videos")
)

# Plot
p_cm <- df_cm %>%
  ggplot(aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), size = 6, fontface = "bold") +
  scale_fill_gradient(high = "#1b9e77", low = "#d95f02") +
  facet_wrap(~Condition, nrow = 1) +
  labs(title = "Confusion Matrices by Emotional Content",
       fill = "Count") +
  theme_classic(base_size = 14) +
  theme(
    strip.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "right"
  )

p_cm

```

Post hoc power analysis on accuracy comparison (two proportions test):

```{r}
library(pwr)

#observerd proportions (taken from above)
p1 <- 42/58  # emotion, correct/total
p2 <- 20/36  # no emotion

# effect size h
h <- ES.h(p1, p2)

power_result <- pwr.2p2n.test(
  h = h,
  n1 = 58,
  n2 = 36,
  sig.level = 0.05,
  alternative = "two.sided"
)

print(power_result)
print(paste("Cohen's h =", round(h, 3)))
print(paste("Achieved power =", round(power_result$power, 3)))
```

```         
   difference of proportion power calculation for binomial distribution (arcsine transformation) 

              h = 0.3534946
             n1 = 58
             n2 = 36
      sig.level = 0.05
          power = 0.38455
    alternative = two.sided

NOTE: different sample sizes

[1] "Cohen's h = 0.353"
[1] "Achieved power = 0.385"
```

### C5.0B

Emotion:

```{r}
c50_model_1_emotion <- caret::confusionMatrix(
  data = predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(c50_model_1_emotion)
```

Without CV:

```         
          Reference
Prediction real fake
      real   18    7
      fake   11   22
                                          
               Accuracy : 0.6897          
                 95% CI : (0.5546, 0.8046)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002677        
                                          
                  Kappa : 0.3793          
                                          
 Mcnemar's Test P-Value : 0.479500        
                                          
            Sensitivity : 0.7586          
            Specificity : 0.6207          
         Pos Pred Value : 0.6667          
         Neg Pred Value : 0.7200          
             Prevalence : 0.5000          
         Detection Rate : 0.3793          
   Detection Prevalence : 0.5690          
      Balanced Accuracy : 0.6897 
```

With CV:

```         
         Reference
Prediction real fake
      real   18    7
      fake   11   22
                                          
               Accuracy : 0.6897          
                 95% CI : (0.5546, 0.8046)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002677        
                                          
                  Kappa : 0.3793          
                                          
 Mcnemar's Test P-Value : 0.479500        
                                          
            Sensitivity : 0.7586          
            Specificity : 0.6207          
         Pos Pred Value : 0.6667          
         Neg Pred Value : 0.7200          
             Prevalence : 0.5000          
         Detection Rate : 0.3793          
   Detection Prevalence : 0.5690          
      Balanced Accuracy : 0.6897   
```

No Emotion:

```{r}
c50_model_1_no_emotion <- caret::confusionMatrix(
  data = predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(c50_model_1_no_emotion)
```

Without CV:

```         
          Reference
Prediction real fake
      real    6    3
      fake   12   15
                                          
               Accuracy : 0.5833          
                 95% CI : (0.4076, 0.7449)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.20252         
                                          
                  Kappa : 0.1667          
                                          
 Mcnemar's Test P-Value : 0.03887         
                                          
            Sensitivity : 0.8333          
            Specificity : 0.3333          
         Pos Pred Value : 0.5556          
         Neg Pred Value : 0.6667          
             Prevalence : 0.5000          
         Detection Rate : 0.4167          
   Detection Prevalence : 0.7500          
      Balanced Accuracy : 0.5833  
```

With CV:

Same.

### SVM

Emotion:

```{r}
svm_model_1_emotion <- caret::confusionMatrix(
  data = predict(svm_model1_cv, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(svm_model_1_emotion)
```

Without CV:

```         
          Reference
Prediction real fake
      real   22   12
      fake    7   17
                                          
               Accuracy : 0.6724          
                 95% CI : (0.5366, 0.7899)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.005964        
                                          
                  Kappa : 0.3448          
                                          
 Mcnemar's Test P-Value : 0.358795        
                                          
            Sensitivity : 0.5862          
            Specificity : 0.7586          
         Pos Pred Value : 0.7083          
         Neg Pred Value : 0.6471          
             Prevalence : 0.5000          
         Detection Rate : 0.2931          
   Detection Prevalence : 0.4138          
      Balanced Accuracy : 0.6724  
```

With CV:

```         
          Reference
Prediction real fake
      real   23   12
      fake    6   17
                                          
               Accuracy : 0.6897          
                 95% CI : (0.5546, 0.8046)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.002677        
                                          
                  Kappa : 0.3793          
                                          
 Mcnemar's Test P-Value : 0.238593        
                                          
            Sensitivity : 0.5862          
            Specificity : 0.7931          
         Pos Pred Value : 0.7391          
         Neg Pred Value : 0.6571          
             Prevalence : 0.5000          
         Detection Rate : 0.2931          
   Detection Prevalence : 0.3966          
      Balanced Accuracy : 0.6897   
```

No Emotion:

```{r}
svm_model_1_no_emotion <- caret::confusionMatrix(
  data = predict(svm_model1_cv, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(svm_model_1_no_emotion)
```

Without CV:

```         
          Reference
Prediction real fake
      real    6    6
      fake   12   12
                                          
               Accuracy : 0.5             
                 95% CI : (0.3292, 0.6708)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.5660          
                                          
                  Kappa : 0               
                                          
 Mcnemar's Test P-Value : 0.2386          
                                          
            Sensitivity : 0.6667          
            Specificity : 0.3333          
         Pos Pred Value : 0.5000          
         Neg Pred Value : 0.5000          
             Prevalence : 0.5000          
         Detection Rate : 0.3333          
   Detection Prevalence : 0.6667          
      Balanced Accuracy : 0.5000 
```

With CV:

```         
          Reference
Prediction real fake
      real    7    6
      fake   11   12
                                          
               Accuracy : 0.5278          
                 95% CI : (0.3549, 0.6959)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.434           
                                          
                  Kappa : 0.0556          
                                          
 Mcnemar's Test P-Value : 0.332           
                                          
            Sensitivity : 0.6667          
            Specificity : 0.3889          
         Pos Pred Value : 0.5217          
         Neg Pred Value : 0.5385          
             Prevalence : 0.5000          
         Detection Rate : 0.3333          
   Detection Prevalence : 0.6389          
      Balanced Accuracy : 0.5278  
```

### Logistic Regression

Emotion:

```{r}
lr_model_1_emotion <- caret::confusionMatrix(
  data = predict(lr_model1, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(lr_model_1_emotion)
```

With CV:

```         
          Reference
Prediction real fake
      real   17   12
      fake   12   17
                                         
               Accuracy : 0.5862         
                 95% CI : (0.4493, 0.714)
    No Information Rate : 0.5            
    P-Value [Acc > NIR] : 0.1185         
                                         
                  Kappa : 0.1724         
                                         
 Mcnemar's Test P-Value : 1.0000         
                                         
            Sensitivity : 0.5862         
            Specificity : 0.5862         
         Pos Pred Value : 0.5862         
         Neg Pred Value : 0.5862         
             Prevalence : 0.5000         
         Detection Rate : 0.2931         
   Detection Prevalence : 0.5000         
      Balanced Accuracy : 0.5862 
```

No emotion:

```{r}
lr_model_1_no_emotion <- caret::confusionMatrix(
  data = predict(lr_model1, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(lr_model_1_no_emotion)
```

With CV:

```         
          Reference
Prediction real fake
      real    7    6
      fake   11   12
                                          
               Accuracy : 0.5278          
                 95% CI : (0.3549, 0.6959)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.434           
                                          
                  Kappa : 0.0556          
                                          
 Mcnemar's Test P-Value : 0.332           
                                          
            Sensitivity : 0.6667          
            Specificity : 0.3889          
         Pos Pred Value : 0.5217          
         Neg Pred Value : 0.5385          
             Prevalence : 0.5000          
         Detection Rate : 0.3333          
   Detection Prevalence : 0.6389          
      Balanced Accuracy : 0.5278   
```

## 40 Held-Out Videos:

### RF Model 1

Emotion:

```{r}
dta_classifier_40_test_scaled_emotion <- dta_classifier_40_test_scaled %>%
  filter(emotion == "y")

rf_model_1_40_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled_emotion),
  reference = dta_classifier_40_test_scaled_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(rf_model_1_40_emotion)

# ROC AUC
y_test  <- factor(dta_classifier_40_test_scaled_emotion$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled_emotion, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

No emotion:

```{r}
dta_classifier_40_test_scaled_no_emotion <- dta_classifier_40_test_scaled %>%
  filter(emotion == "n")

rf_model_1_40_no_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled_no_emotion),
  reference = dta_classifier_40_test_scaled_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(rf_model_1_40_no_emotion)

# ROC AUC
y_test  <- factor(dta_classifier_40_test_scaled_no_emotion$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_cv, newdata = dta_classifier_40_test_scaled_no_emotion, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

Accuracy = 45.83%

95% CI = (0.2555, 0.6718)

NIR = 0.5

P = 0.7294

Kappa = -0.0833

Sensitivity = 41.67%

Specificity = 50%

### RF Model 2

Emotion:

```{r}
dta_classifier_40_test_new_all_emotion <- dta_classifier_40_test_new_all %>%
  filter(emotion == "y")

rf_model_2_40_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model2, newdata = dta_classifier_40_test_new_all_emotion),
  reference = dta_classifier_40_test_new_all_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(rf_model_2_40_emotion)
```

Accuracy = 43.75%

95% CI = (0.1975, 0.7012)

NIR = 0.5

P = 0.7728

Kappa = -0.125

Sensitivity = 62.50%

Specificity = 25%

No emotion:

```{r}
dta_classifier_40_test_new_all_no_emotion <- dta_classifier_40_test_new_all %>%
  filter(emotion == "n")

rf_model_2_40_no_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model2, newdata = dta_classifier_40_test_new_all_no_emotion),
  reference = dta_classifier_40_test_new_all_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(rf_model_2_40_no_emotion)
```

Accuracy = 62.50%

95% CI = (0.4059, 0.812)

NIR = 0.5

P = 0.1537

Kappa = 0.25

Sensitivity = 75%

Specificity = 50%

### C5.0B Model 1

Emotion:

```{r}
c50_model_1_40_emotion <- caret::confusionMatrix(
  data = predict(c50_fit_model1, newdata = dta_classifier_40_test_new_emotion),
  reference = dta_classifier_40_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(c50_model_1_40_emotion)
```

Accuracy = 50.00%

95% CI = (0.2465, 07535)

NIR = 0.5

P = 0.5982

Kappa = 0

Sensitivity = 1

Specificity = 0

No emotion:

```{r}
c50_model_1_40_no_emotion <- caret::confusionMatrix(
  data = predict(c50_fit_model1, newdata = dta_classifier_40_test_new_no_emotion),
  reference = dta_classifier_40_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print("--- No Emotion: ---")
print(c50_model_1_40_no_emotion)
```

Accuracy = 50.00%

95% CI = (0.2912, 0.7088)

NIR = 0.5

P = 0.5806

Kappa = 0

Sensitivity = 83.33%

Specificity = 16.67%

### C5.0B Model 2

Emotion:

```{r}
c50_model_2_40_emotion <- caret::confusionMatrix(
  data = predict(c50_fit_model2, newdata = dta_classifier_40_test_new_all_emotion),
  reference = dta_classifier_40_test_new_all_emotion$isfakeorreal,
  positive = "fake"
)

print("--- Emotion: ---")
print(c50_model_2_40_emotion)
```

# Additional Models

It is a good idea to train a few more classifiers (non-decision tree) to see how they compare to the current decision tree models. This allows us to see whether there is a significant difference between decision tree and non-decision tree classifiers.

## SVM

Support vector machines are a good choice as they are also non-linear but operate in a different way to decision trees (geometrically finds the optimal hyperplane that maximises the margin between different classes).

```{r}
# No CV setup
classifier_control <- caret::trainControl(
  method = "none",
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

svm_model1 <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "svmRadial",  # RBF kernel
  trControl = classifier_control
)
```

Evaluate on 20% test set:

```{r}
svm_model_1_performance <- caret::confusionMatrix(
  data = predict(svm_model1, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(svm_model_1_performance)
```

```         
          Reference
Prediction real fake
      real   29   18
      fake   18   29
                                         
               Accuracy : 0.617          
                 95% CI : (0.511, 0.7154)
    No Information Rate : 0.5            
    P-Value [Acc > NIR] : 0.01488        
                                         
                  Kappa : 0.234          
                                         
 Mcnemar's Test P-Value : 1.00000        
                                         
            Sensitivity : 0.6170         
            Specificity : 0.6170         
         Pos Pred Value : 0.6170         
         Neg Pred Value : 0.6170         
             Prevalence : 0.5000         
         Detection Rate : 0.3085         
   Detection Prevalence : 0.5000         
      Balanced Accuracy : 0.6170  
```

## Linear Regression

This will be the linear benchmark to compare our other non-linear model performances to. A comparable performance might suggest that the data is not non-linear and is best captured by linear models.

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "none",
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

lr_model1 <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "glm",  # Generalised linear models
  family = "binomial",
  trControl = classifier_control
)
```

Evaluate on test set:

```{r}
lr_model1_performance <- caret::confusionMatrix(
  data = predict(lr_model1, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(lr_model1_performance)
```

```         
          Reference
Prediction real fake
      real   24   18
      fake   23   29
                                          
               Accuracy : 0.5638          
                 95% CI : (0.4576, 0.6659)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.1282          
                                          
                  Kappa : 0.1277          
                                          
 Mcnemar's Test P-Value : 0.5322          
                                          
            Sensitivity : 0.6170          
            Specificity : 0.5106          
         Pos Pred Value : 0.5577          
         Neg Pred Value : 0.5714          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.5532          
      Balanced Accuracy : 0.5638  
```

Evaluate on 40 held out videos:

```{r}
lr_model1_40_performance <- caret::confusionMatrix(
  data = predict(lr_model1, newdata = dta_classifier_40_test_new_scaled),
  reference = dta_classifier_40_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(lr_model1_40_performance)
```

## SVM (CV)

```{r}
# Cross-valiation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

svm_model1_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "svmRadial",  # RBF kernel
  trControl = classifier_control
)


# No native feature importance function.
```

Evaluate on 20% test set:

```{r}
svm_model_1_cv_performance <- caret::confusionMatrix(
  data = predict(svm_model1_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(svm_model_1_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(svm_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

5 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   31   18
      fake   16   29
                                          
               Accuracy : 0.6383          
                 95% CI : (0.5327, 0.7349)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.004774        
                                          
                  Kappa : 0.2766          
                                          
 Mcnemar's Test P-Value : 0.863832        
                                          
            Sensitivity : 0.6170          
            Specificity : 0.6596          
         Pos Pred Value : 0.6444          
         Neg Pred Value : 0.6327          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.4787          
      Balanced Accuracy : 0.6383  
```

10 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   30   18
      fake   17   29
                                          
               Accuracy : 0.6277          
                 95% CI : (0.5218, 0.7252)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.008605        
                                          
                  Kappa : 0.2553          
                                          
 Mcnemar's Test P-Value : 1.000000        
                                          
            Sensitivity : 0.6170          
            Specificity : 0.6383          
         Pos Pred Value : 0.6304          
         Neg Pred Value : 0.6250          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.4894          
      Balanced Accuracy : 0.6277 
```

## Logistic Regression (CV)

```{r}
# Cross-valiation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

lr_model1_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "glm",  # Generalised linear models
  family = "binomial",
  trControl = classifier_control
)

plot(caret::varImp(lr_model1_cv)) # To see feature importance
```

Evaluate on 20% test set:

```{r}
lr_model1_cv_performance <- caret::confusionMatrix(
  data = predict(lr_model1_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(lr_model1_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(lr_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

5 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   24   18
      fake   23   29
                                          
               Accuracy : 0.5638          
                 95% CI : (0.4576, 0.6659)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.1282          
                                          
                  Kappa : 0.1277          
                                          
 Mcnemar's Test P-Value : 0.5322          
                                          
            Sensitivity : 0.6170          
            Specificity : 0.5106          
         Pos Pred Value : 0.5577          
         Neg Pred Value : 0.5714          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.5532          
      Balanced Accuracy : 0.5638          
                                          
       'Positive' Class : fake    
```

10 folds, 3 repeats:

```         
          Reference
Prediction real fake
      real   24   18
      fake   23   29
                                          
               Accuracy : 0.5638          
                 95% CI : (0.4576, 0.6659)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.1282          
                                          
                  Kappa : 0.1277          
                                          
 Mcnemar's Test P-Value : 0.5322          
                                          
            Sensitivity : 0.6170          
            Specificity : 0.5106          
         Pos Pred Value : 0.5577          
         Neg Pred Value : 0.5714          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.5532          
      Balanced Accuracy : 0.5638 
```

## LDA (CV)

Linear Discriminant Analysis is another linear model to compare against Logistic Regression. It assumes features are normally distributed and have equal variance.

```{r}
# Cross-valiation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary,
  savePredictions = "final"
)

set.seed(12345)

lda_model_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_scaled[, -(1:2)],
  method = "lda",
  metric = "ROC",
  trControl = classifier_control
)
```

Evaluate on 20% test set:

```{r}
lda_model_cv_performance <- caret::confusionMatrix(
  data = predict(lda_model_cv, newdata = dta_classifier_test_new_scaled, type="raw"),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(lda_model_cv_performance)
```

```         


          Reference
Prediction fake real
      fake   29   22
      real   18   25
                                          
               Accuracy : 0.5745          
                 95% CI : (0.4682, 0.6759)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.08983         
                                          
                  Kappa : 0.1489          
                                          
 Mcnemar's Test P-Value : 0.63526         
                                          
            Sensitivity : 0.6170          
            Specificity : 0.5319          
         Pos Pred Value : 0.5686          
         Neg Pred Value : 0.5814          
             Prevalence : 0.5000          
         Detection Rate : 0.3085          
   Detection Prevalence : 0.5426          
      Balanced Accuracy : 0.5745          
                                          
       'Positive' Class : fake    
```

# Figures

Human vs model performance bar chart:

```{r fig.width=6, fig.height=6}
ggplot(
  data = data.frame(
    Stimuli = factor(
      c("Full Stimuli", "Point Light Display", "Point Light Display"),
      levels = c("Full Stimuli", "Point Light Display")
    ),
    Method = factor(
      c("Human Observers", "Human Observers", "Random Forest"),
      levels = c("Human Observers", "Random Forest")
    ),
    Accuracy = c(88.33, 50.51, 55.00),
    CI_low  = c(80.0,  40.3,  39.8),
    CI_high = c(93.5,  60.7,  69.3)
  ),
  aes(x = Stimuli, y = Accuracy, fill = Method)
) +
  geom_col(position = position_dodge(width = 0.65), width = 0.6) +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    position = position_dodge(width = 0.65),
    width = 0.2, size = 0.8
  ) +
  geom_hline(yintercept = 50, linetype = "dashed",
             colour = "red", size = 0.6, alpha = 0.6) +
  annotate("text", x = 2, y = 50, label = "NIR (0.5)",
           hjust = -1.11, vjust = -0.6, size = 4, colour = "red") +
  scale_y_continuous(limits = c(0, 100),
                     expand = expansion(mult = c(0, 0.08))) +
  scale_fill_manual(values = c("Human Observers" = "gray40",
                               "Random Forest"   = "gray70")) +
  labs(x = NULL, y = "Accuracy (%)", fill = NULL) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 18),
    axis.title.y = element_text(size = 18, face = "bold"),
    axis.text.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    plot.margin  = margin(10, 10, 10, 10)
  )
```

ROC curve figure:

```{r, fig.width=8, fig.height=6}
# Probabilities on the same test set
p_rf  <- predict(rf_fit_model1_cv,  newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
p_c50 <- predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
p_svm <- predict(svm_model1_cv,     newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
y_test <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))

roc_df <- function(truth, prob, name){
  r <- pROC::roc(truth, prob, levels = c("real","fake"), direction = "<", quiet = TRUE)
  c <- pROC::coords(r, x = "all", ret = c("specificity","sensitivity","threshold"), transpose = FALSE)
  c$model <- name
  c$FPR <- 1 - c$specificity
  c$AUC <- as.numeric(pROC::auc(r))
  c
}

roc_all <- bind_rows(
  roc_df(y_test, p_rf,  "Random Forest"),
  roc_df(y_test, p_c50, "C5.0"),
  roc_df(y_test, p_svm, "SVM (RBF)")
)

# Legend labels with AUC
auc_labels <- roc_all |>
  group_by(model) |>
  summarise(AUC = unique(round(AUC, 3)), .groups = "drop") |>
  mutate(label = paste0(model, " (AUC=", sprintf("%.3f", AUC), ")"))
label_map <- setNames(auc_labels$label, auc_labels$model)

ggplot(roc_all, aes(x = FPR, y = sensitivity, color = model)) +
  geom_line(linewidth = 1, alpha = 0.8) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.6, colour = "grey60") +
  coord_equal(xlim = c(0,1), ylim = c(0,1), expand = FALSE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1),
                     labels = scales::number_format(accuracy = 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1),
                     labels = scales::number_format(accuracy = 0.1)) +
  scale_color_brewer(type = "qual", palette = "Dark2",
                     labels = label_map, breaks = names(label_map)) +
  labs(
    x = "False Positive Rate",
    y = "True Positive Rate (Sensitivity)",
    color = NULL,
    title = "ROC Curves on 20% Test Set",
    subtitle = "Random Forest, C5.0, and SVM (positive class = fake)"
  ) +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold")
  )
```

```{r, fig.width=5, fig.height=5}
library(ROCR)

# Extract predicted probabilities (fake class)
rf_probs  <- predict(rf_fit_model1_cv,  newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
c50_probs <- predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
svm_probs <- predict(svm_model1_cv,     newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
lr_probs <- predict(lr_model1_cv,     newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

# True outcome (must be numeric 0/1)
true_class <- ifelse(dta_classifier_test_new_scaled$isfakeorreal == "fake", 1, 0)

# Build ROCR prediction objects
pred_rf  <- prediction(rf_probs,  true_class)
pred_c50 <- prediction(c50_probs, true_class)
pred_svm <- prediction(svm_probs, true_class)
pred_lr <- prediction(lr_probs, true_class)

# Performance objects for TPR/FPR
perf_rf  <- performance(pred_rf,  "tpr", "fpr")
perf_c50 <- performance(pred_c50, "tpr", "fpr")
perf_svm <- performance(pred_svm, "tpr", "fpr")
perf_lr <- performance(pred_lr, "tpr", "fpr")

# Compute AUCs
auc_rf  <- performance(pred_rf,  "auc")@y.values[[1]]
auc_c50 <- performance(pred_c50, "auc")@y.values[[1]]
auc_svm <- performance(pred_svm, "auc")@y.values[[1]]
auc_lr <- performance(pred_lr, "auc")@y.values[[1]]


# Save to PDF for paper
pdf("fig2_panel_B.pdf", width = 5, height = 5)
par(mar = c(4, 4, 3, 1), mgp = c(2, 0.7, 0))

# Plot base ROC curvs
plot(perf_rf,  col = "#E69F00", lwd = 2,
     main = "ROC Curves on Test Set (n=94)",
     xlab = "False Positive Rate", ylab = "True Positive Rate")

plot(perf_c50, col = "#009E73", lwd = 2, add = TRUE)
plot(perf_svm, col = "#CC79A7",   lwd = 2, add = TRUE)
plot(perf_lr, col = "#0072B2", lwd = 2, add = TRUE)

# Diagonal reference line
abline(a = 0, b = 1, col = "#999999", lty = 2)

# Legend with AUC values
legend("bottomright",
       legend = c(
         paste0("Random Forest (AUC = ", sprintf("%.3f", auc_rf), ")"),
         paste0("C5.0 (AUC = ", sprintf("%.3f", auc_c50), ")"),
         paste0("SVM (RBF) (AUC = ", sprintf("%.3f", auc_svm), ")"),
         paste0("LR (AUC = ", sprintf("%.3f", auc_lr), ")")
       ),
       col = c("#E69F00", "#009E73", "#CC79A7", "#0072B2"),
       lwd = 2, cex = 0.9, bty = "n")

dev.off()
```

Class separability:

```{r, fig.width = 5, fig.height = 5}
rf_probs_df <- dta_classifier_test_new_scaled %>%
  mutate(
    truth     = factor(isfakeorreal, levels = c("real","fake")),
    p_fake_rf = predict(rf_fit_model1_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]
  ) %>%
  select(truth, p_fake_rf)


# Save to PDF for paper
pdf("fig2_panel_C.pdf", width = 5, height = 5)
par(mar = c(4, 4, 3, 1))

# Plot densities
ggplot(rf_probs_df, aes(x = p_fake_rf, fill = truth, linetype = truth)) +
  geom_density(alpha = 0.4, adjust = 1, color = "black", linewidth = 0.8) +
  geom_vline(xintercept = 0.5, linewidth = 0.7, linetype = "dotted") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_fill_manual(
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    name = NULL, labels = c("Real", "Fake")
  ) +
  scale_linetype_manual(
    values = c("real" = "solid", "fake" = "dashed"),
    name = NULL, labels = c("Real", "Fake")
  ) +
  labs(
    x = "Predicted probability of 'fake' (Random Forest)",
    y = "Density",
    title = stringr::str_wrap("Distribution of Predicted Probabilities by True Class",
                              width = 40)
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 14)
  )
dev.off()

```

Decision boundary map:

```{r, fig.width=5, fig.height=5}
# two features to plot
f1 <- "diff1x_pacf5_AU17_chin_raiser"
f2 <- "diff1x_pacf5_AU01_inner_brow_raiser"

df_plot <- dta_classifier_test_new_scaled %>%
  mutate(truth = factor(isfakeorreal, levels = c("real","fake")))

# Dense grid covering the observed feature range
make_seq <- function(x, n = 400) {
  rng <- range(x, na.rm = TRUE)
  # tiny expansion to avoid white edge slivers
  seq(rng[1] - 1e-6, rng[2] + 1e-6, length.out = n)
}

grid_2d <- expand.grid(
  setNames(list(make_seq(df_plot[[f1]]), make_seq(df_plot[[f2]])), c(f1, f2)),
  KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE
)

# Build full newdata with all predictors present
proto <- dta_classifier_test_new_scaled[1, , drop = FALSE]
newdata_full <- proto[rep(1, nrow(grid_2d)), , drop = FALSE]
newdata_full[[f1]] <- grid_2d[[f1]]
newdata_full[[f2]] <- grid_2d[[f2]]

# Predict CLASS for background and PROB for the contour
#    type = "raw" -> class labels; type = "prob" -> probabilities
grid_2d$pred_class <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "raw")
probs <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "prob")
grid_2d$prob_fake <- probs[, "fake"]

grid_2d$pred_class <- factor(grid_2d$pred_class, levels = c("real","fake"))

# Plot categorical decision regions + 0.5 contour + points
ggplot() +
  # categorical regions (background)
  geom_raster(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], fill = pred_class),
    alpha = 0.45
  ) +
  # decision boundary at P(fake)=0.5
  geom_contour(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], z = prob_fake),
    breaks = 0.5,
    colour = "black",
    linewidth = 0.85
  ) +
  # data points: shape and fill both mapped to same variable
  geom_point(
    data = df_plot,
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 2.8,
    colour = "black",
    stroke = 0.4
  ) +
  # unified scales for fill and shape
  scale_fill_manual(
    name = "",
    values = c("real" = "#1b9e77", "fake" = "#d95f02"),
    labels = c("Real", "Fake")
  ) +
  scale_shape_manual(
    name = "",
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake")
  ) +
  labs(
    x = f1, y = f2,
    title = "Random Forest Decision Boundaries"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```

Same as before but including training data points:

```{r, fig.width=5, fig.height=5}
library(ggnewscale)

# two features to plot
f1 <- "diff1x_pacf5_AU17_chin_raiser"
f2 <- "diff1x_pacf5_AU01_inner_brow_raiser"

# Data for TEST points
df_plot_test <- dta_classifier_test_new_scaled %>%
  mutate(truth = factor(isfakeorreal, levels = c("real","fake")))

# Data for TRAINING points
df_plot_train <- dta_classifier_train_new_scaled %>%
  mutate(truth = factor(isfakeorreal, levels = c("real","fake")))

# Dense grid covering the observed feature range
make_seq <- function(x, var_name, n = 400) {
  data_rng <- range(x, na.rm = TRUE)
  
  # Check the variable name string to set plot-specific ranges
  plot_rng <- if (identical(f1, var_name)) {
    c(-2, 2)
  } else if (identical(f2, var_name)) {
    c(-4, 2)
  } else {
    data_rng
  }
  
  # Use the widest range for the grid
  final_rng <- c(min(data_rng[1], plot_rng[1]), max(data_rng[2], plot_rng[2]))
  
  seq(final_rng[1] - 1e-6, final_rng[2] + 1e-6, length.out = n)
}
grid_2d <- expand.grid(
  setNames(list(make_seq(df_plot_test[[f1]], var_name = f1), 
                make_seq(df_plot_test[[f2]], var_name = f2)), 
           c(f1, f2)),
  KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE
)
proto <- dta_classifier_test_new_scaled[1, , drop = FALSE]
newdata_full <- proto[rep(1, nrow(grid_2d)), , drop = FALSE]
newdata_full[[f1]] <- grid_2d[[f1]]
newdata_full[[f2]] <- grid_2d[[f2]]
grid_2d$pred_class <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "raw")
probs <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "prob")
grid_2d$prob_fake <- probs[, "fake"]
grid_2d$pred_class <- factor(grid_2d$pred_class, levels = c("real","fake"))



# Save to PDF for paper
pdf("fig2_panel_E.pdf", width = 5, height = 5)
par(mar = c(4, 4, 3, 1))

# Plot
ggplot() +
  # categorical regions (background)
  geom_raster(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], fill = pred_class),
    alpha = 0.45,
    show.legend = FALSE 
  ) +
  # decision boundary at P(fake)=0.5
  geom_contour(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], z = prob_fake),
    breaks = 0.5,
    colour = "black",
    linewidth = 0.85
  ) +

  # TEST DATA POINTS & LEGEND
  geom_point(
    data = df_plot_test,
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 2.8,
    colour = "black",
    stroke = 0.4
  ) +
  # Scales for Test Data
  scale_fill_manual(
    name = "Test",
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 2.8, alpha = 1))
  ) +
  scale_shape_manual(
    name = "Test",
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 2.8, alpha = 1))
  ) +

  ggnewscale::new_scale_fill() +
  ggnewscale::new_scale("shape") +

  geom_point(
    data = df_plot_train, 
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 1.5,
    alpha = 0.4,
    colour = "#000000",
    stroke = 0.2
  ) +
  scale_fill_manual(
    name = "Train",
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 1.5, alpha = 0.4))
  ) +
  scale_shape_manual(
    name = "Train",
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 1.5, alpha = 0.4))
  ) +

  coord_cartesian(
    xlim = c(-2, 2),
    ylim = c(-4, 2),
    expand = FALSE
  ) +
  labs(
    x = f1, y = f2,
    title = "Random Forest Decision Boundaries"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

dev.off()
```

Feature importance for RF Model 1:

```{r, fig.width=5, fig.height=5}
# Extract variable importance
vi <- caret::varImp(rf_fit_model1_cv)
vi_df <- vi$importance %>%
  tibble::rownames_to_column("Feature") %>%
  arrange(desc(Overall))
# Shorten feature names
vi_df <- vi_df %>%
  mutate(Feature = stringr::str_replace(Feature, "(AU\\d+)_.*", "\\1"))


# Save to PDF for paper
pdf("fig2_panel_D.pdf", width = 5, height = 5)
par(mar = c(4, 4, 3, 1))

ggplot(vi_df, aes(x = reorder(Feature, Overall), y = Overall)) +
  geom_col(fill = "#009E73") +
  geom_text(aes(label = round(Overall, 2)), hjust = -0.2, size = 3.5, fontface = "bold") +
  coord_flip(clip = "off") +
  labs(
    title = "Feature Importance - Random Forest",
    x = NULL,
    y = "Importance"
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 11, colour = "black")
  )

dev.off()
```

Classifier confusion matrices:

```{r, fig.width=10, fig.height=4}
# Get predictions for all 4 models on the full test set
pred_rf  <- predict(rf_fit_model1_cv,  newdata = dta_classifier_test_new_scaled)
pred_c50 <- predict(c50_fit_model1_cv, newdata = dta_classifier_test_new_scaled)
pred_svm <- predict(svm_model1_cv,     newdata = dta_classifier_test_new_scaled)
pred_lr  <- predict(lr_model1_cv,      newdata = dta_classifier_test_new_scaled)

# Get the true, actual class
actual_truth <- dta_classifier_test_new_scaled$isfakeorreal

# Create the raw confusion matrix tables
cm_rf  <- table(Predicted = pred_rf,  Actual = actual_truth)
cm_c50 <- table(Predicted = pred_c50, Actual = actual_truth)
cm_svm <- table(Predicted = pred_svm, Actual = actual_truth)
cm_lr  <- table(Predicted = pred_lr,  Actual = actual_truth)

# Tidy function
tidy_cm <- function(cm, title) {
  as.data.frame(cm) %>%
    mutate(Model = title) 
}

# Bind all 4 tables into one data frame
df_cm_models <- bind_rows(
  tidy_cm(cm_rf,  "Random Forest"),
  tidy_cm(cm_c50, "C5.0"),
  tidy_cm(cm_svm, "SVM (RBF)"),
  tidy_cm(cm_lr,  "Logistic Regression")
)

# Optional: Set factor order for the plot
df_cm_models$Model <- factor(df_cm_models$Model, 
    levels = c("Random Forest", "C5.0", "SVM (RBF)", "Logistic Regression"))

# Calculate percentages PER MODEL
df_cm_models <- df_cm_models %>%
  group_by(Model) %>%
  mutate(
    Total = sum(Freq),
    Percent = Freq / Total
  ) %>%
  ungroup()



# Save to PDF for paper
pdf("fig2_panel_F.pdf", width = 10, height = 4)
par(mar = c(4, 4, 3, 1))

ggplot(df_cm_models, aes(x = Actual, y = Predicted, fill = Freq)) +
  # Use black outlines
  geom_tile(color = "black", linewidth = 0.6) + 
  
  # Use sprintf to add percentages
  geom_text(aes(label = sprintf("%d\n(%.0f%%)", Freq, 100*Percent)),
            size = 5, lineheight = 0.95, color = "black") + 
            
  scale_fill_gradient(high = "#009E73", low = "white") +
  
  # Add coord_equal for square tiles
  coord_equal() + 
  
  facet_wrap(~Model, nrow = 1) + 
  labs(title = "Confusion Matrices by Classifier",
       subtitle = "Performance on Test Set (n=94)",
       fill = "Count") +
  theme_classic(base_size = 14) +
  
  # Updated theme to match agreement matrices
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    legend.position = "right",
    axis.title = element_text(face = "bold", color = "black"),
    axis.text = element_text(face = "bold", color = "black")
  )
dev.off()
```

ROC Curves by Emotion:

```{r, fig.width = 10, fig.height=5}
library(ROCR) 

emotion_test_set <- dta_classifier_test_new_scaled %>% filter(emotion == "yes")
no_emotion_test_set <- dta_classifier_test_new_scaled %>% filter(emotion == "no")

# Extract predicted probabilities (fake class)
rf_probs_em  <- predict(rf_fit_model1_cv,  newdata = emotion_test_set, type = "prob")[,"fake"]
c50_probs_em <- predict(c50_fit_model1_cv, newdata = emotion_test_set, type = "prob")[,"fake"]
svm_probs_em <- predict(svm_model1_cv,     newdata = emotion_test_set, type = "prob")[,"fake"]
lr_probs_em <- predict(lr_model1_cv, newdata = emotion_test_set, type = "prob")[,"fake"]

# True outcome (must be numeric 0/1)
true_class_em <- ifelse(emotion_test_set$isfakeorreal == "fake", 1, 0)

# Build ROCR prediction objects
pred_rf_em  <- prediction(rf_probs_em,  true_class_em)
pred_c50_em <- prediction(c50_probs_em, true_class_em)
pred_svm_em <- prediction(svm_probs_em, true_class_em)
pred_lr_em <- prediction(lr_probs_em, true_class_em)

# Performance objects for TPR/FPR
perf_rf_em  <- performance(pred_rf_em,  "tpr", "fpr")
perf_c50_em <- performance(pred_c50_em, "tpr", "fpr")
perf_svm_em <- performance(pred_svm_em, "tpr", "fpr")
perf_lr_em <- performance(pred_lr_em, "tpr", "fpr")

# Compute AUCs
auc_rf_em  <- performance(pred_rf_em,  "auc")@y.values[[1]]
auc_c50_em <- performance(pred_c50_em, "auc")@y.values[[1]]
auc_svm_em <- performance(pred_svm_em, "auc")@y.values[[1]]
auc_lr_em <- performance(pred_lr_em, "auc")@y.values[[1]]

# Extract predicted probabilities (fake class)
rf_probs_no_em  <- predict(rf_fit_model1_cv,  newdata = no_emotion_test_set, type = "prob")[,"fake"]
c50_probs_no_em <- predict(c50_fit_model1_cv, newdata = no_emotion_test_set, type = "prob")[,"fake"]
svm_probs_no_em <- predict(svm_model1_cv,     newdata = no_emotion_test_set, type = "prob")[,"fake"]
lr_probs_no_em <- predict(lr_model1_cv, newdata = no_emotion_test_set, type = "prob")[,"fake"]

# True outcome (must be numeric 0/1)
true_class_no_em <- ifelse(no_emotion_test_set$isfakeorreal == "fake", 1, 0)

# Build ROCR prediction objects
pred_rf_no_em  <- prediction(rf_probs_no_em,  true_class_no_em)
pred_c50_no_em <- prediction(c50_probs_no_em, true_class_no_em)
pred_svm_no_em <- prediction(svm_probs_no_em, true_class_no_em)
pred_lr_no_em <- prediction(lr_probs_no_em, true_class_no_em)

# Performance objects for TPR/FPR
perf_rf_no_em  <- performance(pred_rf_no_em,  "tpr", "fpr")
perf_c50_no_em <- performance(pred_c50_no_em, "tpr", "fpr")
perf_svm_no_em <- performance(pred_svm_no_em, "tpr", "fpr")
perf_lr_no_em <- performance(pred_lr_no_em, "tpr", "fpr")

# Compute AUCs
auc_rf_no_em  <- performance(pred_rf_no_em,  "auc")@y.values[[1]]
auc_c50_no_em <- performance(pred_c50_no_em, "auc")@y.values[[1]]
auc_svm_no_em <- performance(pred_svm_no_em, "auc")@y.values[[1]]
auc_lr_no_em <- performance(pred_lr_no_em, "auc")@y.values[[1]]


# Plot
# Save to PDF for paper
pdf("fig3_panel_C_D.pdf", width = 10, height = 5)
par(mfrow = c(1, 2), mar = c(5, 4, 4, 2) + 0.1, mgp = c(1.8, 0.7, 0)) # mar adjusts margins

# Emotion vids first
plot(perf_rf_em,  col = "#E69F00", lwd = 2,
     main = "Emotion Videos (n=58)",
     xlab = "False Positive Rate", ylab = "True Positive Rate")

plot(perf_c50_em, col = "#009E73", lwd = 2, add = TRUE)
plot(perf_svm_em, col = "#CC79A7",   lwd = 2, add = TRUE)
plot(perf_lr_em, col = "#0072B2", lwd = 2, add = TRUE)

# Diagonal reference line
abline(a = 0, b = 1, col = "#999999", lty = 2)

# Legend with AUC values
legend("bottomright",
       legend = c(
         paste0("Random Forest (AUC = ", sprintf("%.3f", auc_rf_em), ")"),
         paste0("C5.0 (AUC = ", sprintf("%.3f", auc_c50_em), ")"),
         paste0("SVM (RBF) (AUC = ", sprintf("%.3f", auc_svm_em), ")"),
         paste0("LR (AUC =", sprintf("%.3f", auc_lr_em), ")")
       ),
       col = c("#E69F00", "#009E73", "#CC79A7", "#0072B2"),
       lwd = 2, cex = 0.8, bty = "n") # cex=0.8 makes text slightly smaller


# No emotion vids
plot(perf_rf_no_em,  col = "#E69F00", lwd = 2,
     main = "No Emotion Videos (n=36)",
     xlab = "False Positive Rate", ylab = "True Positive Rate")

plot(perf_c50_no_em, col = "#009E73", lwd = 2, add = TRUE)
plot(perf_svm_no_em, col = "#CC79A7",   lwd = 2, add = TRUE)
plot(perf_lr_no_em, col = "#0072B2", lwd = 2, add = TRUE)

# Diagonal reference line
abline(a = 0, b = 1, col = "#999999", lty = 2)

# Legend with AUC values
legend("bottomright",
       legend = c(
         paste0("Random Forest (AUC = ", sprintf("%.3f", auc_rf_no_em), ")"),
         paste0("C5.0 (AUC = ", sprintf("%.3f", auc_c50_no_em), ")"),
         paste0("SVM (RBF) (AUC = ", sprintf("%.3f", auc_svm_no_em), ")"),
         paste0("LR (AUC = ", sprintf("%.3f", auc_lr_no_em), ")")
       ),
       col = c("#E69F00", "#009E73", "#CC79A7", "#0072B2"),
       lwd = 2, cex = 0.8, bty = "n")

# Reset plotting area to default (1x1)
par(mfrow = c(1, 1))
dev.off()
```

Density plots by Emotion:

```{r, fig.width=10, fig.height=5}
emotion_test_set <- dta_classifier_test_new_scaled %>% filter(emotion == "yes")
no_emotion_test_set <- dta_classifier_test_new_scaled %>% filter(emotion == "no")

# Emotion vids
rf_probs_df_em <- emotion_test_set %>%
  mutate(
    truth     = factor(isfakeorreal, levels = c("real","fake")),
    p_fake_rf = predict(rf_fit_model1_cv, newdata = ., type = "prob")[,"fake"]
  ) %>%
  select(truth, p_fake_rf)

# Calculate densities
# We use from=0, to=1 to keep the density calculation within the 0-1 probability range
dens_em_real <- density(rf_probs_df_em$p_fake_rf[rf_probs_df_em$truth == "real"], from = 0, to = 1, adjust = 1)
dens_em_fake <- density(rf_probs_df_em$p_fake_rf[rf_probs_df_em$truth == "fake"], from = 0, to = 1, adjust = 1)


# No Emotion vids
rf_probs_df_no_em <- no_emotion_test_set %>%
  mutate(
    truth     = factor(isfakeorreal, levels = c("real","fake")),
    p_fake_rf = predict(rf_fit_model1_cv, newdata = ., type = "prob")[,"fake"]
  ) %>%
  select(truth, p_fake_rf)

# Calculate densities
dens_no_em_real <- density(rf_probs_df_no_em$p_fake_rf[rf_probs_df_no_em$truth == "real"], from = 0, to = 1, adjust = 1)
dens_no_em_fake <- density(rf_probs_df_no_em$p_fake_rf[rf_probs_df_no_em$truth == "fake"], from = 0, to = 1, adjust = 1)

# Plot
# Save to PDF for paper
pdf("fig3_panel_E_F.pdf", width = 10, height = 5)

par(mfrow = c(1, 2), mar = c(5, 4, 4, 2) + 0.1, mgp = c(1.7, 0.7, 0))

col_real <- "#009E73"
col_fake <- "#E69F00"
# Create transparent versions
col_real_fill <- adjustcolor(col_real, alpha.f = 0.4)
col_fake_fill <- adjustcolor(col_fake, alpha.f = 0.4)


# Get Y-axis limits to fit both densities
#ylim_em <- range(0, dens_em_real$y, dens_em_fake$y)

# Shared y-axis limit for both plots
ylim_shared <- range(0, dens_em_real$y, dens_em_fake$y, 
                        dens_no_em_real$y, dens_no_em_fake$y)

# Set up the empty plot
plot(NULL,
     main = "Emotion Videos (n=58)",
     xlab = "Predicted probability of 'fake'", ylab = "Density",
     xlim = c(0, 1), ylim = ylim_shared
)

# Add shaded polygons
polygon(dens_em_real, col = col_real_fill, border = NA)
polygon(dens_em_fake, col = col_fake_fill, border = NA)

# Add lines on top
lines(dens_em_real, lty = "solid", lwd = 2, col = "black")
lines(dens_em_fake, lty = "dashed", lwd = 2, col = "black")

# Add vline and legend
abline(v = 0.5, lty = "dotted", lwd = 2, col = "black")
legend("topright",
       legend = c("Real", "Fake"),
       fill = c(col_real_fill, col_fake_fill),
       border = c("black", "black"),
       lty = c("solid", "dashed"),
       lwd = 2,
       bty = "n", 
       cex = 0.9
)


# Get Y-axis limits
#ylim_no_em <- range(0, dens_no_em_real$y, dens_no_em_fake$y)

# Set up the empty plot
plot(NULL,
     main = "No Emotion Videos (n=36)",
     xlab = "Predicted probability of 'fake'", ylab = "Density",
     xlim = c(0, 1), ylim = ylim_shared
)

# Add shaded polygons
polygon(dens_no_em_real, col = col_real_fill, border = NA)
polygon(dens_no_em_fake, col = col_fake_fill, border = NA)

# Add lines on top
lines(dens_no_em_real, lty = "solid", lwd = 2, col = "black")
lines(dens_no_em_fake, lty = "dashed", lwd = 2, col = "black")

# Add vline and legend
abline(v = 0.5, lty = "dotted", lwd = 2, col = "black")
legend("topright",
       legend = c("Real", "Fake"),
       fill = c(col_real_fill, col_fake_fill),
       border = c("black", "black"),
       lty = c("solid", "dashed"),
       lwd = 2,
       bty = "n",
       cex = 0.9
)

# Reset plotting area to default (1x1)
par(mfrow = c(1, 1))
dev.off()
```

Decision boundary map by Emotion:

```{r, fig.width=8, fig.height=5}
# two features to plot (same as before)
f1 <- "diff1x_pacf5_AU17_chin_raiser"
f2 <- "diff1x_pacf5_AU01_inner_brow_raiser"

# grid calculation
make_seq <- function(x, n = 400) {
  rng <- range(x, na.rm = TRUE)
  seq(rng[1] - 1e-6, rng[2] + 1e-6, length.out = n)
}

grid_2d <- expand.grid(
  setNames(list(make_seq(dta_classifier_test_new_scaled[[f1]]), 
                make_seq(dta_classifier_test_new_scaled[[f2]])), c(f1, f2)),
  KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE
)

proto <- dta_classifier_test_new_scaled[1, , drop = FALSE]
newdata_full <- proto[rep(1, nrow(grid_2d)), , drop = FALSE]
newdata_full[[f1]] <- grid_2d[[f1]]
newdata_full[[f2]] <- grid_2d[[f2]]

grid_2d$pred_class <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "raw")
probs <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "prob")
grid_2d$prob_fake <- probs[, "fake"]
grid_2d$pred_class <- factor(grid_2d$pred_class, levels = c("real","fake"))


df_plot_faceted <- dta_classifier_test_new_scaled %>%
  mutate(
    truth = factor(isfakeorreal, levels = c("real","fake")),
    Condition = factor(emotion, 
                       levels = c("yes", "no"), 
                       labels = c("Emotion Videos (n=58)", "No Emotion Videos (n=36)"))
  )


# Plot
ggplot() +
  # background
  geom_raster(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], fill = pred_class),
    alpha = 0.45
  ) +
  # boundary
  geom_contour(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], z = prob_fake),
    breaks = 0.5,
    colour = "black",
    linewidth = 0.85
  ) +
  # data points
  geom_point(
    data = df_plot_faceted, 
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 2.8,
    colour = "black",
    stroke = 0.4
  ) +
  
  scale_fill_manual(
    name = "",
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    labels = c("Real", "Fake")
  ) +
  scale_shape_manual(
    name = "",
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake")
  ) +
  
  facet_wrap(~ Condition) +
  
  labs(
    x = f1, y = f2,
    title = "Random Forest Decision Boundaries by Emotional Content"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 12)
  )
```

Decision boundary map by Emotion with train data overlay:

```{r, fig.width=8, fig.height=5}
library(ggnewscale)

# two features to plot
f1 <- "diff1x_pacf5_AU17_chin_raiser"
f2 <- "diff1x_pacf5_AU01_inner_brow_raiser"

df_plot_test <- dta_classifier_test_new_scaled %>%
  mutate(
    truth = factor(isfakeorreal, levels = c("real","fake")),
    Condition = factor(emotion, 
                       levels = c("yes", "no"), 
                       labels = c("Emotion Videos (n=58)", "No Emotion Videos (n=36)")) 
  )

df_plot_train <- dta_classifier_train_new_scaled %>%
  mutate(
    truth = factor(isfakeorreal, levels = c("real","fake")),
    Condition = factor(emotion, 
                       levels = c("yes", "no"), 
                       labels = c("Emotion Videos (n=58)", "No Emotion Videos (n=36)"))
  )
  
make_seq <- function(x, var_name, n = 400) {
  data_rng <- range(x, na.rm = TRUE)
  plot_rng <- if (identical(f1, var_name)) {
    c(-2, 2)
  } else if (identical(f2, var_name)) {
    c(-4, 2)
  } else {
    data_rng
  }
  final_rng <- c(min(data_rng[1], plot_rng[1]), max(data_rng[2], plot_rng[2]))
  seq(final_rng[1] - 1e-6, final_rng[2] + 1e-6, length.out = n)
}
grid_2d <- expand.grid(
  setNames(list(make_seq(dta_classifier_test_new_scaled[[f1]], var_name = f1), 
                make_seq(dta_classifier_test_new_scaled[[f2]], var_name = f2)), 
           c(f1, f2)),
  KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE
)
proto <- dta_classifier_test_new_scaled[1, , drop = FALSE]
newdata_full <- proto[rep(1, nrow(grid_2d)), , drop = FALSE]
newdata_full[[f1]] <- grid_2d[[f1]]
newdata_full[[f2]] <- grid_2d[[f2]]
grid_2d$pred_class <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "raw")
probs <- predict(rf_fit_model1_cv, newdata = newdata_full, type = "prob")
grid_2d$prob_fake <- probs[, "fake"]
grid_2d$pred_class <- factor(grid_2d$pred_class, levels = c("real","fake"))

# Plot
# Save to PDF for paper
#pdf("fig3_panel_G_H.pdf", width = 8, height = 5)

ggplot() +
  # Background and boundary 
  geom_raster(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], fill = pred_class),
    alpha = 0.45,
    show.legend = FALSE
  ) +
  geom_contour(
    data = grid_2d,
    aes(x = .data[[f1]], y = .data[[f2]], z = prob_fake),
    breaks = 0.5, colour = "black", linewidth = 0.85
  ) +

  geom_point(
    data = df_plot_test, 
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 2.8,
    colour = "black",
    stroke = 0.4
  ) +
  scale_fill_manual(
    name = "Test", 
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 2.8, alpha = 1)) 
  ) +
  scale_shape_manual(
    name = "Test", 
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 2.8, alpha = 1)) 
  ) +

  ggnewscale::new_scale_fill() +
  ggnewscale::new_scale("shape") +

  geom_point(
    data = df_plot_train, 
    aes(
      x = .data[[f1]],
      y = .data[[f2]],
      shape = truth,
      fill = truth
    ),
    size = 1.5,
    alpha = 0.4, 
    colour = "black",
    stroke = 0.2
  ) +
  scale_fill_manual(
    name = "Train", 
    values = c("real" = "#009E73", "fake" = "#E69F00"),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 1.5, alpha = 0.4)) 
  ) +
  scale_shape_manual(
    name = "Train", 
    values = c("real" = 21, "fake" = 24),
    labels = c("Real", "Fake"),
    guide = guide_legend(override.aes = list(size = 1.5, alpha = 0.4)) 
  ) +

  facet_wrap(~ Condition) + 
  
  coord_cartesian(
    xlim = c(-2, 2),
    ylim = c(-4, 2),
    expand = FALSE
  ) +
  labs(
    x = f1, y = f2,
    title = ""
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 12)
  )
#dev.off()
```

Emotion bar chart revised:

```{r, fig.width=3, fig.height=5}
# Save to PDF for paper
pdf("fig3_panel_A.pdf", width = 3, height = 5)
par(mar = c(4, 4, 3, 1))

ggplot(
  data = data.frame(
    Condition = factor(c("Emotion", "No Emotion"),
                       levels = c("Emotion", "No Emotion")),
    Accuracy  = c(72.4, 55.6),
    CI_low    = c(59.8, 39.6),
    CI_high   = c(82.2, 70.5),
    p_label   = c("p < .001", "p = .309"),
    vjust_val = c(-4, -6)
  ),
  aes(x = Condition, y = Accuracy, fill = Condition) 
) +
  geom_col(width = 0.6, color = "black") + 
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high),
                width = 0.2, size = 0.8, color = "black") + 
  geom_hline(yintercept = 50, linetype = "dashed",
             colour = "black", size = 0.6) + 
  annotate("text", x = 2.2, y = 50, label = "",
           hjust = -1.1, vjust = -0.5, size = 4, colour = "black") +
  
  geom_text(aes(label = p_label, vjust = vjust_val), size = 5) + 
  
  scale_fill_manual(
    values = c("Emotion" = "#009E73", "No Emotion" = "#009E73"),
    guide = "none" 
  ) +
  
  scale_y_continuous(limits = c(0, 100),
                     expand = expansion(mult = c(0, 0.08))) +
  labs(y = "Accuracy (%)", x = NULL) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x  = element_text(size = 12, color = "black", margin = margin(t=5)), 
    axis.text.y = element_text(size = 14, color = "black"), 
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.margin  = margin(10, 10, 10, 10)
  )
dev.off()
```

Updated PLD vs Full Stim bar chart:

```{r, fig.width=6, fig.height=6}
ggplot(
  data = data.frame(
    Stimuli = factor(
      c("Full Stimuli", "Point Light Display", "Point Light Display"),
      levels = c("Full Stimuli", "Point Light Display")
    ),
    Method = factor(
      c("Human Observers", "Random Forest", "Human Observers"),
      levels = c("Human Observers", "Random Forest")
    ),
    Accuracy = c(88.33, 55.00, 50.51),
    CI_low  = c(80.0,  39.8,  40.3),
    CI_high = c(93.5,  69.3,  60.7)
  ),
  aes(x = Stimuli, y = Accuracy, fill = Method)
) +
  geom_col(position = position_dodge(width = 0.65), width = 0.6, color = "black") + # <-- Added outline
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    position = position_dodge(width = 0.65),
    width = 0.2, size = 0.8, color = "black" # <-- Set color to black
  ) +
  geom_hline(yintercept = 50, linetype = "dashed",
             colour = "black", size = 0.6) + # <-- Changed to black
  annotate("text", x = 1.55, y = 50, label = "NIR (0.5)", # <-- Adjusted x-position
           hjust = -2.7, vjust = -0.5, size = 4, colour = "black") + # <-- Changed to black
  scale_y_continuous(limits = c(0, 100),
                     expand = expansion(mult = c(0, 0.08))) +
  
  # --- Use your established color palette ---
  scale_fill_manual(values = c("Human Observers" = "#1b9e77",
                               "Random Forest"   = "#d95f02")) +
  
  labs(x = NULL, y = "Accuracy (%)", fill = NULL) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, face = "bold", color = "black"),
    axis.text.y = element_text(size = 18, color = "black"),
    axis.text.x = element_text(size = 18, color = "black"),
    plot.margin  = margin(10, 10, 10, 10)
  )
```

Updated Emotion vs No Emotion confusion matrices:

```{r, fig.width=7, fig.height=5}
# Filter subsets
emotion_test <- dta_classifier_test_new_scaled %>% filter(emotion == "yes")
no_emotion_test <- dta_classifier_test_new_scaled %>% filter(emotion == "no")

# Model predictions
pred_emotion    <- predict(rf_fit_model1_cv, newdata = emotion_test)
pred_no_emotion <- predict(rf_fit_model1_cv, newdata = no_emotion_test)

# Raw confusion matrices
cm_emotion <- table(Predicted = pred_emotion, Actual = emotion_test$isfakeorreal)
cm_no_emotion <- table(Predicted = pred_no_emotion, Actual = no_emotion_test$isfakeorreal)

# Convert to tidy format
tidy_cm <- function(cm, title) {
  as.data.frame(cm) %>%
    mutate(Condition = title)
}

df_cm <- bind_rows(
  tidy_cm(cm_emotion,    "Emotion Videos"),
  tidy_cm(cm_no_emotion, "No Emotion Videos")
)

# calculate percentages PER FACET
# This calculates the % within each 2x2 grid
df_cm <- df_cm %>%
  group_by(Condition) %>%
  mutate(
    Total = sum(Freq),
    Percent = Freq / Total
  ) %>%
  ungroup()


p_cm <- df_cm %>%
  ggplot(aes(x = Actual, y = Predicted, fill = Freq)) +
  
  # black outlines
  geom_tile(color = "black", linewidth = 0.6) +
  
  # percentages
  geom_text(aes(label = sprintf("%d\n(%.0f%%)", Freq, 100*Percent)),
            size = 5, lineheight = 0.95, color = "black") +
            
  scale_fill_gradient(low = "white", high = "#009E73") +
  
  # Add coord_equal for square tiles
  coord_equal() +
  
  facet_wrap(~Condition, nrow = 1) +
  labs(title = "Confusion Matrices by Emotional Content",
       subtitle = "Performance on Test Set (n=94)",
       fill = "Count") +
  theme_classic(base_size = 14) +
  
  # Updated theme to match agreement matrices
  theme(
    strip.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    legend.position = "right",
    axis.title = element_text(face = "bold", color = "black"), 
    axis.text = element_text(face = "bold", color = "black")
  )

# Save to PDF for paper
pdf("fig3_panel_B.pdf", width=7, height=5)
par(mar = c(4, 4, 3, 1))
p_cm
dev.off()
```

Feature graphs:

```{r, fig.width=5, fig.height=5}
library(ggplot2)
library(dplyr)
library(tidyr)

boruta_selected_features <- c(
  "diff1x_pacf5_AU17_chin_raiser",
  "diff1x_pacf5_AU01_inner_brow_raiser",
  "diff2_acf1_AU01_inner_brow_raiser",
  "diff2_acf10_AU12_lip_corner_puller",
  "max_kl_shift_AU12_lip_corner_puller",
  "diff1_acf1_AU12_lip_corner_puller",
  "lumpiness_AU12_lip_corner_puller",
  "diff2_acf1_AU12_lip_corner_puller"
)

# pretty_names <- c(
#   "diff1x_pacf5_AU17_chin_raiser"       = "PACF(5) \u0394 AU17",
#   "diff1x_pacf5_AU01_inner_brow_raiser" = "PACF(5) \u0394 AU01",
#   "diff2_acf1_AU01_inner_brow_raiser"   = "ACF(1) \u0394\u00B2 AU01",
#   "diff2_acf10_AU12_lip_corner_puller"  = "ACF(10) \u0394\u00B2 AU12",
#   "max_kl_shift_AU12_lip_corner_puller" = "Max KL Shift AU12",
#   "diff1_acf1_AU12_lip_corner_puller"   = "ACF(1) \u0394 AU12",
#   "lumpiness_AU12_lip_corner_puller"    = "Lumpiness AU12",
#   "diff2_acf1_AU12_lip_corner_puller"   = "ACF(1) \u0394\u00B2 AU12"
# )

pretty_names <- c(
  "diff1x_pacf5_AU17_chin_raiser"       = "diff1x_pacf5_AU17",
  "diff1x_pacf5_AU01_inner_brow_raiser" = "diff1x_pacf5_AU01",
  "diff2_acf1_AU01_inner_brow_raiser"   = "diff2_acf1_AU01",
  "diff2_acf10_AU12_lip_corner_puller"  = "diff2_acf10_AU12",
  "max_kl_shift_AU12_lip_corner_puller" = "max_kl_shift_AU12",
  "diff1_acf1_AU12_lip_corner_puller"   = "diff1_acf1_AU12",
  "lumpiness_AU12_lip_corner_puller"    = "lumpiness_AU12",
  "diff2_acf1_AU12_lip_corner_puller"   = "diff2_acf1_AU12"
)


plot_df <- dta_classifier_train_new_scaled %>%
  select(isfakeorreal, all_of(boruta_selected_features)) %>%
  pivot_longer(
    cols = all_of(boruta_selected_features),
    names_to = "Feature",
    values_to = "Value"
  ) %>%
  mutate(
    Feature = factor(Feature,
                     levels = boruta_selected_features,
                     labels = pretty_names)
  )

fill_pal <- c("real" = "#009E73", "fake" = "#E69F00")


# Save to PDF for paper
pdf("fig5_panel_C.pdf", width = 5, height = 5)

ggplot(plot_df, aes(x = isfakeorreal, y = Value, fill = isfakeorreal)) +
  
  # Violin plot
  geom_violin(
    width = 0.9, 
    trim = TRUE, 
    alpha = 0.3,
    color = "black", 
    linewidth = 0.3
  ) +
  
  # Boxplot
  geom_boxplot(
    width = 0.25, 
    outlier.shape = 16, 
    outlier.size = 0.8,
    color = "black", 
    linewidth = 0.4, 
    alpha = 0.9
  ) +
  
  facet_wrap(
    ~ Feature, 
    ncol = 2, 
    nrow = 4, 
    scales = "free_y"
  ) +
  
  # Apply colors and labels
  scale_fill_manual(values = fill_pal, labels = c("Real", "Fake")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  labs(
    title = "Distribution of Final Selected Features",
    x = NULL,
    y = "Z-scored Value",
    fill = "Class"
  ) +
  
  theme_classic(base_size = 10) + 
  theme(
    legend.position = "top",
    legend.title = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 9),
    
    strip.text = element_text(size = 9, face = "bold", color = "black"),
    strip.background = element_rect(fill = "grey90", color = "grey50"),
    
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 10, face = "bold"),
    
    axis.text.x = element_text(size = 9, color = "black"),
    axis.text.y = element_text(size = 8, color = "black"),
    
    panel.border = element_rect(color = "grey80", fill = NA, linewidth = 0.5)
  )
dev.off()
```

```{r, fig.height=5, fig.width=5}
# --- 1. Final selected features (from your script) ---
boruta_selected_features <- c(
  "diff1x_pacf5_AU17_chin_raiser",
  "diff1x_pacf5_AU01_inner_brow_raiser",
  "diff2_acf1_AU01_inner_brow_raiser",
  "diff2_acf10_AU12_lip_corner_puller",
  "max_kl_shift_AU12_lip_corner_puller",
  "diff1_acf1_AU12_lip_corner_puller",
  "lumpiness_AU12_lip_corner_puller",
  "diff2_acf1_AU12_lip_corner_puller"
)

# --- 2. Pretty names for plot facets (using Unicode for Δ and ²) ---
pretty_names <- c(
  "diff1x_pacf5_AU17_chin_raiser"       = "PACF(5) \u0394 AU17",
  "diff1x_pacf5_AU01_inner_brow_raiser" = "PACF(5) \u0394 AU01",
  "diff2_acf1_AU01_inner_brow_raiser"   = "ACF(1) \u0394\u00B2 AU01",
  "diff2_acf10_AU12_lip_corner_puller"  = "ACF(10) \u0394\u00B2 AU12",
  "max_kl_shift_AU12_lip_corner_puller" = "Max KL Shift AU12",
  "diff1_acf1_AU12_lip_corner_puller"   = "ACF(1) \u0394 AU12",
  "lumpiness_AU12_lip_corner_puller"    = "Lumpiness AU12",
  "diff2_acf1_AU12_lip_corner_puller"   = "ACF(1) \u0394\u00B2 AU12"
)

# --- 3. Prepare data for plotting ---
plot_df <- dta_classifier_train_new_scaled %>%
  select(isfakeorreal, all_of(boruta_selected_features)) %>%
  pivot_longer(
    cols = all_of(boruta_selected_features),
    names_to = "Feature",
    values_to = "Value"
  ) %>%
  # Apply the pretty names and set factor order
  mutate(
    Feature = factor(Feature,
                     levels = boruta_selected_features,
                     labels = pretty_names),
    # Ensure consistent factor levels for isfakeorreal
    isfakeorreal = factor(isfakeorreal, levels = c("real", "fake"))
  )
plot_df <- plot_df %>%
  group_by(Feature) %>%
  filter(
    Value >= quantile(Value, 0.005, na.rm = TRUE),
    Value <= quantile(Value, 0.995, na.rm = TRUE)
  ) %>%
  ungroup()

# --- 4. Define color palette ---
fill_pal <- c("real" = "#1b9e77", "fake" = "#d95f02")

# --- 5. Create the 2x4 plot ---
p <- ggplot(plot_df, aes(x = isfakeorreal, y = Value, fill = isfakeorreal)) +
  
  # Violin plot
  geom_violin(
    width = 0.8, 
    trim = TRUE, 
    alpha = 0.3,
    color = "black", 
    linewidth = 0.3
  ) +
  
  # Boxplot
  geom_boxplot(
    width = 0.25, 
    outlier.shape = 16, 
    outlier.size = 0.5,
    outlier.alpha = 0.5,
    color = "black", 
    linewidth = 0.35, 
    alpha = 0.8
  ) +
  
  # Facet into 2 columns and 4 rows
  facet_wrap(
    ~ Feature, 
    ncol = 2, 
    nrow = 4, 
    scales = "free_y"
  ) +
  
  # Apply colors and labels
  scale_fill_manual(
    values = fill_pal, 
    labels = c("real" = "Real", "fake" = "Fake")
  ) +
  scale_x_discrete(
    labels = c("real" = "Real", "fake" = "Fake")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  
  labs(
    title = "Distribution of Final Selected Features",
    x = NULL,
    y = "Standardised Value (z-score)",
    fill = "Video Class"
  ) +
  
  # --- 6. Theme optimized for publication ---
  theme_classic(base_size = 8) +
  theme(
    # Font family
    text = element_text(family = "Arial"),
    
    # Legend
    legend.position = "top",
    legend.title = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.4, "cm"),
    legend.margin = margin(0, 0, 3, 0),
    
    # Facet titles (the feature names)
    strip.text = element_text(size = 8, face = "bold", color = "black"),
    strip.background = element_rect(fill = "grey90", color = "grey50", linewidth = 0.3),
    
    # Title and axis labels
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 8, face = "bold"),
    axis.title.x = element_blank(),
    
    # Axis text
    axis.text.x = element_text(size = 7, color = "black"),
    axis.text.y = element_text(size = 7, color = "black"),
    
    # Panel
    panel.border = element_rect(color = "grey70", fill = NA, linewidth = 0.3),
    panel.spacing = unit(0.3, "cm"),
    
    # Lines
    axis.line = element_line(linewidth = 0.3),
    axis.ticks = element_line(linewidth = 0.3),
    
    # Margins
    plot.margin = margin(5, 5, 5, 5)
  )

print(p)


ggsave(
  "feature_distributions.pdf",
  plot = p,
  width = 5,
  height = 5,
  device = cairo_pdf
)

```

AU Before and After Smoothing By Real and Fake:

```{r, fig.height=5, fig.width=10}
dta_agg_long <- dta_preprocessed_smth %>%
  # Make sure to select the new variable
  select(frame, AU17_r, smth_AU17_r, isfakeorreal) %>% 
  tidyr::pivot_longer(
    cols = c(AU17_r, smth_AU17_r),
    names_to = "Type",
    values_to = "Intensity"
  ) %>%
  group_by(isfakeorreal, frame, Type) %>%
  # mean intensity for real and fake
  summarise(
    Intensity = mean(Intensity, na.rm = TRUE),
    .groups = "drop"
  )

base_theme <- theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 18, face = "bold", color = "black"),
    axis.title.x = element_text(size = 18, color = "black"),
    axis.text.y  = element_text(size = 16, color = "black"),
    axis.text.x  = element_text(size = 16, color = "black"),
    plot.title   = element_text(size = 18, face = "bold", color = "black", hjust = 0.5),
    plot.margin  = margin(10, 10, 10, 10),
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(size = 16, face = "bold", color = "black")
  )

# Calculate common y axis range
y_limits <- range(dta_agg_long$Intensity, na.rm = T)

# plot 1
p_raw <- dta_agg_long %>%
  filter(Type == "AU17_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#1b9e77", alpha = 0.6, linewidth = 0.8) +
  facet_wrap(~ isfakeorreal) +
  labs(
    title = "Raw",
    x = NULL,
    y = "AU Intensity"
  ) +
  scale_y_continuous(limits = y_limits, expand = expansion(mult = c(0, 0.08))) +
  base_theme

# plot 2
p_smth <- dta_agg_long %>%
  filter(Type == "smth_AU17_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#d95f02", linewidth = 0.8) +
  facet_wrap(~ isfakeorreal) +
  labs(
    title = "Smoothed",
    x = "Frame",
    y = "AU Intensity"
  ) +
  scale_y_continuous(limits = y_limits, expand = expansion(mult = c(0, 0.08))) +
  base_theme

# combine plots
p_raw / p_smth +
  plot_annotation(
    title = "AU17 - Before vs. After Smoothing", 
    theme = theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5, color = "black")
    )
  )
```

For two videos only:

```{r}
REAL_VIDEO_FILENAME <- "21__kitchen_still" 
FAKE_VIDEO_FILENAME <- "21_07__kitchen_still__MKU99DVX" 

dta_agg_long <- dta_preprocessed_smth %>%
  select(video_filename, frame, AU17_r, smth_AU17_r, isfakeorreal) %>% 
  filter(video_filename %in% c(REAL_VIDEO_FILENAME, FAKE_VIDEO_FILENAME)) %>%
  tidyr::pivot_longer(
    cols = c(AU17_r, smth_AU17_r),
    names_to = "Type",
    values_to = "Intensity"
  )

base_theme <- theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 18, face = "bold", color = "black"),
    axis.title.x = element_text(size = 18, color = "black"),
    axis.text.y  = element_text(size = 16, color = "black"),
    axis.text.x  = element_text(size = 16, color = "black"),
    plot.title   = element_text(size = 18, face = "bold", color = "black", hjust = 0.5),
    plot.margin  = margin(10, 10, 10, 10),
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(size = 16, face = "bold", color = "black")
  )

y_limits <- range(dta_agg_long$Intensity, na.rm = T)

# plot 1
p_raw <- dta_agg_long %>%
  filter(Type == "AU17_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#1b9e77", alpha = 0.6, linewidth = 0.8) +
  facet_wrap(~ isfakeorreal) +
  labs(
    title = "Raw",
    x = NULL,
    y = "AU Intensity"
  ) +
  scale_y_continuous(limits = y_limits, expand = expansion(mult = c(0, 0.08))) +
  base_theme

# plot 2
p_smth <- dta_agg_long %>%
  filter(Type == "smth_AU17_r") %>%
  ggplot(aes(x = frame, y = Intensity)) +
  geom_line(color = "#d95f02", linewidth = 0.8) +
  facet_wrap(~ isfakeorreal) +
  labs(
    title = "Smoothed",
    x = "Frame",
    y = "AU Intensity"
  ) +
  scale_y_continuous(limits = y_limits, expand = expansion(mult = c(0, 0.08))) +
  base_theme

# combine plots
p_raw / p_smth +
  plot_annotation(
    title = "AU17 - Before vs. After Smoothing", 
    subtitle = "21_07__kitchen_still__MKU99DVX + 21__kitchen_still",
    theme = theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5, color = "black")
    )
  )
```

Feature bar charts:

```{r, fig.width=5, fig.height=5}
REAL_VIDEO_FILENAME <- "21__kitchen_still" 
FAKE_VIDEO_FILENAME <- "21_07__kitchen_still__MKU99DVX" 

au17_feature <- "diff1x_pacf5_AU17_chin_raiser"

dta_feature_compare <- dta_cmfts_output_imputed_zerosd %>%
  filter(video_filename %in% c(REAL_VIDEO_FILENAME, FAKE_VIDEO_FILENAME)) %>%
  select(video_filename, isfakeorreal, all_of(au17_feature)) %>%
  rename(Value = all_of(au17_feature))


ggplot(dta_feature_compare, aes(x = isfakeorreal, y = Value, fill = isfakeorreal)) +
  
  # Use geom_col() for a bar chart
  geom_col(width = 0.6, color = "black") +
  
  # Add text labels to show the exact value
  geom_text(aes(label = round(Value, 2)), vjust = -0.5, size = 5) +
  
  # Use the same colors as your other plots
  scale_fill_manual(
    values = c("real" = "#1b9e77", "fake" = "#d95f02"),
    guide = "none" # Hide the legend, as the x-axis is clear
  ) +
  
  labs(
    title = au17_feature,
    x = NULL,
    y = "Feature Value (Z-scored)"
  ) +
  
  # Use a theme similar to your others
  theme_classic(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(size = 18, face = "bold"),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_text(size = 16)
  )
```

# Classifiers w/ Balanced Emotion

## Random Forest

```{r}
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

rf_fit_model1_balanced_cv <- caret::train(
  isfakeorreal ~ ., # nmf
  data = dta_classifier_train_new_balanced_scaled[, -(1:2)],
  method = "rf",
  metric = "ROC",
  trControl = classifier_control,
  tuneLength = 5
)

print(rf_fit_model1_balanced_cv)
plot(caret::varImp(rf_fit_model1_balanced_cv)) # To see feature importance
```

Evaluate on test data:

```{r}
rf_model_1_balanced_cv_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model1_balanced_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_balanced_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(rf_fit_model1_balanced_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

```         
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   33   16
      real   14   31
                                          
               Accuracy : 0.6809          
                 95% CI : (0.5767, 0.7733)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.000294        
                                          
                  Kappa : 0.3617          
                                          
 Mcnemar's Test P-Value : 0.855132        
                                          
            Sensitivity : 0.7021          
            Specificity : 0.6596          
         Pos Pred Value : 0.6735          
         Neg Pred Value : 0.6889          
             Prevalence : 0.5000          
         Detection Rate : 0.3511          
   Detection Prevalence : 0.5213          
      Balanced Accuracy : 0.6809          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
Area under the curve: 0.6815
95% CI: 0.5707-0.7923 (DeLong)
```

### Sensitivity Analysis - Emotion Subsets with Balanced Model

```{r}
cat("--- EMOTION VIDEOS (n=58) ---\n")

# Get predictions on emotion subset
rf_balanced_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_balanced_cv, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)

print(rf_balanced_emotion)

# Calculate ROC-AUC for emotion subset
y_test_emotion <- factor(dta_classifier_test_new_emotion$isfakeorreal, levels = c("real", "fake"))
p_fake_emotion <- predict(rf_fit_model1_balanced_cv, 
                          newdata = dta_classifier_test_new_emotion, 
                          type = "prob")[, "fake"]

roc_obj_emotion <- pROC::roc(y_test_emotion, p_fake_emotion, levels = c("real", "fake"))
auc_emotion <- pROC::auc(roc_obj_emotion)
ci_emotion <- pROC::ci.auc(roc_obj_emotion)

cat(sprintf("\nROC-AUC: %.4f\n", auc_emotion))
cat(sprintf("95%% CI: %.4f - %.4f\n\n", ci_emotion[1], ci_emotion[3]))
```

No emotion:

```{r}
cat("--- NO EMOTION VIDEOS (n=36) ---\n")

# Get predictions on no emotion subset
rf_balanced_no_emotion <- caret::confusionMatrix(
  data = predict(rf_fit_model1_balanced_cv, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)

print(rf_balanced_no_emotion)

# Calculate ROC-AUC for no emotion subset
y_test_no_emotion <- factor(dta_classifier_test_new_no_emotion$isfakeorreal, levels = c("real", "fake"))
p_fake_no_emotion <- predict(rf_fit_model1_balanced_cv, 
                             newdata = dta_classifier_test_new_no_emotion, 
                             type = "prob")[, "fake"]

roc_obj_no_emotion <- pROC::roc(y_test_no_emotion, p_fake_no_emotion, levels = c("real", "fake"))
auc_no_emotion <- pROC::auc(roc_obj_no_emotion)
ci_no_emotion <- pROC::ci.auc(roc_obj_no_emotion)

cat(sprintf("\nROC-AUC: %.4f\n", auc_no_emotion))
cat(sprintf("95%% CI: %.4f - %.4f\n\n", ci_no_emotion[1], ci_no_emotion[3]))
```

Comparison:

```{r}
cat("ORIGINAL MODEL (trained on imbalanced data: 222 emotion, 148 no emotion):\n")
cat(sprintf("  Emotion videos (n=58):    Accuracy = %.1f%%, Sensitivity = %.1f%%, Specificity = %.1f%%\n",
            rf_model_1_emotion$overall["Accuracy"] * 100,
            rf_model_1_emotion$byClass["Sensitivity"] * 100,
            rf_model_1_emotion$byClass["Specificity"] * 100))
cat(sprintf("  No Emotion videos (n=36): Accuracy = %.1f%%, Sensitivity = %.1f%%, Specificity = %.1f%%\n",
            rf_model_1_no_emotion$overall["Accuracy"] * 100,
            rf_model_1_no_emotion$byClass["Sensitivity"] * 100,
            rf_model_1_no_emotion$byClass["Specificity"] * 100))
cat(sprintf("  >> Performance gap: %.1f%% accuracy difference\n\n",
            (rf_model_1_emotion$overall["Accuracy"] - rf_model_1_no_emotion$overall["Accuracy"]) * 100))

cat("BALANCED MODEL (trained on balanced data: 222 emotion, 222 no emotion):\n")
cat(sprintf("  Emotion videos (n=58):    Accuracy = %.1f%%, Sensitivity = %.1f%%, Specificity = %.1f%%\n",
            rf_balanced_emotion$overall["Accuracy"] * 100,
            rf_balanced_emotion$byClass["Sensitivity"] * 100,
            rf_balanced_emotion$byClass["Specificity"] * 100))
cat(sprintf("  No Emotion videos (n=36): Accuracy = %.1f%%, Sensitivity = %.1f%%, Specificity = %.1f%%\n",
            rf_balanced_no_emotion$overall["Accuracy"] * 100,
            rf_balanced_no_emotion$byClass["Sensitivity"] * 100,
            rf_balanced_no_emotion$byClass["Specificity"] * 100))
cat(sprintf("  >> Performance gap: %.1f%% accuracy difference\n\n",
            (rf_balanced_emotion$overall["Accuracy"] - rf_balanced_no_emotion$overall["Accuracy"]) * 100))


```

## C5.0B

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

c50_fit_model1_balanced_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_balanced_scaled[, -(1:2)],
  method = "C5.0",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(
    model = "tree",
    trials = c(1, 10, 20, 30),  # number of boosting iterations
    winnow = FALSE
  )
)

print(c50_fit_model1_balanced_cv)
plot(caret::varImp(c50_fit_model1_balanced_cv)) # To see feature importanc
```

Evaluate on test data:

```{r}
c50_model_1_balanced_cv_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model1_balanced_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(c50_model_1_balanced_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(c50_fit_model1_balanced_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

```         
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   30   23
      real   17   24
                                          
               Accuracy : 0.5745          
                 95% CI : (0.4682, 0.6759)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.08983         
                                          
                  Kappa : 0.1489          
                                          
 Mcnemar's Test P-Value : 0.42920         
                                          
            Sensitivity : 0.6383          
            Specificity : 0.5106          
         Pos Pred Value : 0.5660          
         Neg Pred Value : 0.5854          
             Prevalence : 0.5000          
         Detection Rate : 0.3191          
   Detection Prevalence : 0.5638          
      Balanced Accuracy : 0.5745          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
Area under the curve: 0.56
95% CI: 0.4423-0.6776 (DeLong)
```

## SVM

```{r}
# Cross-valiation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

svm_model1_balanced_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_balanced_scaled[, -(1:2)],
  method = "svmRadial",  # RBF kernel
  trControl = classifier_control
)
```

Evaluate on test data:

```{r}
svm_model_1_balanced_cv_performance <- caret::confusionMatrix(
  data = predict(svm_model1_balanced_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(svm_model_1_balanced_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(svm_model1_balanced_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

```         
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   30   14
      real   17   33
                                          
               Accuracy : 0.6702          
                 95% CI : (0.5656, 0.7638)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.0006307       
                                          
                  Kappa : 0.3404          
                                          
 Mcnemar's Test P-Value : 0.7194375       
                                          
            Sensitivity : 0.6383          
            Specificity : 0.7021          
         Pos Pred Value : 0.6818          
         Neg Pred Value : 0.6600          
             Prevalence : 0.5000          
         Detection Rate : 0.3191          
   Detection Prevalence : 0.4681          
      Balanced Accuracy : 0.6702          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
Area under the curve: 0.713
95% CI: 0.608-0.818 (DeLong)
```

## Logistic Regression

```{r}
# Cross-valiation setup
classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

lr_model1_balanced_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_balanced_scaled[, -(1:2)],
  method = "glm",  # Generalised linear models
  family = "binomial",
  trControl = classifier_control
)

plot(caret::varImp(lr_model1_balanced_cv)) # To see feature importance
```

Evaluate on test data:

```{r}
lr_model1_balanced_cv_performance <- caret::confusionMatrix(
  data = predict(lr_model1_balanced_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)

print(lr_model1_balanced_cv_performance)


# ROC AUC
y_test  <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real","fake"))
p_fake  <- predict(lr_model1_balanced_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[,"fake"]

roc_obj <- pROC::roc(y_test, p_fake, levels = c("real","fake"))
pROC::auc(roc_obj)       # ROC-AUC
pROC::ci.auc(roc_obj)    # 95% CI (DeLong)
```

```         
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   22   21
      real   25   26
                                          
               Accuracy : 0.5106          
                 95% CI : (0.4054, 0.6152)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.4590          
                                          
                  Kappa : 0.0213          
                                          
 Mcnemar's Test P-Value : 0.6583          
                                          
            Sensitivity : 0.4681          
            Specificity : 0.5532          
         Pos Pred Value : 0.5116          
         Neg Pred Value : 0.5098          
             Prevalence : 0.5000          
         Detection Rate : 0.2340          
   Detection Prevalence : 0.4574          
      Balanced Accuracy : 0.5106          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
Area under the curve: 0.5405
95% CI: 0.4226-0.6584 (DeLong)
```

# Classifiers w/ Balanced Emotion Classes (Downsampling)

Previously I balanced the emotion class using bootstrap oversampling. I did this to prevent discarding many rows of data, however, this oversampling method works by adding duplicate rows to upsampling the minority class. A reviewer could argue that the classifier trained on this balanced train set is not actually learning anything different as the upsampled rows are the same. Therefore, I will run this again but this time using downsampling. This will discard data but will allow a fair comparison. Classes should be n=148 each after downsampling.

```{r}
set.seed(12345)
# Separate by emotion class
emotion_majority <- dta_cmfts_output_imputed_zerosd %>%
  filter(emotion == "yes")  # n = 222

emotion_minority <- dta_cmfts_output_imputed_zerosd %>%
  filter(emotion == "no")   # n = 148

cat(sprintf("Before downsampling:\n  Emotion: %d\n  No Emotion: %d\n\n",
            nrow(emotion_majority), nrow(emotion_minority)))

# Downsample majority (Emotion) to match minority (No Emotion)
emotion_majority_downsampled <- emotion_majority %>%
  sample_n(size = nrow(emotion_minority), replace = FALSE)

# Combine
dta_balanced_downsampled <- bind_rows(emotion_majority_downsampled, emotion_minority)

cat(sprintf("After downsampling:\n  Emotion: %d\n  No Emotion: %d\n  Total: %d\n\n",
            sum(dta_balanced_downsampled$emotion == "yes"),
            sum(dta_balanced_downsampled$emotion == "no"),
            nrow(dta_balanced_downsampled)))
```

```{r}
dta_train_ds <- dta_balanced_downsampled %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

# scale using same params as origina model
feature_columns <- names(dta_train_ds)[!names(dta_train_ds) %in% 
                                         c("video_filename", "emotion", "isfakeorreal")]

dta_train_ds_scaled <- dta_train_ds
dta_train_ds_scaled[, feature_columns] <- predict(scaling_model,
                                                    newdata = dta_train_ds[, feature_columns])

# Set factor levels
dta_train_ds_scaled$isfakeorreal <- factor(
  dta_train_ds_scaled$isfakeorreal,
  levels = c("fake", "real")
)

# train rf model

classifier_control_ds <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

rf_downsampled_cv <- caret::train(
  isfakeorreal ~ .,
  data = dta_train_ds_scaled[, -(1:2)],  # exclude video_filename and emotion
  method = "rf",
  metric = "ROC",
  trControl = classifier_control_ds,
  tuneLength = 5
)

cat("--- DOWNSAMPLED RF MODEL ---\n")
print(rf_downsampled_cv)
```

Evaluate on test data:

```{r}
library(pROC)
library(caret)

cat("\n--- FULL TEST SET (n=94) ---\n")
rf_ds_full <- confusionMatrix(
  data = predict(rf_downsampled_cv, newdata = dta_classifier_test_new_scaled),
  reference = dta_classifier_test_new_scaled$isfakeorreal,
  positive = "fake"
)
print(rf_ds_full)

y_full <- factor(dta_classifier_test_new_scaled$isfakeorreal, levels = c("real", "fake"))
p_full <- predict(rf_downsampled_cv, newdata = dta_classifier_test_new_scaled, type = "prob")[, "fake"]
roc_full <- pROC::roc(y_full, p_full, levels = c("real", "fake"))
cat(sprintf("ROC-AUC: %.4f [%.4f, %.4f]\n\n",
            pROC::auc(roc_full), ci.auc(roc_full)[1], ci.auc(roc_full)[3]))


cat("--- EMOTION VIDEOS (n=58) ---\n")
rf_ds_emotion <- confusionMatrix(
  data = predict(rf_downsampled_cv, newdata = dta_classifier_test_new_emotion),
  reference = dta_classifier_test_new_emotion$isfakeorreal,
  positive = "fake"
)
print(rf_ds_emotion)

y_em <- factor(dta_classifier_test_new_emotion$isfakeorreal, levels = c("real", "fake"))
p_em <- predict(rf_downsampled_cv, newdata = dta_classifier_test_new_emotion, type = "prob")[, "fake"]
roc_em <- pROC::roc(y_em, p_em, levels = c("real", "fake"))
auc_em <- pROC::auc(roc_em)
ci_em <- ci.auc(roc_em)
cat(sprintf("ROC-AUC: %.4f [%.4f, %.4f]\n\n", auc_em, ci_em[1], ci_em[3]))


cat("--- NO EMOTION VIDEOS (n=36) ---\n")
rf_ds_no_emotion <- confusionMatrix(
  data = predict(rf_downsampled_cv, newdata = dta_classifier_test_new_no_emotion),
  reference = dta_classifier_test_new_no_emotion$isfakeorreal,
  positive = "fake"
)
print(rf_ds_no_emotion)

y_no <- factor(dta_classifier_test_new_no_emotion$isfakeorreal, levels = c("real", "fake"))
p_no <- predict(rf_downsampled_cv, newdata = dta_classifier_test_new_no_emotion, type = "prob")[, "fake"]
roc_no <- roc(y_no, p_no, levels = c("real", "fake"))
auc_no <- auc(roc_no)
ci_no <- ci.auc(roc_no)
cat(sprintf("ROC-AUC: %.4f [%.4f, %.4f]\n\n", auc_no, ci_no[1], ci_no[3]))
```

```         
--- FULL TEST SET (n=94) ---
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   35   22
      real   12   25
                                          
               Accuracy : 0.6383          
                 95% CI : (0.5327, 0.7349)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.004774        
                                          
                  Kappa : 0.2766          
                                          
 Mcnemar's Test P-Value : 0.122713        
                                          
            Sensitivity : 0.7447          
            Specificity : 0.5319          
         Pos Pred Value : 0.6140          
         Neg Pred Value : 0.6757          
             Prevalence : 0.5000          
         Detection Rate : 0.3723          
   Detection Prevalence : 0.6064          
      Balanced Accuracy : 0.6383          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
ROC-AUC: 0.6573 [0.5451, 0.7695]

--- EMOTION VIDEOS (n=58) ---
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   19    9
      real   10   20
                                          
               Accuracy : 0.6724          
                 95% CI : (0.5366, 0.7899)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.005964        
                                          
                  Kappa : 0.3448          
                                          
 Mcnemar's Test P-Value : 1.000000        
                                          
            Sensitivity : 0.6552          
            Specificity : 0.6897          
         Pos Pred Value : 0.6786          
         Neg Pred Value : 0.6667          
             Prevalence : 0.5000          
         Detection Rate : 0.3276          
   Detection Prevalence : 0.4828          
      Balanced Accuracy : 0.6724          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
ROC-AUC: 0.7265 [0.5964, 0.8566]

--- NO EMOTION VIDEOS (n=36) ---
Confusion Matrix and Statistics

          Reference
Prediction fake real
      fake   16   13
      real    2    5
                                          
               Accuracy : 0.5833          
                 95% CI : (0.4076, 0.7449)
    No Information Rate : 0.5             
    P-Value [Acc > NIR] : 0.202516        
                                          
                  Kappa : 0.1667          
                                          
 Mcnemar's Test P-Value : 0.009823        
                                          
            Sensitivity : 0.8889          
            Specificity : 0.2778          
         Pos Pred Value : 0.5517          
         Neg Pred Value : 0.7143          
             Prevalence : 0.5000          
         Detection Rate : 0.4444          
   Detection Prevalence : 0.8056          
      Balanced Accuracy : 0.5833          
                                          
       'Positive' Class : fake            
                                          
Setting direction: controls < cases
ROC-AUC: 0.5463 [0.3434, 0.7491]
```

Comparison:

```{r}

cat("================================================================\n")
cat("COMPARISON: ORIGINAL vs DOWNSAMPLED MODEL\n")
cat("================================================================\n\n")

cat("ORIGINAL MODEL (trained on: 222 Emotion, 148 No Emotion = 370 total):\n")
cat(sprintf("  Emotion (n=58):    Acc = %.1f%%, Sens = %.1f%%, Spec = %.1f%%, AUC = 0.765\n",
            rf_model_1_emotion$overall["Accuracy"] * 100,
            rf_model_1_emotion$byClass["Sensitivity"] * 100,
            rf_model_1_emotion$byClass["Specificity"] * 100))
cat(sprintf("  No Emotion (n=36): Acc = %.1f%%, Sens = %.1f%%, Spec = %.1f%%, AUC = 0.605\n",
            rf_model_1_no_emotion$overall["Accuracy"] * 100,
            rf_model_1_no_emotion$byClass["Sensitivity"] * 100,
            rf_model_1_no_emotion$byClass["Specificity"] * 100))
cat(sprintf("  >> Gap: %.1f pp\n\n",
            (rf_model_1_emotion$overall["Accuracy"] - 
               rf_model_1_no_emotion$overall["Accuracy"]) * 100))

cat(sprintf("DOWNSAMPLED MODEL (trained on: 148 Emotion, 148 No Emotion = 296 total):\n"))
cat(sprintf("  Emotion (n=58):    Acc = %.1f%%, Sens = %.1f%%, Spec = %.1f%%, AUC = %.3f\n",
            rf_ds_emotion$overall["Accuracy"] * 100,
            rf_ds_emotion$byClass["Sensitivity"] * 100,
            rf_ds_emotion$byClass["Specificity"] * 100,
            auc_em))
cat(sprintf("  No Emotion (n=36): Acc = %.1f%%, Sens = %.1f%%, Spec = %.1f%%, AUC = %.3f\n",
            rf_ds_no_emotion$overall["Accuracy"] * 100,
            rf_ds_no_emotion$byClass["Sensitivity"] * 100,
            rf_ds_no_emotion$byClass["Specificity"] * 100,
            auc_no))
cat(sprintf("  >> Gap: %.1f pp\n",
            (rf_ds_emotion$overall["Accuracy"] - 
               rf_ds_no_emotion$overall["Accuracy"]) * 100))
```

Pattern holds. The gap is less (16.9 to 8.9 pp) but key finding is preserved (emotion videos are significantly above chance (67.2%, p = .006) whereas no emotion videos remain non-significant (58.3%, p =.203)).

# Emotion Classification Model

To test whether the emotional signal itself is weaker in fakes.

```{r}
# Filter training data to include ONLY real videos
dta_emotion_train_real_only <- dta_classifier_train_new_scaled %>%
  filter(isfakeorreal == "real")

cat(sprintf("Original training set size: %d videos\n", nrow(dta_classifier_train_new_scaled)))
cat(sprintf("Real-only training set size: %d videos\n", nrow(dta_emotion_train_real_only)))

# Check class distribution
emotion_distribution <- table(dta_emotion_train_real_only$emotion)
cat("\nEmotion distribution in real-only training data:\n")
print(emotion_distribution)
```

Balance emotion class

```{r}
# Determine majority and minority classes
if (emotion_distribution["yes"] > emotion_distribution["no"]) {
  majority_class <- "yes"
  minority_class <- "no"
  n_majority <- emotion_distribution["yes"]
  n_minority <- emotion_distribution["no"]
} else {
  majority_class <- "no"
  minority_class <- "yes"
  n_majority <- emotion_distribution["no"]
  n_minority <- emotion_distribution["yes"]
}

cat(sprintf("\nMajority class: %s (n=%d)\n", majority_class, n_majority))
cat(sprintf("Minority class: %s (n=%d)\n", minority_class, n_minority))

# Option 1: Undersample majority class (recommended - no data duplication)
set.seed(12345)
majority_data <- dta_emotion_train_real_only %>%
  filter(emotion == majority_class)
minority_data <- dta_emotion_train_real_only %>%
  filter(emotion == minority_class)

# Randomly sample from majority to match minority size
majority_data_downsampled <- majority_data %>%
  sample_n(size = n_minority, replace = FALSE)

# Combine to create balanced dataset
dta_emotion_train_balanced <- bind_rows(majority_data_downsampled, minority_data) %>%
  arrange(video_filename)  # Shuffle is done in train() function

cat(sprintf("\nBalanced training set size: %d videos\n", nrow(dta_emotion_train_balanced)))
cat("\nBalanced emotion distribution:\n")
print(table(dta_emotion_train_balanced$emotion))

# Convert emotion to factor with explicit levels
dta_emotion_train_balanced$emotion <- factor(
  dta_emotion_train_balanced$emotion,
  levels = c("no", "yes")
)
```

Train classifier:

```{r}
emotion_classifier_control <- caret::trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

# Train Random Forest to predict EMOTION (not isfakeorreal)
set.seed(12345)

cat("\nTraining Random Forest to classify emotion (yes/no) using only real videos...\n")

rf_emotion_classifier <- caret::train(
  emotion ~ .,  # Predict emotion instead of isfakeorreal
  data = dta_emotion_train_balanced[, !names(dta_emotion_train_balanced) %in% 
                                      c("video_filename", "isfakeorreal")],
  method = "rf",
  metric = "ROC",
  trControl = emotion_classifier_control,
  tuneLength = 5
)

print(rf_emotion_classifier)

# Plot feature importance
cat("\nFeature Importance for Emotion Classification:\n")
plot(caret::varImp(rf_emotion_classifier))
```

Evaluate on REAL test data:

```{r}
dta_test_real_only <- dta_classifier_test_new_scaled %>%
  filter(isfakeorreal == "real")

cat(sprintf("Number of real videos in test set: %d\n", nrow(dta_test_real_only)))

# Ensure emotion is a factor with correct levels
dta_test_real_only$emotion <- factor(
  dta_test_real_only$emotion,
  levels = c("no", "yes")
)

# Get predictions
emotion_pred_real <- caret::confusionMatrix(
  data = predict(rf_emotion_classifier, newdata = dta_test_real_only),
  reference = dta_test_real_only$emotion,
  positive = "yes"
)

cat("\nPerformance on REAL videos:\n")
print(emotion_pred_real)

# Calculate ROC-AUC for real videos
y_test_real <- factor(dta_test_real_only$emotion, levels = c("no", "yes"))
p_emotion_real <- predict(rf_emotion_classifier, 
                          newdata = dta_test_real_only, 
                          type = "prob")[, "yes"]

roc_obj_real <- pROC::roc(y_test_real, p_emotion_real, levels = c("no", "yes"))
auc_real <- pROC::auc(roc_obj_real)
ci_real <- pROC::ci.auc(roc_obj_real)

cat(sprintf("\nROC-AUC on REAL videos: %.4f [%.4f, %.4f]\n", 
            auc_real, ci_real[1], ci_real[3]))
```

Evaluate on FAKE test data:

```{r}
cat("\n\n--- Step 5: Testing on FAKE Videos ---\n")

# Filter test set to fake videos only
dta_test_fake_only <- dta_classifier_test_new_scaled %>%
  filter(isfakeorreal == "fake")

cat(sprintf("Number of fake videos in test set: %d\n", nrow(dta_test_fake_only)))

# Ensure emotion is a factor with correct levels
dta_test_fake_only$emotion <- factor(
  dta_test_fake_only$emotion,
  levels = c("no", "yes")
)

# Get predictions
emotion_pred_fake <- caret::confusionMatrix(
  data = predict(rf_emotion_classifier, newdata = dta_test_fake_only),
  reference = dta_test_fake_only$emotion,
  positive = "yes"
)

cat("\nPerformance on FAKE videos:\n")
print(emotion_pred_fake)

# Calculate ROC-AUC for fake videos
y_test_fake <- factor(dta_test_fake_only$emotion, levels = c("no", "yes"))
p_emotion_fake <- predict(rf_emotion_classifier, 
                          newdata = dta_test_fake_only, 
                          type = "prob")[, "yes"]

roc_obj_fake <- pROC::roc(y_test_fake, p_emotion_fake, levels = c("no", "yes"))
auc_fake <- pROC::auc(roc_obj_fake)
ci_fake <- pROC::ci.auc(roc_obj_fake)

cat(sprintf("\nROC-AUC on FAKE videos: %.4f [%.4f, %.4f]\n", 
            auc_fake, ci_fake[1], ci_fake[3]))
```

Comparison:

```{r}
cat("\n\n=== COMPARISON: REAL vs FAKE VIDEOS ===\n\n")

cat("Emotion Classifier Performance:\n")
cat("(Trained on REAL videos only, tested on both REAL and FAKE)\n\n")

cat(sprintf("REAL videos (n=%d):\n", nrow(dta_test_real_only)))
cat(sprintf("  Accuracy:    %.1f%%\n", emotion_pred_real$overall["Accuracy"] * 100))
cat(sprintf("  Sensitivity: %.1f%% (correctly identifies emotion='yes')\n", 
            emotion_pred_real$byClass["Sensitivity"] * 100))
cat(sprintf("  Specificity: %.1f%% (correctly identifies emotion='no')\n", 
            emotion_pred_real$byClass["Specificity"] * 100))
cat(sprintf("  ROC-AUC:     %.4f [%.4f, %.4f]\n", auc_real, ci_real[1], ci_real[3]))

cat(sprintf("\nFAKE videos (n=%d):\n", nrow(dta_test_fake_only)))
cat(sprintf("  Accuracy:    %.1f%%\n", emotion_pred_fake$overall["Accuracy"] * 100))
cat(sprintf("  Sensitivity: %.1f%% (correctly identifies emotion='yes')\n", 
            emotion_pred_fake$byClass["Sensitivity"] * 100))
cat(sprintf("  Specificity: %.1f%% (correctly identifies emotion='no')\n", 
            emotion_pred_fake$byClass["Specificity"] * 100))
cat(sprintf("  ROC-AUC:     %.4f [%.4f, %.4f]\n", auc_fake, ci_fake[1], ci_fake[3]))

# Calculate performance drop
accuracy_drop <- (emotion_pred_real$overall["Accuracy"] - 
                  emotion_pred_fake$overall["Accuracy"]) * 100
auc_drop <- auc_real - auc_fake

cat(sprintf("\n>> PERFORMANCE DROP on FAKE videos:\n"))
cat(sprintf("   Accuracy: %.1f%% lower\n", accuracy_drop))
cat(sprintf("   ROC-AUC:  %.4f lower\n", auc_drop))
```

Statistical tests:

```{r}
# Fisher's exact test to compare accuracy between real and fake
real_correct <- sum(emotion_pred_real$table[1,1], emotion_pred_real$table[2,2])
real_incorrect <- sum(emotion_pred_real$table[1,2], emotion_pred_real$table[2,1])
fake_correct <- sum(emotion_pred_fake$table[1,1], emotion_pred_fake$table[2,2])
fake_incorrect <- sum(emotion_pred_fake$table[1,2], emotion_pred_fake$table[2,1])

contingency_emotion <- matrix(c(real_correct, real_incorrect,
                                fake_correct, fake_incorrect),
                              nrow = 2, byrow = TRUE,
                              dimnames = list(c("Real", "Fake"),
                                            c("Correct", "Incorrect")))

cat("Contingency table for accuracy:\n")
print(contingency_emotion)

fisher_emotion <- fisher.test(contingency_emotion)
cat(sprintf("\nFisher's exact test (Real vs Fake accuracy): p = %.4f\n", 
            fisher_emotion$p.value))

# McNemar's test is not appropriate here since we're comparing different videos
# But we can do a proportion test
prop_test_emotion <- prop.test(x = c(real_correct, fake_correct),
                               n = c(nrow(dta_test_real_only), nrow(dta_test_fake_only)))
cat(sprintf("Two-proportion z-test (Real vs Fake accuracy): p = %.4f\n", 
            prop_test_emotion$p.value))
```
