---
title: "final_analysis_emotion"
output:
  pdf_document: default
---

# Introduction

This Rmd will be the same as `final_analysis.Rmd` but will just use videos with emotion. Previous model performance indicated that there are differences between videos with emotion and no emotion. By splitting the problem into emotion vs. no emotion, this should reduce complexity.

# Load Libraries & Data

```{r}
library(tidyverse)
library(dplyr)

# Load data
dta_nmf_input <- read.csv("dta_nmf_input.csv")
```

Split into separate datasets by emotion:

```{r}
# Function to add emotion column
add_emotion_column <- function(df) {
  df %>%
    mutate(
      emotion = case_when(
        str_detect(video_filename, "_laughing") ~ "yes",
        str_detect(video_filename, "_happy") ~ "yes",
        str_detect(video_filename, "_serious") ~ "yes",
        str_detect(video_filename, "_angry") ~ "yes",
        str_detect(video_filename, "_disgusted") ~ "yes",
        TRUE ~ "no"
      ),
      emotion = factor(emotion, levels = c("yes", "no"))
    )
}


dta_nmf_input <- add_emotion_column(dta_nmf_input)
# Reorder columns
dta_nmf_input <- dta_nmf_input %>%
  select(video_filename, corresponding_video, scene_name, isfakeorreal, emotion, frame, everything())


# Emotion only
dta_nmf_input_emotion <- dta_nmf_input %>%
  filter(emotion == "yes")

# No Emotion only
dta_nmf_input_no_emotion <- dta_nmf_input %>%
  filter(emotion == "no")


# Video counts
print(length(unique(dta_nmf_input_emotion$video_filename)))
print(length(unique(dta_nmf_input_no_emotion$video_filename)))
```

Emotion dataset: 280 videos

No Emotion dataset: 184 videos

Save datasets:

```{r}
write.csv(dta_nmf_input_emotion, "dta_nmf_input_emotion.csv", row.names = F)
write.csv(dta_nmf_input_no_emotion, "dta_nmf_input_no_emotion.csv", row.names = F)
```

# Split 80/20

```{r}
set.seed(12345)

real_videos <- dta_nmf_input_emotion %>%
  filter(isfakeorreal == "real") %>%
  distinct(video_filename)

# Sample 80% of real videos for train set
train_real_videos <- sample(real_videos$video_filename,
                            size = floor(0.8 * nrow(real_videos)))

# Get corresponding fake videos for selected real videos
train_fake_videos <- dta_nmf_input_emotion %>%
  filter(video_filename %in% train_real_videos) %>%
  pull(corresponding_video) %>%
  unique()

# Train set
dta_nmf_input_train <- dta_nmf_input_emotion %>%
  filter(video_filename %in% c(train_real_videos, train_fake_videos))

# Test set is everything else
dta_nmf_input_test <- dta_nmf_input_emotion %>%
  filter(!video_filename %in% c(train_real_videos, train_fake_videos))


# Check distribution
print(length(unique(dta_nmf_input_train$video_filename)))
print(length(unique(dta_nmf_input_test$video_filename)))

# Check pairs are preserved
train_videos <- unique(dta_nmf_input_train$video_filename)
test_videos <- unique(dta_nmf_input_test$video_filename)

train_corresponding_videos <- dta_nmf_input_emotion %>%
  filter(video_filename %in% train_videos) %>%
  pull(corresponding_video) %>%
  unique()

if (length(intersect(train_corresponding_videos, test_videos)) == 0) {
  cat("Success.\n")
} else {
  cat("Error, the following pairs were split:\n")
  print(intersect(train_corresponding_videos, test_videos))
}
```

Training set: 224 videos.

Test set: 56 videos.

# NMF

## Optimal K

```{r}
# Installs required dependencies automatically and prevents update prompts
#remotes::install_github("zdebruine/RcppML", dependencies = TRUE, upgrade = "never")
set.seed(12345)

ranks <- 2:10
mse_values <- numeric(length(ranks))
names(mse_values) <- ranks

input_matrix <- as.matrix(t(dta_nmf_input_train[, 7:23]))

for (i in 1:length(ranks)) {
  k <- ranks[i]
  cat("Running RcppML NMF for rank =", k, "\n")
  res_rcppml <- tryCatch({
    RcppML::nmf(input_matrix,
                k = k,
                verbose = FALSE,
                tol = 1e-4,
                maxit = 200)
  }, error = function(e) {
    cat("Error running RcppML NMF for rank", k, ":", conditionMessage(e), "\n")
    return(NULL)
  })
  
  if (!is.null(res_rcppml)) {
    W <- res_rcppml$w
    H <- res_rcppml$h
    D <- res_rcppml$d
    V_hat <- W %*% diag(D) %*% H  # reconstruction
    
    # Caclulate RSS
    rss <- sum((input_matrix - V_hat)^2)
    # Calculate MSE
    mse_values[i] <- rss / (nrow(input_matrix) * ncol(input_matrix))
    cat("Rank", k, "completed. MSE:", mse_values[i], "\n")
    
  } else {
    mse_values[i] <- NA
  }
  rm(res_rcppml, W, H, V_hat); gc()
}

# Plot
dta_rank_metrics_mse <- data.frame(
  Rank = ranks,
  MSE = mse_values
)

dta_rank_metrics_mse %>%
  ggplot(aes(x = Rank, y = MSE)) +
  geom_line(color = "blue") +
  geom_point(color = "blue", size = 3) +
  scale_x_continuous(breaks = dta_rank_metrics_mse$Rank) +
  labs(title = "NMF Rank Estimation (RccpML): MSE vs Rank") +
  theme_minimal()
```

## Run NMF

```{r}
model_nmf <- RcppML::nmf(input_matrix, 
                         k = 3, 
                         tol = 1e-6, 
                         maxit = 750, 
                         verbose = TRUE)


# Extract H and scaling vector D
H <- model_nmf$h
D <- model_nmf$d

# Scale H matrix using d scaling vector
H_scaled <- sweep(H, 1, D, "*")

# Transpose H matrix to get correct row order
dta_h_matrix <- as.data.frame(t(H_scaled))

# Change column names
colnames(dta_h_matrix) <- paste0("nmf", 1:ncol(dta_h_matrix))

# Check row order in h_df matches row order in metadata
if(nrow(dta_h_matrix) != nrow(dta_nmf_input_train[, 2:5])) {
  stop("Number of samples in H matrix does not match number of rows in the metadata.")
}

# Add h to dta_nmf_input
dta_nmf_output <- dta_nmf_input_train %>%
  select(video_filename, corresponding_video, scene_name, isfakeorreal, emotion, frame) %>%
  bind_cols(dta_h_matrix)
```

## Visualisations

```{r}
# Function for min-max normalisation, used in visualisations
fn_maxnormalise <- function(x, ...) {
  rng <- range(x, na.rm = TRUE)
  if (anyNA(rng) || diff(rng) == 0) {
    return(rep(0, length(x)))  # safe fallback: all 0s
  } else {
    return((x - rng[1]) / diff(rng))
  }
}
```

```{r}
library(RColorBrewer)

au_names <- c("AU01_inner_brow_raiser", "AU02_outer_brow_raiser", "AU04_brow_lowerer",
              "AU05_upper_lid_raiser", "AU06_cheek_raiser", "AU07_lid_tightener",
              "AU09_nose_wrinkler", "AU10_upper_lip_raiser", "AU12_lip_corner_puller",
              "AU14_dimpler", "AU15_lip_corner_depressor", "AU17_chin_raiser",
              "AU20_lip_stretcher", "AU23_lip_tightener", "AU25_lips_part",
              "AU26_jaw_drop", "AU45_blink")

rownames(model_nmf$w) <- au_names


# Basis matrix
NMF::aheatmap(
  model_nmf$w, 
  main = "Basis Matrix (W)", 
  color = colorRampPalette(brewer.pal(6, "Blues"))(100), 
  Rowv = NA, Colv = NA
)
# Normalise across rows for comparison
# Group by component for normalisation
NMF::aheatmap(
  apply(model_nmf$w, 2, fn_maxnormalise), 
  main = "Basis Matrix (W) - Normalised", 
  color = colorRampPalette(brewer.pal(6, "Blues"))(100), 
  scale = "none", 
  Rowv = NA, Colv = NA
)

apply(model_nmf$w, 2, fn_maxnormalise)

# Applied fix to normalised basis matrix by changing 1 to 2 (before it was normalising by rows not columns)
```

High loading AUs:

Component 1:

-   AU06 - cheek raiser (1)

-   AU07 - lid tightener (0.75864228)

-   AU10 - upper lip raiser (0.90533776)

-   AU12 - lip corner puller (0.99279814)

-   AU14 - dimpler (0.87149821)

Component 2:

-   AU04 - brow lowerer (0.752474531)

-   AU07 - lid tightener (0.594085123)

-   AU15 - lip corner depressor (0.561341988)

-   AU 17 - chin raiser (1)

-   AU23 - lip tightener (0.513841587)

Component 3:

-   AU01 - inner brow raiser (1)

-   AU02 - outer brow raiser (0.8697229)

-   AU05 - upper lid raiser (0.6727540)

-   AU25 - lips part (0.5623135)

-   AU26 - jaw drop (0.5614809)

-   AU45 - blink (0.5261286)

```{r}
dta_nmf_output %>%
  pivot_longer(
    cols = starts_with("nmf"),
    names_to = "Component",
    values_to = "Activation"
  ) %>%
  group_by(video_filename, Component) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Average across videos for each frame & component
  group_by(isfakeorreal, Component, frame) %>%
  summarise(Activation = mean(Activation, na.rm=TRUE)) %>%
  # Normalise again by component
  group_by(Component)%>%
  mutate(Activation = fn_maxnormalise(Activation))%>%
  ungroup() %>%
  # Plot heatmap
  ggplot( aes(x = frame, y = Component, fill = Activation)) +
  geom_tile() +
  facet_wrap(~ isfakeorreal, ncol = 2) +
  scale_fill_gradientn(colours = viridis::magma(100)) +
  labs(
    title = "Heatmap of Averaged NMF Component Activations (Emotion Only Videos)",
    x = "Frame",
    y = "NMF Component",
    fill = "Activation"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 14))
```

```{r}
# Real vs. fake
dta_nmf_output %>%
  pivot_longer(
    cols = starts_with("nmf"),
    names_to = "Component",
    values_to = "Activation"
  ) %>%
  # Normalise within each video and component
  group_by(video_filename, Component) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Average across videos (by component and frame)
  group_by(isfakeorreal, Component, frame) %>%
  summarise(Activation = mean(Activation, na.rm = TRUE)) %>%
  # Normalise averaged values within each component
  group_by(Component) %>%
  mutate(Activation = fn_maxnormalise(Activation)) %>%
  ungroup() %>%
  # Plot
  ggplot(aes(x = frame, y = Activation, colour = Component)) +
  geom_smooth() +
  facet_wrap(~ isfakeorreal, ncol = 1) +
  labs(
    title = "NMF Components Over Time For Emotion Only Videos (Real vs. Fake)"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 14)  # Increase font size for facet labels
  )
```

# CMFTS

## Install CMFTS Package

```{r}
# Install dependencies
#devtools::install_github("cbergmeir/tsExpKit")

# Install CMFTS package (uncomment if first run)
#devtools::install_github("fjbaldan/CMFTS")
```

## Perform CMFTS Feature Extraction

```{r}
dta_cmfts_input <- dta_nmf_output %>%
  select(video_filename, isfakeorreal, emotion, frame, nmf1, nmf2, nmf3) %>%
  pivot_longer(cols = c("nmf1", "nmf2", "nmf3"),
               names_to = "component",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_cmfts_input), value = TRUE)

head(dta_cmfts_input)
```

Function for calculating CMFTS features:

```{r}
fn_extract_cmfts_features <- function(ts_values) {
  ts_matrix <- matrix(as.numeric(ts_values), nrow = 1)
  cmfts_features <- cmfts::cmfts(dataset = ts_matrix)
  return(cmfts_features)
}
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_cmfts_output <- apply(
  dta_cmfts_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_cmfts_input[, c("video_filename", "isfakeorreal", "emotion", "component")],
    .
  )

head(dta_cmfts_output)
```

## CMFTS Preprocessing

```{r}
dta_cmfts_output_concat <- dta_cmfts_output %>%
  pivot_wider(
    names_from = component,
    values_from = -c(video_filename, component),
    names_glue = "{.value}_{component}"
  )

# Remove duplicate columns
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -3:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_cmfts_output_concat)[2] <- c("isfakeorreal")
colnames(dta_cmfts_output_concat)[3] <- c("emotion")

head(dta_cmfts_output_concat)
```

### Missing Values

```{r}
dta_cmfts_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

We can remove these columns:

```{r}
# Drop permutation entropy columns
dta_cmfts_output_concat <- dta_cmfts_output_concat[, -13:-15]
```

### Infinite Values

```{r}
dta_cmfts_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

Only shannon entropy columns have infinite values. We can convert these to NA and then impute with missForest.

```{r}
# Convert Infs to NA
for(col in names(dta_cmfts_output_concat)) {
  dta_cmfts_output_concat[[col]][is.infinite(dta_cmfts_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_output_imputed <- dta_cmfts_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_output_imputed[, sapply(dta_cmfts_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE: 0.0005 (very low reconstruction error).

### Features with Zero SD

Features with 0 SD contribute no meaningful information and can be removed.

```{r}
sds_zerosd <- sapply(dta_cmfts_output_imputed, function(x) {
  if(is.numeric(x)) sd(x, na.rm=T) else NA
})

# List features that do not meet the threshold
names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd == 0)]

# Drop features with zero SD
dta_cmfts_output_imputed_zerosd <- dta_cmfts_output_imputed[, names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd > 0)]]
```

# Feature Selection

## Correlation

Starting with a quick correlation-based elimination is a good way to find out which columns can be removed to reduce pairwise correlations. Columns that are highly correlated with one another are likely to be redundant for the classifier.

```{r}
# Cutoff 0.9
cat("ZeroSD\n\nCutoff = 0.9\n")
caret::findCorrelation(cor(dta_cmfts_output_imputed_zerosd[, 4:114]), cutoff = 0.9)

# Cutoff 0.95
cat("\nCutoff = 0.95\n")
caret::findCorrelation(cor(dta_cmfts_output_imputed_zerosd[, 4:114]), cutoff = 0.95)
```

A cutoff of 0.9 flagged 8 features for removal due to being highly correlated with at least one other feature. A cutoff point of 0.95 flagged no features for removal, same as before.

## Recursive Feature Elimination (RFE)

```{r}
x_train_zerosd <- dta_cmfts_output_imputed_zerosd[, 4:114]
y_train_zerosd <- as.factor(dta_cmfts_output_imputed_zerosd$isfakeorreal)

# Set up RFE control using random forest (fitting params)
control <- caret::rfeControl(functions = caret::rfFuncs,
                      method = "cv",
                      number = 10,
                      verbose = TRUE)  # 10-fold cross validation

# Define subset sizes to evaluate
# Model will start with 5 predictors, incrementing by 5
subsetSizes <- seq(5, ncol(x_train_zerosd), by = 5)

# RFE
set.seed(12345)
rfeResults_zerosd <- caret::rfe(x = x_train_zerosd,
                     y = y_train_zerosd,
                     sizes = subsetSizes,
                     rfeControl = control)

# Results
print(rfeResults_zerosd)
plot(rfeResults_zerosd)
```

RFE selected an optimum feature set of 25 variables, achieving 45.31% accuracy.

## Boruta

```{r}
set.seed(12345)

# Create a subset of data including only the outcome and predictors
boruta_train_zsd <- dta_cmfts_output_imputed_zerosd[, c("isfakeorreal", 
                                                        names(dta_cmfts_output_imputed_zerosd)[4:114])]

# Make sure outcome variable is factor
boruta_train_zsd$isfakeorreal <- as.factor(boruta_train_zsd$isfakeorreal)

# Boruta
boruta_output_zsd <- Boruta::Boruta(isfakeorreal ~ ., data = boruta_train_zsd, doTrace = 2)

# Output
print(boruta_output_zsd)
plot(boruta_output_zsd, las = 2)

# Solve tentative features
Boruta::TentativeRoughFix(boruta_output_zsd)
```

Boruta initially identified no features as important, but one as tentative:

-   shannon_entropy_SG_nmf2

After roughfixing, this feature was deemed as important.

# Create Test Set

To create the test set, we need to project the NMF model onto the AU data from the test set, then derive CMFTS features from these.

## Project onto Test Set

```{r}
library(Matrix)
library(nnls)

# 1. Extract scaled basis matrix from model
project_W_train_raw <- model_nmf$w
project_d_train <- as.numeric(model_nmf$d)
project_W_train <- as.matrix(project_W_train_raw %*% Diagonal(x = project_d_train))

# 2. Set up loop to project each video

# Empty list to store final activation matrices
project_activations_list <- list()

for (video_name in unique(dta_nmf_input_test$video_filename)) {
  # Prepare matrix for current video
  project_V_new <- dta_nmf_input_test %>%
    filter(video_filename == video_name) %>%
    select(starts_with("AU")) %>%
    as.matrix()
  
  # Transpose so that features are rows and frames are columns
  project_v_new <- t(project_V_new)
  
  # Set up variables for inner NNLS loop
  project_k <- ncol(project_W_train)
  project_n_new <- ncol(project_v_new)
  project_h_new <- matrix(NA_real_, nrow = project_k, ncol = project_n_new)
  
  # Project each frame using NNLS
  for (i in seq_len(project_n_new)) {
    project_h_new[, i] <- nnls(project_W_train, project_v_new[, i])$x
  }
  
  # Apply scaling to projected H matrix
  project_h_new_scaled <- sweep(project_h_new, 1, project_d_train, "*")
  
  # Transpose final activation matrix so its frames x components
  project_h_new_transposed <- t(project_h_new_scaled)
  
  # Store result in list
  project_activations_list[[video_name]] <- project_h_new_transposed
}

cat("Projection complete.\n")
cat(paste("The list now contains", length(project_activations_list), "activation matrices.\n\n"))
cat("Dimensions of activation matrix for first video (frames x components):\n")
print(dim(project_activations_list[[1]]))
```

## Validate

```{r}
project_V_orig <- dta_nmf_input_test %>%
  #filter(video_filename == names(project_activations_list)[1]) %>%
  select(starts_with("AU")) %>%
  as.matrix()

project_H_new <- do.call(rbind, project_activations_list)

# Reconstruct
project_V_recon <- project_H_new %*% t(model_nmf$w)

# Calculate R squared
project_r2_validation <- 1 - sum((project_V_orig - project_V_recon)^2) / sum((project_V_orig - mean(project_V_orig))^2)

cat(sprintf("Reconstruction R2 for entire test set: %.4f\n", project_r2_validation))
```

R2 = 0.4892

This means that the trained NMF model can explain approximately 49% of the total variance in the unseen test set. This is marginally better than in the previous analysis (emotion and no emotion videos combined), suggesting that splitting the problem into two separate pipelines simplifies things for the model somewhat.

## Create Final Test Set

```{r}
dta_h_matrix_test <- as.data.frame(project_H_new)
colnames(dta_h_matrix_test) <- paste0("nmf", 1:ncol(dta_h_matrix_test))

dta_metadata_test <- dta_nmf_input_test %>%
  select(video_filename, corresponding_video, scene_name, isfakeorreal, emotion, frame)

dta_nmf_test_output <- bind_cols(dta_metadata_test, dta_h_matrix_test)

head(dta_nmf_test_output)
```

## Create CMFTS Features

```{r}
dta_cmfts_test_input <- dta_nmf_test_output %>%
  select(video_filename, isfakeorreal, emotion, frame, nmf1, nmf2, nmf3) %>%
  pivot_longer(cols = c("nmf1", "nmf2", "nmf3"),
               names_to = "component",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_cmfts_test_input), value = TRUE)

head(dta_cmfts_test_input)
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_cmfts_test_output <- apply(
  dta_cmfts_test_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_cmfts_test_input[, c("video_filename", "isfakeorreal", "emotion", "component")],
    .
  )

head(dta_cmfts_test_output)
```

Now to preprocess in the same way as before.

## CMFTS Preprocessing

```{r}
dta_cmfts_test_output_concat <- dta_cmfts_test_output %>%
  pivot_wider(
    names_from = component,
    values_from = -c(video_filename, component),
    names_glue = "{.value}_{component}"
  )

# Remove duplicate columns
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -3:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_cmfts_test_output_concat)[2] <- c("isfakeorreal")
colnames(dta_cmfts_test_output_concat)[3] <- c("emotion")

head(dta_cmfts_test_output_concat)
```

Missing values:

```{r}
dta_cmfts_test_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

```{r}
# Drop permutation entropy columns
dta_cmfts_test_output_concat <- dta_cmfts_test_output_concat[, -13:-15]
```

Infinite values:

```{r}
dta_cmfts_test_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

```{r}
# Convert Infs to NA
for(col in names(dta_cmfts_test_output_concat)) {
  dta_cmfts_test_output_concat[[col]][is.infinite(dta_cmfts_test_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_test_output_imputed <- dta_cmfts_test_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_test_output_imputed[, sapply(dta_cmfts_test_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_test_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE: 0.00037

Features with zero SD:

```{r}
sds_zerosd <- sapply(dta_cmfts_test_output_imputed, function(x) {
  if(is.numeric(x)) sd(x, na.rm=T) else NA
})

# List features that do not meet the threshold
names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd == 0)]

# Drop features with zero SD
dta_cmfts_test_output_imputed_zerosd <- dta_cmfts_test_output_imputed[, names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd > 0)]]
```

The test set is now ready: `dta_cmfts_test_output_imputed_zerosd`.

# Classifiers

I will create two models for each classifier.

-   Model 1 = Only features confirmed as important by Boruta (n=2)

-   Model 2 = Same feature types but across all components (n=6)

The reason for this stems from the idea that the three NMF components represent interrelated aspects of facial expression, so if a feature type was most discriminative for one component, its behaviour in other components might still hold relevant information for classification.

```{r}
set.seed(12345)

# Train set (UNCOMMENT FOR IMBALANCED EMOTION)
dta_cmfts_output_imputed_zerosd$isfakeorreal <- factor(
  dta_cmfts_output_imputed_zerosd$isfakeorreal,
  levels = c("real", "fake")
)

# dta_cmfts_emotion_balanced$isfakeorreal <- factor(
#   dta_cmfts_emotion_balanced$isfakeorreal,
#   levels = c("real", "fake")
# )

boruta_selected_features <- c(
  "shannon_entropy_SG_nmf2"
)

# Select these features plus the outcome variable
dta_classifier_train_new <- dta_cmfts_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

# dta_classifier_train_new <- dta_cmfts_emotion_balanced %>%
#   select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

base_feature_types <- c(
  "shannon_entropy_SG")

all_component_features <- character()
for (base_feat in base_feature_types) {
  for (comp_num in 1:3) {
    all_component_features <- c(all_component_features, paste0(base_feat, "_nmf", comp_num))
  }
}

dta_classifier_train_new_all <- dta_cmfts_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(all_component_features))

# dta_classifier_train_new_all <- dta_cmfts_emotion_balanced %>%
#   select(video_filename, emotion, isfakeorreal, all_of(all_component_features))

dta_classifier_train_new$video_filename <- dta_cmfts_output_imputed_zerosd$video_filename

# dta_classifier_train_new$video_filename <- dta_cmfts_emotion_balanced$video_filename



# Test set
dta_cmfts_test_output_imputed_zerosd$isfakeorreal <- factor(
  dta_cmfts_test_output_imputed_zerosd$isfakeorreal,
  levels = c("real", "fake")
)

dta_classifier_test_new <- dta_cmfts_test_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

dta_classifier_test_new_all <- dta_cmfts_test_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(all_component_features))

dta_classifier_test_new$video_filename <- dta_cmfts_test_output_imputed_zerosd$video_filename
```

## Random Forest

### Model 1

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "none",
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

rf_fit_model1 <- caret::train(
  isfakeorreal ~ ., # nmf
  data = dta_classifier_train_new[, 3:4],
  method = "rf",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(mtry = 4)
)

print(rf_fit_model1)
plot(caret::varImp(rf_fit_model1)) # To see feature importance
```

Evaluate on unseen data:

```{r}
rf_model_1_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model1, newdata = dta_classifier_test_new),
  reference = dta_classifier_test_new$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_performance)
```

Accuracy = 48.21%

95% CI = (0.3466, 0.6197)

NIR = 0.5

P = 0.6556

Kappa = -0.0357

Sensitivity = 46.43%

Specificity = 50%

### Model 2

```{r}
# Cross validation setup
classifier_control <- caret::trainControl(
  method = "none",
  classProbs = TRUE,
  summaryFunction = caret::twoClassSummary
)

set.seed(12345)

rf_fit_model2 <- caret::train(
  isfakeorreal ~ ., # nmf
  data = dta_classifier_train_new_all[, 3:6],
  method = "rf",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(mtry = 4)
)

print(rf_fit_model2)
plot(caret::varImp(rf_fit_model2)) # To see feature importance
```

Evaluate on unseen data:

```{r}
rf_model_2_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model2, newdata = dta_classifier_test_new_all),
  reference = dta_classifier_test_new_all$isfakeorreal,
  positive = "fake"
)

print(rf_model_2_performance)
```

Accuracy = 51.79%

95% CI = (0.3803, 0.6534)

NIR = 0.5

P = 0.4469

Kappa = 0.0357

Sensitivity = 50%

Specificity = 53.57%

## C5.0B

### Model 1

```{r}
set.seed(12345)

c50_fit_model1 <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new[, 3:4],
  method = "C5.0",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(
    model = "tree",
    trials = 20,  # number of boosting iterations
    winnow = FALSE
  )
)

print(c50_fit_model1)
plot(varImp(c50_fit_model1)) # To see feature importanc
```

Evaluate on unseen data:

```{r}
c50_model_1_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model1, newdata = dta_classifier_test_new),
  reference = dta_classifier_test_new$isfakeorreal,
  positive = "fake"
)

print(c50_model_1_performance)
```

Accuracy = 58.93%

95% CI = (0.4498, 0.719)

NIR = 0.5

P = 0.1144

Kappa = 0.1786

Sensitivity = 67.86%

Specificity = 50%

## Model 2

```{r}
set.seed(12345)

c50_fit_model2 <- caret::train(
  isfakeorreal ~ .,
  data = dta_classifier_train_new_all[, 3:6],
  method = "C5.0",
  metric = "ROC",
  trControl = classifier_control,
  tuneGrid = expand.grid(
    model = "tree",
    trials = 20,  # number of boosting iterations
    winnow = FALSE
  )
)

print(c50_fit_model2)
plot(varImp(c50_fit_model2)) # To see feature importanc
```

Evaluate on unseen data:

```{r}
c50_model_2_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model2, newdata = dta_classifier_test_new_all),
  reference = dta_classifier_test_new_all$isfakeorreal,
  positive = "fake"
)

print(c50_model_2_performance)
```

Accuracy = 58.93%

95% CI = (0.4498, 0.719)

NIR = 0.5

P = 0.1144

Sensitivity = 67.86%

Specificity = 50%

Same as Model 1.

# 40 Held-Out Videos

I will test all models on the 40 held-out videos from the human participant experiment as another test set. I will do this first for the entire 40 videos and then for just emotion only.

But first, I will have to derive the CMFTS features for these videos by projecting the NMF model first.

## Project NMF to AU Data

```{r}
dta_video_au_40 <- read.csv("../human_data_preprocess/dta_video_au_final.csv")

# UNCOMMENT FOR EMOTION VIDEOS ONLY
dta_video_au_40 <- dta_video_au_40 %>%
  filter(emotion_yes_no_new == "y")

library(Matrix)
library(nnls)

# 1. Extract scaled basis matrix from model
project_W_train_raw <- model_nmf$w
project_d_train <- as.numeric(model_nmf$d)
project_W_train <- as.matrix(project_W_train_raw %*% Diagonal(x = project_d_train))

# 2. Set up loop to project each video

# Empty list to store final activation matrices
project_40_activations_list <- list()

for (video_name in unique(dta_video_au_40$stimnopath)) {
  # Prepare matrix for current video
  project_V_new <- dta_video_au_40 %>%
    filter(stimnopath == video_name) %>%
    select(starts_with("au")) %>%
    as.matrix()
  
  # Transpose so that features are rows and frames are columns
  project_v_new <- t(project_V_new)
  
  # Set up variables for inner NNLS loop
  project_k <- ncol(project_W_train)
  project_n_new <- ncol(project_v_new)
  project_h_new <- matrix(NA_real_, nrow = project_k, ncol = project_n_new)
  
  # Project each frame using NNLS
  for (i in seq_len(project_n_new)) {
    project_h_new[, i] <- nnls(project_W_train, project_v_new[, i])$x
  }
  
  # Apply scaling to projected H matrix
  project_h_new_scaled <- sweep(project_h_new, 1, project_d_train, "*")
  
  # Transpose final activation matrix so its frames x components
  project_h_new_transposed <- t(project_h_new_scaled)
  
  # Store result in list
  project_40_activations_list[[video_name]] <- project_h_new_transposed
}

cat("Projection complete.\n")
cat(paste("The list now contains", length(project_40_activations_list), "activation matrices.\n\n"))
cat("Dimensions of activation matrix for first video (frames x components):\n")
print(dim(project_40_activations_list[[1]]))
```

## Validation

```{r}
project_40_V_orig <- dta_video_au_40 %>%
  #filter(video_filename == names(project_activations_list)[1]) %>%
  select(starts_with("AU")) %>%
  as.matrix()

project_40_H_new <- do.call(rbind, project_40_activations_list)

# Reconstruct
project_40_V_recon <- project_40_H_new %*% t(model_nmf$w)

# Calculate R squared
project_40_r2_validation <- 1 - sum((project_40_V_orig - project_40_V_recon)^2) / sum((project_40_V_orig - mean(project_40_V_orig))^2)

cat(sprintf("Reconstruction R2 for entire test set: %.4f\n", project_40_r2_validation))
```

R2 (All 40 videos) = 0.3702

R2 (Only emotion) = 0.4078

This suggests that performance is better when we stick with emotion only videos.

## Create Final Test Set

```{r}
dta_40_h_matrix_test <- as.data.frame(project_40_H_new)
colnames(dta_40_h_matrix_test) <- paste0("nmf", 1:ncol(dta_40_h_matrix_test))

dta_40_metadata_test <- dta_video_au_40 %>%
  select(stimnopath, isfakeorreal, emotion_yes_no_new, frame)

dta_40_nmf_test_output <- bind_cols(dta_40_metadata_test, dta_40_h_matrix_test)

head(dta_40_nmf_test_output)
```

## CMFTS Features

```{r}
# Make sure max frames is 241
dta_40_nmf_test_output <- dta_40_nmf_test_output %>%
  filter(frame < 242)


dta_cmfts_40_test_input <- dta_40_nmf_test_output %>%
  select(stimnopath, isfakeorreal, emotion_yes_no_new, frame, nmf1, nmf2, nmf3) %>%
  pivot_longer(cols = c("nmf1", "nmf2", "nmf3"),
               names_to = "component",
               values_to = "activation") %>%
  pivot_wider(names_from = frame,
              values_from = activation,
              names_prefix = "fr_")

frame_cols <- grep("^fr_", names(dta_cmfts_40_test_input), value = TRUE)

head(dta_cmfts_40_test_input)
```

```{r}
# Apply feature extraction row-wise over the frame columns
dta_cmfts_40_test_output <- apply(
  dta_cmfts_40_test_input[, frame_cols], 
  1, 
  fn_extract_cmfts_features
  ) %>%
  do.call(rbind, .) %>%
  as.data.frame() %>%
  cbind(
    dta_cmfts_40_test_input[, c("stimnopath", "isfakeorreal", "emotion_yes_no_new", "component")],
    .
  )

head(dta_cmfts_40_test_output)
```

## CMFTS Preprocessing

```{r}
dta_cmfts_40_test_output_concat <- dta_cmfts_40_test_output %>%
  pivot_wider(
    names_from = component,
    values_from = -c(stimnopath, component),
    names_glue = "{.value}_{component}"
  )

# Remove duplicate columns
dta_cmfts_40_test_output_concat <- dta_cmfts_40_test_output_concat[, -3:-6]
# Rename isfakeorreal & emotion columns
colnames(dta_cmfts_40_test_output_concat)[2] <- c("isfakeorreal")
colnames(dta_cmfts_40_test_output_concat)[3] <- c("emotion")
colnames(dta_cmfts_40_test_output_concat)[1] <- c("video_filename")

head(dta_cmfts_40_test_output_concat)
```

Missing values:

```{r}
dta_cmfts_40_test_output_concat %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "column",
               values_to = "n_missing") %>%
  filter(n_missing > 0)
```

```{r}
# Drop permutation entropy columns
dta_cmfts_40_test_output_concat <- dta_cmfts_40_test_output_concat[, -13:-15]
```

Infinite values:

```{r}
dta_cmfts_40_test_output_concat %>%
  summarise(across(
    .cols = where(is.numeric),
    .fns  = ~ sum(is.infinite(.))
  )) %>%
  pivot_longer(
    cols       = everything(),
    names_to   = "column",
    values_to  = "n_infinite"
  )
```

```{r}
# Convert Infs to NA
for(col in names(dta_cmfts_40_test_output_concat)) {
  dta_cmfts_40_test_output_concat[[col]][is.infinite(dta_cmfts_40_test_output_concat[[col]]) ] <- NA
}

set.seed(12345)

# Copy dta_cmfts_output_concat
dta_cmfts_40_test_output_imputed <- dta_cmfts_40_test_output_concat

# Get all numeric columns into a matrix
tmp <- as.matrix(dta_cmfts_40_test_output_imputed[, sapply(dta_cmfts_40_test_output_imputed, is.numeric)])
tmp[!is.finite(tmp)] <- NA

# Run missForest
impute_res <- missForest::missForest(tmp,
                         ntree = 100,
                         maxiter = 10,
                         verbose = TRUE)

dta_cmfts_40_test_output_imputed[, colnames(tmp)] <- impute_res$ximp

impute_res$OOBerror
```

NRMSE (Emotion only) = 0.00088

Same NRMSE for all videos.

Features with zero SD:

```{r}
sds_zerosd <- sapply(dta_cmfts_40_test_output_imputed, function(x) {
  if(is.numeric(x)) sd(x, na.rm=T) else NA
})

# List features that do not meet the threshold
names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd == 0)]

# Drop features with zero SD
dta_cmfts_40_test_output_imputed_zerosd <- dta_cmfts_40_test_output_imputed[, names(sds_zerosd)[is.na(sds_zerosd) | (sds_zerosd > 0)]]
```

The final dataset for this is ready: `dta_cmfts_40_test_output_imputed_zerosd`.

## Model Predictions

First to refine the test set so that it only has the relevant features.

```{r}
# Test set
dta_cmfts_40_test_output_imputed_zerosd$isfakeorreal <- factor(
  dta_cmfts_40_test_output_imputed_zerosd$isfakeorreal,
  levels = c("real", "fake")
)

dta_classifier_40_test_new <- dta_cmfts_40_test_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(boruta_selected_features))

dta_classifier_40_test_new_all <- dta_cmfts_40_test_output_imputed_zerosd %>%
  select(video_filename, emotion, isfakeorreal, all_of(all_component_features))

dta_classifier_40_test_new$video_filename <- dta_cmfts_40_test_output_imputed_zerosd$video_filename

head(dta_classifier_40_test_new_all)
```

### RF Model 1

```{r}
rf_model_1_40_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model1, newdata = dta_classifier_40_test_new),
  reference = dta_classifier_40_test_new$isfakeorreal,
  positive = "fake"
)

print(rf_model_1_40_performance)
```

Emotion videos only (n=16):

Accuracy = 50%

95% CI = (0.2465, 0.7535)

NIR = 0.5

P = 0.5982

Kappa = 0

Sensitivity = 50%

Specificity = 50%

Precision = 50%

F1 = 50%

All videos (n=40):

Accuracy = 47.5%

95% CI = (0.3151, 0.6387)

NIR = 0.5

P = 0.6821

Kappa = -0.05

Sensitivity = 55%

Specificity = 40%

Precision = 47.83%

F1 = 51.16%

### RF Model 2

```{r}
rf_model_2_40_performance <- caret::confusionMatrix(
  data = predict(rf_fit_model2, newdata = dta_classifier_40_test_new_all),
  reference = dta_classifier_40_test_new_all$isfakeorreal,
  positive = "fake"
)

print(rf_model_2_40_performance)
```

Emotion videos only (n=16):

-   Same as model 1.

All videos (n=40):

Accuracy = 50%

95% CI = (0.338, 0.662)

NIR = 0.5

P = 0.5627

Kappa = 0

Sensitivity = 55%

Specificity = 45%

Precision = 50%

F1 = 52.38%

### C5.0B Model 1

```{r}
c50_model_1_40_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model1, newdata = dta_classifier_40_test_new),
  reference = dta_classifier_40_test_new$isfakeorreal,
  positive = "fake"
)

print(c50_model_1_40_performance)
```

Emotion videos only (n=16):

Accuracy = 62.50%

95% CI = (0.3543, 0.848)

NIR = 0.5

P = 0.2272

Kappa = 0.25

Sensitivity = 75%

Specificity = 50%

Precision = 60%

F1 = 66.67%

All videos (n=40):

Accuracy = 47.50%

95% CI = (0.3151, 0.6387)

NIR = 0.5

P = 0.6821

Kappa = -0.05

Sensitivity = 65%

Specificity = 30%

Precision = 48.15%

F1 = 55.32%

### C5.0B Model 2

```{r}
c50_model_2_40_performance <- caret::confusionMatrix(
  data = predict(c50_fit_model2, newdata = dta_classifier_40_test_new_all),
  reference = dta_classifier_40_test_new_all$isfakeorreal,
  positive = "fake"
)

print(c50_model_2_40_performance)
```

Emotion videos only (n=16):

-   Same as model 1.

All videos (n=40):

-   Same as model 1.
