---
title: "movement_analysis"
output:
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

Previous sensitivity analysis (`final_analysis_NMF_check_valence_2.Rmd`)
found that there are different degradation patterns between real and
fake videos depending on emotion (positive, neutral, negative). The main
goal of that analysis was to determine whether the finding from the main
script (`final_analysis_NMF_check.Rmd`) that real vs fake classifying
performance was better in emotion videos was due to emotional
expressions having more movement or due to genuine degradation in
emotional signal.

However, this is still unsolved. This current script aims to directly
test whether there are differences in facial movement between real and
fake videos, as well as between different emotions.

# Load Libraries & Data

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
# Load data
dta_preprocessed <- read.csv("../../CSV files/OpenFace files/full_dataset_preprocessed.csv")
dta_nmf_input_train <- read.csv("../../CSV files/dta_nmf_input_train.csv")
dta_nmf_input_test <- read.csv("../../CSV files/dta_nmf_input_test.csv")
```

# Movement Metrics

## Frame-Level

```{r}
# Function to calculate derivatives
fn_calculate_derivatives <- function(x) {
  # Velocity (first derivative)
  velocity <- c(0, diff(x))
  
  # Acceleration (second derivative)
  acceleration <- c(0, diff(velocity))
  
  # Jerk (third derivative)
  jerk <- c(0, diff(acceleration))
  
  return(list(
    velocity = velocity,
    acceleration = acceleration,
    jerk = jerk
  ))
}

au_cols <- names(dta_nmf_input_train)[grepl("^AU", names(dta_nmf_input_train))]

# Calculate derivatives for each AU per video
dta_frame_movement_train <- dta_nmf_input_train %>%
  group_by(video_filename) %>%
  mutate(
    # For each AU, calculate derivatives
    across(all_of(au_cols),
           ~ calculate_derivatives(.)$velocity,
           .names = "vel_{col}"),
    across(all_of(au_cols),
           ~ abs(calculate_derivatives(.)$velocity),  # Absolute velocity
           .names = "abs_vel_{col}"),
    across(all_of(au_cols),
           ~ calculate_derivatives(.)$acceleration,
           .names = "acc_{col}"),
    across(all_of(au_cols),
           ~ abs(calculate_derivatives(.)$acceleration),
           .names = "abs_acc_{col}"),
    across(all_of(au_cols),
           ~ calculate_derivatives(.)$jerk,
           .names = "jerk_{col}")
  ) %>%
  ungroup()

# Calculate total movement per frame (sum of absolute velocities across all AUs)
vel_cols <- names(dta_frame_movement_train)[grepl("^abs_vel_AU", names(dta_frame_movement_train))]

dta_frame_movement_train <- dta_frame_movement_train %>%
  mutate(
    total_movement = rowSums(select(., all_of(vel_cols)), na.rm = TRUE),
    total_acceleration = rowSums(select(., starts_with("abs_acc_AU")), na.rm = TRUE)
  )
```

We now have the following metrics:

-   Velocity

-   Absolute velocity

-   Acceleration

-   Absolute acceleration

-   Jerk

-   Total movement (sum of all absolute velocity)

-   Total acceleration

Absolute values were calculated for velocity and acceleration because
they can have a negative and positive value, which can sometimes lead to
cancelling out and making it look like no movement occurred (e.g., +10,
then -10 = 0, suggesting the face never moved).

## Video-Level

```{r}
# Calculate video-level summaries
dta_video_movement_train <- dta_frame_movement_train %>%
  group_by(video_filename, corresponding_video, 
           scene_name, isfakeorreal, emotion, valence) %>%
  summarise(
    # Global movement metrics
    mean_movement = mean(total_movement, na.rm = TRUE),
    sd_movement = sd(total_movement, na.rm = TRUE),
    max_movement = max(total_movement, na.rm = TRUE),
    movement_95th = quantile(total_movement, 0.95, na.rm = TRUE),
    
    # Movement variability (coefficient of variation)
    movement_cv = sd_movement / mean_movement,
    
    # Peak movement frequency (proportion of frames exceeding mean + 2*SD)
    threshold = mean(total_movement, na.rm = TRUE) + 2 * sd(total_movement, na.rm = TRUE),
    peak_freq = sum(total_movement > threshold, na.rm = TRUE) / n(),
    
    # Acceleration metrics
    mean_acceleration = mean(total_acceleration, na.rm = TRUE),
    max_acceleration = max(total_acceleration, na.rm = TRUE),
    
    .groups = 'drop'
  ) %>%
  select(-threshold)

# Calculate movement smoothness (spectral arc length)
fn_calculate_smoothness <- function(movement_signal) {
  if(length(movement_signal) < 10) return(NA)
  
  # Compute FFT
  fft_result <- abs(fft(movement_signal))
  
  # Spectral arc length (simplified version)
  arc_length <- sum(sqrt(1 + diff(fft_result[1:(length(fft_result)/2)])^2))
  
  # Normalize by signal length
  return(-log(arc_length / length(movement_signal)))  # Higher = smoother
}

# Add smoothness metric
dta_smoothness_train <- dta_frame_movement_train %>%
  group_by(video_filename) %>%
  summarise(
    movement_smoothness = fn_calculate_smoothness(total_movement),
    .groups = 'drop'
  )

dta_video_movement_train <- dta_video_movement_train %>%
  left_join(dta_smoothness_train, by = "video_filename")

cat("Video-level movement metrics calculated.\n")
cat("Total videos:", nrow(dta_video_movement_train), "\n")
cat("Real videos:", sum(dta_video_movement_train$isfakeorreal == "real"), "\n")
cat("Fake videos:", sum(dta_video_movement_train$isfakeorreal == "fake"), "\n")
```

Now also have the following video-level movement metrics:

-   mean_movement (avg of total movement across all frames)

-   max_movement (highest single spike in entire video)

-   movement_95th (95th percentile of movement - upper limit of normal
    high activity for that video)

-   sd_movement (how much movement varies from avg)

-   movement_cv (coefficient of variation - how inconsistent movement is
    relative to how much movement there is).

-   peak_freq (percentage of frames where movement exceeds threshold of
    mean+2SD)

-   mean_acceleration

-   max_acceleration

-   movement_smoothness (calculated using FFT and spectral arc length,
    looks at the frequency of movement signal. A high value = smooth,
    biological movement. Low value = jittery, noisy or robotic movement)

# Paired T-Tests

```{r}
dta_video_movement_train <- dta_video_movement_train %>%
  mutate(
    # Remove file extensions (e.g., .mp4, .avi) to ensure matches
    clean_filename = str_remove(video_filename, "\\.[^.]+$"),
    clean_corresponding = str_remove(corresponding_video, "\\.[^.]+$"),
    
    # LOGIC: 
    # If it's REAL, its pair ID is its own name.
    # If it's FAKE, its pair ID is the corresponding source video.
    pair_id = ifelse(isfakeorreal == "real", 
                     clean_filename, 
                     clean_corresponding)
  )

# 2. Check if the IDs actually overlap now
real_ids <- dta_video_movement_train %>% filter(isfakeorreal == "real") %>% pull(pair_id)
fake_ids <- dta_video_movement_train %>% filter(isfakeorreal == "fake") %>% pull(pair_id)

cat("Unique Real IDs:", length(unique(real_ids)), "\n")
cat("Unique Fake IDs:", length(unique(fake_ids)), "\n")
cat("Intersecting IDs:", length(intersect(real_ids, fake_ids)), "\n")

# 3. Perform the Split and Join using the new 'pair_id'
dta_real <- dta_video_movement_train %>% 
  filter(isfakeorreal == "real") %>%
  select(pair_id, everything()) %>%
  rename_with(~paste0(., "_real"), -c(pair_id))

dta_fake <- dta_video_movement_train %>% 
  filter(isfakeorreal == "fake") %>%
  select(pair_id, everything()) %>%
  rename_with(~paste0(., "_fake"), -c(pair_id))

# Join strictly on pair_id
dta_paired_train <- inner_join(dta_real, dta_fake, by = "pair_id")

cat("\n=== PAIRING DIAGNOSTICS (ATTEMPT 2) ===\n")
cat("Successfully Paired:", nrow(dta_paired_train), "\n")
```

Now, `dta_paired_train` has two of the same columns for everything,
except one is for the real video and the other is for the fake video in
a pair.

## Real vs. Fake

```{r}
cat("\n=== OVERALL PAIRED T-TESTS (Real vs Fake) ===\n\n")

# Test mean movement
t_mean <- t.test(dta_paired_train$mean_movement_real, 
                 dta_paired_train$mean_movement_fake, 
                 paired = TRUE)
cat("Mean Movement:\n")
cat(sprintf("  Real: %.3f ± %.3f\n", 
            mean(dta_paired_train$mean_movement_real, na.rm = TRUE),
            sd(dta_paired_train$mean_movement_real, na.rm = TRUE)))
cat(sprintf("  Fake: %.3f ± %.3f\n", 
            mean(dta_paired_train$mean_movement_fake, na.rm = TRUE),
            sd(dta_paired_train$mean_movement_fake, na.rm = TRUE)))
cat(sprintf("  Difference: %.3f (95%% CI: [%.3f, %.3f])\n",
            t_mean$estimate, t_mean$conf.int[1], t_mean$conf.int[2]))
cat(sprintf("  t(%d) = %.2f, p = %.4f\n\n", 
            t_mean$parameter, t_mean$statistic, t_mean$p.value))

# Test movement variability
t_cv <- t.test(dta_paired_train$movement_cv_real, 
               dta_paired_train$movement_cv_fake, 
               paired = TRUE)
cat("Movement Variability (CV):\n")
cat(sprintf("  Real: %.3f ± %.3f\n", 
            mean(dta_paired_train$movement_cv_real, na.rm = TRUE),
            sd(dta_paired_train$movement_cv_real, na.rm = TRUE)))
cat(sprintf("  Fake: %.3f ± %.3f\n", 
            mean(dta_paired_train$movement_cv_fake, na.rm = TRUE),
            sd(dta_paired_train$movement_cv_fake, na.rm = TRUE)))
cat(sprintf("  Difference: %.3f (95%% CI: [%.3f, %.3f])\n",
            t_cv$estimate, t_cv$conf.int[1], t_cv$conf.int[2]))
cat(sprintf("  t(%d) = %.2f, p = %.4f\n\n", 
            t_cv$parameter, t_cv$statistic, t_cv$p.value))

# Test peak frequency
t_peak <- t.test(dta_paired_train$peak_freq_real, 
                 dta_paired_train$peak_freq_fake, 
                 paired = TRUE)
cat("Peak Movement Frequency:\n")
cat(sprintf("  Real: %.3f ± %.3f\n", 
            mean(dta_paired_train$peak_freq_real, na.rm = TRUE),
            sd(dta_paired_train$peak_freq_real, na.rm = TRUE)))
cat(sprintf("  Fake: %.3f ± %.3f\n", 
            mean(dta_paired_train$peak_freq_fake, na.rm = TRUE),
            sd(dta_paired_train$peak_freq_fake, na.rm = TRUE)))
cat(sprintf("  Difference: %.3f (95%% CI: [%.3f, %.3f])\n",
            t_peak$estimate, t_peak$conf.int[1], t_peak$conf.int[2]))
cat(sprintf("  t(%d) = %.2f, p = %.4f\n\n", 
            t_peak$parameter, t_peak$statistic, t_peak$p.value))
```

Mean movement:

-   No statistical difference in average speed of the face between real
    and fake.

Movement variability (CV):

-   Highly significant (p \< 0.0001) Real: 0.690 \| Fake: 0.643

-   Real videos have a significantly higher Coefficient of Variation
    than fake videos.

-   CV = SD/Mean, and since we know that the means are the same, we know
    that real videos have a higher SD.

-   Real humans are dynamic, we have a lot of variability in our
    movement. In fakes, their movement variability hovers closer to the
    average movement more often than in real videos. They lack the full
    dynamic range of a real human. They are unnaturally consistent.

Peak movement frequency:

-   Not significant, both groups (real + fake) cross the high movement
    threshold (mean + 2SD) about 4.5% of the time.

-   Deepfakes do not have sudden glitch spikes in their facial movement
    (sudden huge movement jumps).

-   It could be possible that there are glitches in movement but they
    are micro ones and do not cross the threshold. Or its possible that
    deepfakes are quite smooth.

Overall, these findings suggest that the emotional signal is not getting
degraded, it is infact being preserved (mean movement finding). What is
actually happening is signal homogenisation, where the faceswap is
preserving the average motion but smoothing out the variability. The
degradation is actually a loss of dynamic contrast.

```{r}
cat("\n=== MISSING PAIRED T-TESTS (Force & Intensity) ===\n")

# 1. Test Mean Acceleration
t_mean_acc <- t.test(dta_paired_train$mean_acceleration_real, 
                     dta_paired_train$mean_acceleration_fake, 
                     paired = TRUE)
cat("Mean Acceleration:\n")
cat(sprintf("  Real: %.3f\n", mean(dta_paired_train$mean_acceleration_real, na.rm=T)))
cat(sprintf("  Fake: %.3f\n", mean(dta_paired_train$mean_acceleration_fake, na.rm=T)))
cat(sprintf("  Diff: %.3f | p-value: %.4f %s\n\n", 
            t_mean_acc$estimate, t_mean_acc$p.value,
            ifelse(t_mean_acc$p.value < 0.05, "*SIG*", "")))

# 2. Test Max Acceleration
t_max_acc <- t.test(dta_paired_train$max_acceleration_real, 
                    dta_paired_train$max_acceleration_fake, 
                    paired = TRUE)
cat("Max Acceleration:\n")
cat(sprintf("  Real: %.3f\n", mean(dta_paired_train$max_acceleration_real, na.rm=T)))
cat(sprintf("  Fake: %.3f\n", mean(dta_paired_train$max_acceleration_fake, na.rm=T)))
cat(sprintf("  Diff: %.3f | p-value: %.4f %s\n\n", 
            t_max_acc$estimate, t_max_acc$p.value,
            ifelse(t_max_acc$p.value < 0.05, "*SIG*", "")))

# 3. Test Max Movement
t_max_mov <- t.test(dta_paired_train$max_movement_real, 
                    dta_paired_train$max_movement_fake, 
                    paired = TRUE)
cat("Max Movement:\n")
cat(sprintf("  Real: %.3f\n", mean(dta_paired_train$max_movement_real, na.rm=T)))
cat(sprintf("  Fake: %.3f\n", mean(dta_paired_train$max_movement_fake, na.rm=T)))
cat(sprintf("  Diff: %.3f | p-value: %.4f %s\n\n", 
            t_max_mov$estimate, t_max_mov$p.value,
            ifelse(t_max_mov$p.value < 0.05, "*SIG*", "")))

# 4. Test 95th Percentile Movement
t_95_mov <- t.test(dta_paired_train$movement_95th_real, 
                   dta_paired_train$movement_95th_fake, 
                   paired = TRUE)
cat("95th Percentile Movement:\n")
cat(sprintf("  Real: %.3f\n", mean(dta_paired_train$movement_95th_real, na.rm=T)))
cat(sprintf("  Fake: %.3f\n", mean(dta_paired_train$movement_95th_fake, na.rm=T)))
cat(sprintf("  Diff: %.3f | p-value: %.4f %s\n\n", 
            t_95_mov$estimate, t_95_mov$p.value,
            ifelse(t_95_mov$p.value < 0.05, "*SIG*", "")))
```

Acceleration:

-   Mean acceleration not significant. Suggests that the fakes are not
    jittery or have sudden movement snaps.

-   Max acceleration is also not significant.

Max movement:

-   Significantly lower in fakes compared to real videos.

-   Fakes have less extreme movement peaks. When an huge emotional
    expression is made in a real video, the fake fails to reach the full
    intensity of the real movement.

95th percentile movement:

-   Significantly lower in fake compared to real again.

-   Fakes cut off the top 5% of the expression intensity.

Overall, these findings suggest that the faceswap autoencoder largely
preserves the dynamics of facial expressions, but the big movements are
muted. This gives a potential explanation for why the classifier works
better on emotion videos. In neutral videos, the face isn't moving a
lot, so there are no large peaks (max movement). In emotion videos,
there are bigger facial movements. The faceswap process fails to reach
the peak of these movements, and this difference is what the classifier
is likely detecting.

## Emotion

```{r}
# Get list of unique valences (usually Positive, Neutral, Negative)
valences <- unique(dta_paired_train$valence_real)

# Loop through each valence and run the tests
for (v in valences) {
  
  # 1. Filter data for just this valence
  dta_subset <- dta_paired_train %>% filter(valence_real == v)
  
  cat(paste0("\n=========================================\n"))
  cat(paste0("   ANALYSIS FOR VALENCE: ", toupper(v), "\n"))
  cat(paste0("   (n = ", nrow(dta_subset), " pairs)\n"))
  cat(paste0("=========================================\n"))

  # --- TEST 1: Mean Movement ---
  if(nrow(dta_subset) > 2) {
    t_mean <- t.test(dta_subset$mean_movement_real, 
                     dta_subset$mean_movement_fake, 
                     paired = TRUE)
    
    cat("\n1. Mean Movement:\n")
    cat(sprintf("   Real: %.3f (SD: %.3f)\n", mean(dta_subset$mean_movement_real, na.rm=T), sd(dta_subset$mean_movement_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f (SD: %.3f)\n", mean(dta_subset$mean_movement_fake, na.rm=T), sd(dta_subset$mean_movement_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_mean$estimate, t_mean$p.value, 
                ifelse(t_mean$p.value < 0.05, "*SIG*", "")))
  }

  # --- TEST 2: Movement Variability (CV) ---
  if(nrow(dta_subset) > 2) {
    t_cv <- t.test(dta_subset$movement_cv_real, 
                   dta_subset$movement_cv_fake, 
                   paired = TRUE)
    
    cat("\n2. Movement Variability (CV):\n")
    cat(sprintf("   Real: %.3f (SD: %.3f)\n", mean(dta_subset$movement_cv_real, na.rm=T), sd(dta_subset$movement_cv_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f (SD: %.3f)\n", mean(dta_subset$movement_cv_fake, na.rm=T), sd(dta_subset$movement_cv_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_cv$estimate, t_cv$p.value, 
                ifelse(t_cv$p.value < 0.05, "*SIG*", "")))
  }

  # --- TEST 3: Peak Frequency ---
  if(nrow(dta_subset) > 2) {
    t_peak <- t.test(dta_subset$peak_freq_real, 
                     dta_subset$peak_freq_fake, 
                     paired = TRUE)
    
    cat("\n3. Peak Frequency:\n")
    cat(sprintf("   Real: %.3f\n", mean(dta_subset$peak_freq_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f\n", mean(dta_subset$peak_freq_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_peak$estimate, t_peak$p.value, 
                ifelse(t_peak$p.value < 0.05, "*SIG*", "")))
  }
  
  # --- TEST 4: Smoothness ---
  if(nrow(dta_subset) > 2) {
    valid_smooth <- dta_subset %>% filter(!is.na(movement_smoothness_real) & !is.na(movement_smoothness_fake))
    
    if(nrow(valid_smooth) > 2) {
      t_smooth <- t.test(valid_smooth$movement_smoothness_real, 
                         valid_smooth$movement_smoothness_fake, 
                         paired = TRUE)
      
      cat("\n4. Movement Smoothness:\n")
      cat(sprintf("   Real: %.3f\n", mean(valid_smooth$movement_smoothness_real)))
      cat(sprintf("   Fake: %.3f\n", mean(valid_smooth$movement_smoothness_fake)))
      cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                  t_smooth$estimate, t_smooth$p.value, 
                  ifelse(t_smooth$p.value < 0.05, "*SIG*", "")))
    } else {
      cat("\n4. Movement Smoothness: Not enough valid data points.\n")
    }
  }

  # --- TEST 5: Mean Acceleration ---
  if(nrow(dta_subset) > 2) {
    t_mean_acc <- t.test(dta_subset$mean_acceleration_real, 
                         dta_subset$mean_acceleration_fake, 
                         paired = TRUE)
    
    cat("\n5. Mean Acceleration:\n")
    cat(sprintf("   Real: %.3f\n", mean(dta_subset$mean_acceleration_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f\n", mean(dta_subset$mean_acceleration_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_mean_acc$estimate, t_mean_acc$p.value,
                ifelse(t_mean_acc$p.value < 0.05, "*SIG*", "")))
  }

  # --- TEST 6: Max Acceleration ---
  if(nrow(dta_subset) > 2) {
    t_max_acc <- t.test(dta_subset$max_acceleration_real, 
                        dta_subset$max_acceleration_fake, 
                        paired = TRUE)
    
    cat("\n6. Max Acceleration:\n")
    cat(sprintf("   Real: %.3f\n", mean(dta_subset$max_acceleration_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f\n", mean(dta_subset$max_acceleration_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_max_acc$estimate, t_max_acc$p.value,
                ifelse(t_max_acc$p.value < 0.05, "*SIG*", "")))
  }

  # --- TEST 7: Max Movement ---
  if(nrow(dta_subset) > 2) {
    t_max_mov <- t.test(dta_subset$max_movement_real, 
                        dta_subset$max_movement_fake, 
                        paired = TRUE)
    
    cat("\n7. Max Movement:\n")
    cat(sprintf("   Real: %.3f\n", mean(dta_subset$max_movement_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f\n", mean(dta_subset$max_movement_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_max_mov$estimate, t_max_mov$p.value,
                ifelse(t_max_mov$p.value < 0.05, "*SIG*", "")))
  }

  # --- TEST 8: 95th Percentile ---
  if(nrow(dta_subset) > 2) {
    t_95 <- t.test(dta_subset$movement_95th_real, 
                   dta_subset$movement_95th_fake, 
                   paired = TRUE)
    
    cat("\n8. 95th Percentile Movement:\n")
    cat(sprintf("   Real: %.3f\n", mean(dta_subset$movement_95th_real, na.rm=T)))
    cat(sprintf("   Fake: %.3f\n", mean(dta_subset$movement_95th_fake, na.rm=T)))
    cat(sprintf("   Diff: %.3f | p-value: %.4f %s\n", 
                t_95$estimate, t_95$p.value,
                ifelse(t_95$p.value < 0.05, "*SIG*", "")))
  }

  cat("\n\n")
}
```

Neutral:

-   Fakes are significantly smoother (movement smoothness) and have
    significantly lower variability (CV) than real.

-   A real face has small movements that the faceswap process smooths
    out. The face becomes too perfect and unnaturally stable.

-   Real videos have significantly higher peak movements (max movements)
    than fakes.

-   Mean movement had no significant difference, suggesting the volume
    of movement is the same but the quality (complexity) is lost.

-   The faceswap autoencoder acts as a low-pass filter by removing
    subtle biological signals.

Negative:

-   Fakes fail a lot here with significant differences for all tests.

-   Real faces move more (mean movement), have more average acceleration
    (mean acceleration), snap/twitch much faster (max acceleration), and
    have deform much further (max movement).

-   The faceswap autoencoder essentially creates a lazy version of the
    negative expression because it cannot accurately preserve. Negative
    expressions are known to be very subtle and nuanced, and since we
    know already that the fake process is smoothing out a lot of subtle
    signal, we can see it do the same here.

Positive:

-   Positive expressions typically involve large deformations/movements,
    but unlike negative emotions, the faceswap process seems to add
    energy rather than remove it.

-   Fakes have significantly higher mean movement and mean acceleration
    than real videos.

-   This could be due to positive expressions having more jitter in
    fakes, where the textures jitter or vibrate when the faceswap
    process struggles to latch onto the facial landmarks due to their
    extreme movements.

Overall, we know that the faceswap process smooths out subtle movements
in facial expressions. There is a loss of dynamic range in facial
movements in fake compared to real videos, which is likely what the
classifier picks up on to detect fakes.

# Mixed Effect Model

The paired t-tests so far have investigated whether there is a
difference between real and fake movement, but a mixed effect model will
tell us whether the difference depends on emotion, while controlling for
video pairs (some actors may be more naturally expressive than others).

The model below tests whether the real/fake difference in movement
metrics depend on emotional valence, while accounting for the paired
structure of the data.

```{r}
library(lme4)
library(lmerTest)

cat("\n=== MIXED EFFECTS MODEL: Movement Variability (CV) ===\n\n")

# Change baseline reference to neutral
dta_video_movement_train$valence <- relevel(factor(dta_video_movement_train$valence), ref = "neutral")

model_cv <- lmer(movement_cv ~ isfakeorreal * valence + (1| pair_id),
                 data = dta_video_movement_train)

summary(model_cv)

# Type III ANOVA table (tests each fixed effect)
cat("\nANOVA Table:\n")
anova(model_cv)


cat("\n=== MIXED EFFECTS MODEL: Max Movement ===\n\n")

model_max <- lmer(max_movement ~ isfakeorreal * valence + (1 | pair_id),
                  data = dta_video_movement_train)

summary(model_max)
anova(model_max)


cat("\n=== MIXED EFFECTS MODEL: 95th Percentile Movement ===\n\n")

model_95th <- lmer(movement_95th ~ isfakeorreal * valence + (1 | pair_id),
                   data = dta_video_movement_train)

summary(model_95th)
anova(model_95th)


cat("\n=== MIXED EFFECTS MODEL: Mean Movement ===\n\n")

model_mean <- lmer(mean_movement ~ isfakeorreal * valence + (1 | pair_id),
                   data = dta_video_movement_train)

summary(model_mean)
anova(model_mean)
```

Movement Variability (CV):

-   Intercept represents fake neutral videos (CV = 0.625).

-   Real videos have significantly higher CV overall (+0.057, p =
    0.001), confirming fakes are unnaturally consistent.

-   Interaction with positive valence is borderline although not
    significant (p = 0.07). The real/fake gap in CV is smaller for
    positive emotions than for neutral.

-   Negative emotions show no significant interaction (real/fake gap is
    similar to neutral).

Max movement:

-   Real videos have higher max movement overall (+0.445, p = 0.048).

-   Positive interaction is borderline significant and negative (-0.628,
    p = 0.053), meaning the real/fake gap in peak movement is reduced
    (or even reversed) for positive expressions.

-   This aligns with the earlier finding that fakes seem to
    overcompensate during positive expressions.

95th percentile movement:

-   Similar pattern to max movement. Real videos trend higher (+0.195, p
    = 0.069) but not significant.

-   Positive interaction is again negative (-0.263, p = 0.088) but not
    significant.

-   Overall interaction is highly significant (p = 0.007), driven by the
    divergent patterns across emotions.

Mean movement:

-   Main effect is non-significant (p = 0.208), meaning that on average
    real and fake videos have the same amount of movement.

-   But, interaction is highly significant (p = 0.002). Specifically,
    the real:positive coefficient is -0.098 (p = 0.015), suggesting that
    for positive emotions, fakes actually have more mean movement than
    real videos. This is the overcompensation effect seen earlier.

Overall, the faceswap autoencoder appears to be performing some kind of
movement intensity compression towards the mean. For neutral videos, the
fakes lose subtle micro-movements (lower CV, smoother), but mean
movement is preserved. For negative videos, fakes underperform - real
faces reach higher peaks and have more variability but in fakes, the
autoencoder cannot capture the nuanced and subtle expressions). For
positive videos, fakes overcompensate - they add movement/jitter,
possibly because the extreme deformations of positive expressions
compared to neutral/negative (wide smiles, raised brows, jaw drop) cause
the autoencoder to struggle with texture stability.

The classifier likely detects fakes better in emotion videos not because
they have more movement, but because fakes compress the dynamic range in
opposite directions depending on valence - they are too flat for
negative emotions, and too noisy for positive ones.

# Post-Hoc Contrasts

The interaction terms from the mixed effects model tell us that patterns
differ, but not the exact pairwise comparisons. This section will find
the estimated real/fake difference at each valence level.

```{r}
library(emmeans)

# For each model, extract the real-fake contrast at each valence

??emmeans
emm_cv <- emmeans(model_cv, pairwise ~ valence | isfakeorreal, adjust = "bonf")
emm_max <- emmeans(model_max, pairwise ~ valence | isfakeorreal, adjust = "bonf")
emm_95th <- emmeans(model_95th, pairwise ~ valence | isfakeorreal, adjust = "bonf")
emm_mean <- emmeans(model_mean, pairwise ~ valence | isfakeorreal, adjust = "bonf")

cat("\n=== POST-HOC: Movement Variability (CV) ===\n")
summary(emm_cv$contrasts)

cat("\n=== POST-HOC: Max Movement ===\n")
summary(emm_max$contrasts)

cat("\n=== POST-HOC: 95th Percentile Movement ===\n")
summary(emm_95th$contrasts)

cat("\n=== POST-HOC: Mean Movement ===\n")
summary(emm_mean$contrasts)
```

## Get Effect Sizes

This gets the standardised effect sizes (Cohens D) for the real/fake
difference at each valence.

```{r}

fn_get_cohens_d <- function(model, emm_contrasts) {
  residual_sd <- sigma(model)
  contrasts_df <- as.data.frame(summary(emm_contrasts))
  contrasts_df$cohens_d <- contrasts_df$estimate / residual_sd
  return(contrasts_df[, c("valence", "estimate", "SE", "p.value", "cohens_d")])
}

cat("\n=== EFFECT SIZES (Cohen's d) ===\n")
cat("\nMovement CV:\n")
print(fn_get_cohens_d(model_cv, emm_cv$contrasts))

cat("\nMax Movement:\n")
print(fn_get_cohens_d(model_max, emm_max$contrasts))
```

Negative:

-   Across all 4 metrics, negative emotional expressions show the
    largest and most significant real/fake gaps.

-   CV: d = -0.90 (large effect), p = 0.0002

-   Max movement: d = -0.71 (medium-large), p = 0.003

-   95th percentile: p = 0.0008

-   Mean movement: p = 0.012

-   Real negative faces have more variability, higher peaks, and more
    overall movement than fakes. The autoencoder essentially is creating
    a dampened version of negative expressions, where it cannot
    reproduce the subtleties of negative expressions.

Neutral:

-   Significant differences for CV (d = -0.54, p = 0.001) and max
    movement (d = -0.32, p = 0.048), with 95th percentile movement
    borderline (p = 0.069).

-   Mean movement shows no difference (p = 0.208).

-   This suggests fakes preserve the amount of movement but lose the
    intensity of the movement (the micro-movements and dynamic
    variability). This confirms earlier findings.

Positive:

-   No significant differences in CV, max movement, or 95th percentile.

-   mean movement is significantly higher in fakes (+0.063, p = 0.031).

-   The autoencoder isn't failing to reach the peaks - it is actually
    adding movement (likely jitter or instability).

```{r}
# Residual plots for key model
par(mfrow = c(2, 2))
plot(model_cv, main = "CV Model Diagnostics")

# QQ plot
qqnorm(resid(model_cv))
qqline(resid(model_cv))
```

# ICC

The intra-class correlation (ICC) to show how much variance is explained
by the pairing structure (how much of the variance is down to which
pair/actor it is).

```{r}
# ICC = variance_pair / (variance_pair + variance_residual)
calc_icc <- function(model) {
  vc <- as.data.frame(VarCorr(model))
  icc <- vc$vcov[1] / sum(vc$vcov)
  return(icc)
}

cat("\n=== INTRA-CLASS CORRELATIONS ===\n")
cat(sprintf("CV Model ICC: %.3f\n", calc_icc(model_cv)))
cat(sprintf("Max Movement ICC: %.3f\n", calc_icc(model_max)))
cat(sprintf("95th Percentile ICC: %.3f\n", calc_icc(model_95th)))
cat(sprintf("Mean Movement ICC: %.3f\n", calc_icc(model_mean)))
```

-   Mean movement ICC = 0.84

    -   84% of variance in mean movement is explained by video pair
        identity.

    -   Some actors move more than others, and this is strongly
        preserved across pairs.

    -   Who the actor/pair is almost completely determines the average
        amount of movement.

-   CV ICC = 0.58

    -   Moderate clustering. Individual expressiveness matters but there
        is also substantial within-pair variation that affects CV.

-   Max / 95th ICC = \~0.5

    -   About half the variance is pair-level, the other half is
        real-fake/emotion driven.

Overall, the autoencoder appears to work like a selective compression
filter, where it preserves gross movement quantity (mean movement) but
compresses the dynamic range of the movement. This compression effect is
asymmetric by valence - negative emotions are compressed towards
flatness, whereas positive emotions are compressed towards the mean but
with added noise (overcompensation).

# Figures

```{r}
library(ggplot2)
library(emmeans)
library(patchwork)

# Define consistent colors
colors_auth <- c("real" = "#2E86AB", "fake" = "#E94F37")


# ============================================
# FIGURE 1: Interaction Plot (Estimated Marginal Means)
# Shows the real-fake gap at each valence level
# ============================================

# Extract emmeans for plotting
emm_cv_plot <- as.data.frame(emmeans(model_cv, ~ isfakeorreal * valence))
emm_max_plot <- as.data.frame(emmeans(model_max, ~ isfakeorreal * valence))
emm_mean_plot <- as.data.frame(emmeans(model_mean, ~ isfakeorreal * valence))
emm_95th_plot <- as.data.frame(emmeans(model_95th, ~ isfakeorreal * valence))

# Add metric labels
emm_cv_plot$metric <- "Movement Variability (CV)"
emm_max_plot$metric <- "Max Movement"
emm_mean_plot$metric <- "Mean Movement"
emm_95th_plot$metric <- "95th Percentile Movement"

# Combine all
emm_all <- bind_rows(emm_cv_plot, emm_max_plot, emm_mean_plot, emm_95th_plot)

# Reorder valence for logical display
emm_all$valence <- factor(emm_all$valence, levels = c("negative", "neutral", "positive"))

# Rename authenticity for cleaner legend
emm_all$Authenticity <- factor(emm_all$isfakeorreal, 
                                levels = c("fake", "real"),
                                labels = c("Fake", "Real"))

# Create the interaction plot
fig_interaction <- ggplot(emm_all, aes(x = valence, y = emmean, 
                                        color = Authenticity, group = Authenticity)) +
  geom_line(linewidth = 0.8, position = position_dodge(0.1)) +
  geom_point(size = 3, position = position_dodge(0.1)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.15, linewidth = 0.6, 
                position = position_dodge(0.1)) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_color_manual(values = colors_auth) +
  scale_x_discrete(labels = c("Negative", "Neutral", "Positive")) +
  labs(
    x = "Emotional Valence",
    y = "Estimated Marginal Mean",
    title = "Real-Fake Movement Differences Depend on Emotional Content"
  ) +
  theme_classic()

print(fig_interaction)


# ============================================
# FIGURE 2: Contrast Plot (Real - Fake Difference)
# Shows the magnitude and direction of the real-fake gap
# ============================================

# Extract contrasts
contrast_cv <- as.data.frame(summary(emm_cv$contrasts))
contrast_max <- as.data.frame(summary(emm_max$contrasts))
contrast_mean <- as.data.frame(summary(emm_mean$contrasts))
contrast_95th <- as.data.frame(summary(emm_95th$contrasts))

contrast_cv$metric <- "Movement Variability (CV)"
contrast_max$metric <- "Max Movement"
contrast_mean$metric <- "Mean Movement"
contrast_95th$metric <- "95th Percentile Movement"

contrasts_all <- bind_rows(contrast_cv, contrast_max, contrast_mean, contrast_95th)

# Clean up valence labels
contrasts_all$valence <- factor(contrasts_all$valence, 
                                 levels = c("negative", "neutral", "positive"),
                                 labels = c("Negative", "Neutral", "Positive"))

# Add significance stars
contrasts_all$sig <- case_when(
  contrasts_all$p.value < 0.001 ~ "***",
  contrasts_all$p.value < 0.01 ~ "**",
  contrasts_all$p.value < 0.05 ~ "*",
  contrasts_all$p.value < 0.1 ~ "†",
  TRUE ~ ""
)

# Calculate confidence intervals (estimate ± 1.96*SE for 95% CI)
contrasts_all$lower <- contrasts_all$estimate - 1.96 * contrasts_all$SE
contrasts_all$upper <- contrasts_all$estimate + 1.96 * contrasts_all$SE

# Create contrast plot
fig_contrasts <- ggplot(contrasts_all, aes(x = valence, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.5) +
  geom_point(size = 3, color = "#2E86AB") +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.2, linewidth = 0.6, color = "#2E86AB") +
  geom_text(aes(label = sig, y = upper + 0.05 * max(abs(c(lower, upper)))), 
            size = 4, vjust = 0) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  labs(
    x = "Emotional Valence",
    y = "Real − Fake Difference (EMM)",
    title = "Magnitude of Real-Fake Gap by Emotion",
    subtitle = "Positive values indicate Real > Fake; † p < .10, * p < .05, ** p < .01, *** p < .001"
  ) +
  theme_classic() +
  theme(plot.subtitle = element_text(size = 9, color = "gray40"))

print(fig_contrasts)


# ============================================
# FIGURE 3 (Optional): Raw Data with Model Overlay
# Shows individual data points with model estimates
# ============================================

# Prepare raw data
dta_video_movement_train$valence <- factor(dta_video_movement_train$valence, 
                                            levels = c("negative", "neutral", "positive"))
dta_video_movement_train$Authenticity <- factor(dta_video_movement_train$isfakeorreal,
                                                 levels = c("fake", "real"),
                                                 labels = c("Fake", "Real"))

fig_raw_cv <- ggplot(dta_video_movement_train, 
                      aes(x = valence, y = movement_cv, color = Authenticity)) +
  geom_point(alpha = 0.3, position = position_jitterdodge(jitter.width = 0.15, 
                                                           dodge.width = 0.4),
             size = 1.5) +
  geom_point(data = emm_cv_plot %>% 
               mutate(valence = factor(valence, levels = c("negative", "neutral", "positive")),
                      Authenticity = factor(isfakeorreal, levels = c("fake", "real"), 
                                            labels = c("Fake", "Real"))),
             aes(y = emmean), 
             size = 4, position = position_dodge(0.4)) +
  geom_errorbar(data = emm_cv_plot %>% 
                  mutate(valence = factor(valence, levels = c("negative", "neutral", "positive")),
                         Authenticity = factor(isfakeorreal, levels = c("fake", "real"), 
                                               labels = c("Fake", "Real"))),
                aes(y = emmean, ymin = lower.CL, ymax = upper.CL),
                width = 0.15, linewidth = 0.8, position = position_dodge(0.4)) +
  scale_color_manual(values = colors_auth) +
  scale_x_discrete(labels = c("Negative", "Neutral", "Positive")) +
  labs(
    x = "Emotional Valence",
    y = "Movement Variability (CV)",
    title = "Movement Variability by Authenticity and Emotion"
  ) +
  theme_classic()

print(fig_raw_cv)

```

Paired t-test figure:

```{r}
# ============================================
# FIGURE: Paired T-Test Results Summary
# Shows real vs fake means with significance for each metric × valence
# ============================================

# First, calculate summary statistics from your paired data
# We need means and SEs for real and fake at each valence level

# Create a function to summarize one metric
summarize_metric <- function(data, metric_name, real_col, fake_col) {
  data %>%
    group_by(valence_real) %>%
    summarise(
      real_mean = mean(.data[[real_col]], na.rm = TRUE),
      real_se = sd(.data[[real_col]], na.rm = TRUE) / sqrt(n()),
      fake_mean = mean(.data[[fake_col]], na.rm = TRUE),
      fake_se = sd(.data[[fake_col]], na.rm = TRUE) / sqrt(n()),
      n = n(),
      .groups = 'drop'
    ) %>%
    mutate(metric = metric_name) %>%
    rename(valence = valence_real)
}

# Summarize each metric
summary_cv <- summarize_metric(dta_paired_train, "Movement Variability (CV)", 
                                "movement_cv_real", "movement_cv_fake")
summary_max <- summarize_metric(dta_paired_train, "Max Movement", 
                                 "max_movement_real", "max_movement_fake")
summary_95th <- summarize_metric(dta_paired_train, "95th Percentile Movement", 
                                  "movement_95th_real", "movement_95th_fake")
summary_mean <- summarize_metric(dta_paired_train, "Mean Movement", 
                                  "mean_movement_real", "mean_movement_fake")

# Combine all summaries
summary_all <- bind_rows(summary_cv, summary_max, summary_95th, summary_mean)

# Pivot to long format for plotting
summary_long <- summary_all %>%
  pivot_longer(
    cols = c(real_mean, fake_mean),
    names_to = "authenticity",
    values_to = "mean"
  ) %>%
  mutate(
    se = ifelse(authenticity == "real_mean", real_se, fake_se),
    authenticity = ifelse(authenticity == "real_mean", "Real", "Fake")
  ) %>%
  select(valence, metric, authenticity, mean, se, n)

# Order factors
summary_long$valence <- factor(summary_long$valence, 
                                levels = c("negative", "neutral", "positive"),
                                labels = c("Negative", "Neutral", "Positive"))
summary_long$authenticity <- factor(summary_long$authenticity, 
                                     levels = c("Fake", "Real"))
summary_long$metric <- factor(summary_long$metric,
                               levels = c("Movement Variability (CV)", 
                                          "Mean Movement",
                                          "Max Movement", 
                                          "95th Percentile Movement"))

# Add significance annotations from post-hoc results
sig_annotations <- data.frame(
  metric = factor(c(
    # CV
    "Movement Variability (CV)", "Movement Variability (CV)", "Movement Variability (CV)",
    # Mean Movement
    "Mean Movement", "Mean Movement", "Mean Movement",
    # Max Movement
    "Max Movement", "Max Movement", "Max Movement",
    # 95th Percentile
    "95th Percentile Movement", "95th Percentile Movement", "95th Percentile Movement"
  ), levels = c("Movement Variability (CV)", "Mean Movement", 
                "Max Movement", "95th Percentile Movement")),
  valence = factor(rep(c("Negative", "Neutral", "Positive"), 4),
                    levels = c("Negative", "Neutral", "Positive")),
  p_value = c(
    # CV
    0.0002, 0.001, 0.485,
    # Mean Movement
    0.012, 0.208, 0.031,
    # Max Movement
    0.003, 0.048, 0.430,
    # 95th Percentile
    0.0008, 0.069, 0.537
  )
)

sig_annotations <- sig_annotations %>%
  mutate(
    sig_label = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      p_value < 0.1 ~ "†",
      TRUE ~ "ns"
    )
  )

# Calculate y positions for significance bars
y_positions <- summary_long %>%
  group_by(metric, valence) %>%
  summarise(y_max = max(mean + se), .groups = 'drop') %>%
  left_join(sig_annotations, by = c("metric", "valence"))

# Define colors
colors_auth <- c("Real" = "#2E86AB", "Fake" = "#E94F37")

# Create the plot
fig_paired_ttest <- ggplot(summary_long, 
                            aes(x = valence, y = mean, fill = authenticity)) +
  geom_bar(stat = "identity", position = position_dodge(0.8), 
           width = 0.7, color = "black", linewidth = 0.3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(0.8), width = 0.2, linewidth = 0.4) +
  geom_text(data = y_positions,
            aes(x = valence, y = y_max * 1.08, label = sig_label, fill = NULL),
            size = 4, vjust = 0) +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = colors_auth, name = "Authenticity") +
  labs(
    x = "Emotional Valence",
    y = "Mean ± SE",
    title = "Facial Movement Metrics by Authenticity and Emotional Valence",
    caption = "† p < .10, * p < .05, ** p < .01, *** p < .001 (paired t-tests)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "gray95", color = NA),
    plot.title = element_text(face = "bold", size = 13, hjust = 0.5),
    plot.caption = element_text(size = 9, color = "gray40", hjust = 0),
    axis.title = element_text(size = 11),
    panel.spacing = unit(1, "lines")
  )

print(fig_paired_ttest)


# ============================================
# ALTERNATIVE: Difference Plot (Real - Fake)
# Shows the contrast directly with CIs
# ============================================

# Create contrast data from post-hoc results
contrast_data <- data.frame(
  metric = factor(c(
    rep("Movement Variability (CV)", 3),
    rep("Mean Movement", 3),
    rep("Max Movement", 3),
    rep("95th Percentile Movement", 3)
  ), levels = c("Movement Variability (CV)", "Mean Movement", 
                "Max Movement", "95th Percentile Movement")),
  valence = factor(rep(c("Negative", "Neutral", "Positive"), 4),
                    levels = c("Negative", "Neutral", "Positive")),
  # Note: contrasts were fake - real, so we flip signs for real - fake
  estimate = -c(
    # CV
    -0.0946, -0.0568, -0.0123,
    # Mean Movement
    -0.1038, -0.0353, 0.0631,
    # Max Movement
    -0.9917, -0.4449, 0.1831,
    # 95th Percentile
    -0.5334, -0.1950, 0.0683
  ),
  se = c(
    # CV
    0.0248, 0.0170, 0.0176,
    # Mean Movement
    0.0409, 0.0280, 0.0289,
    # Max Movement
    0.3274, 0.2238, 0.2315,
    # 95th Percentile
    0.156, 0.107, 0.110
  ),
  p_value = c(
    # CV
    0.0002, 0.001, 0.485,
    # Mean Movement
    0.012, 0.208, 0.031,
    # Max Movement
    0.003, 0.048, 0.430,
    # 95th Percentile
    0.0008, 0.069, 0.537
  )
)

contrast_data <- contrast_data %>%
  mutate(
    lower = estimate - 1.96 * se,
    upper = estimate + 1.96 * se,
    sig_label = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      p_value < 0.1 ~ "†",
      TRUE ~ ""
    ),
    # Color by direction
    direction = ifelse(estimate > 0, "Real > Fake", "Fake > Real")
  )

fig_contrasts <- ggplot(contrast_data, aes(x = valence, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.5) +
  geom_point(aes(color = direction), size = 3.5) +
  geom_errorbar(aes(ymin = lower, ymax = upper, color = direction), 
                width = 0.15, linewidth = 0.7) +
  geom_text(aes(label = sig_label, y = upper + 0.08 * max(abs(c(lower, upper)))),
            size = 4.5, fontface = "bold") +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_color_manual(values = c("Real > Fake" = "#2E86AB", "Fake > Real" = "#E94F37"),
                     name = "Direction") +
  labs(
    x = "Emotional Valence",
    y = "Real − Fake Difference",
    title = "Direction and Magnitude of Real-Fake Movement Differences",
    caption = "Points show mean difference ± 95% CI. † p < .10, * p < .05, ** p < .01, *** p < .001"
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "gray95", color = NA),
    plot.title = element_text(face = "bold", size = 13, hjust = 0.5),
    plot.caption = element_text(size = 9, color = "gray40", hjust = 0),
    axis.title = element_text(size = 11),
    panel.spacing = unit(1, "lines")
  )

print(fig_contrasts)

```
